### 2014-01-20 paper/letter.tex
First version of the paper

\documentclass[]{letter}

\usepackage{graphicx}

\usepackage[margin=1.375in]{geometry}

%% \topmargin 0.0cm
%% \oddsidemargin 0.0cm
%% \evensidemargin 0.0cm
%% \textwidth 5.75in
%% \textheight 9.25in

%% \usepackage{hyperref}

\signature{Eric G. Hunsberger}
\address{Centre for Theoretical Neuroscience\\200 University Ave. W.\\Waterloo, ON, Canada N2L 3G1}
\date{\today}

\begin{document}
\begin{letter}{Editoral Board\\Neural Computation}
\opening{Dear Sir or Madam:}

I am writing on behalf of myself and my co-authors, Matthew Scott and Chris Eliasmith, to present you with a manuscript that we have prepared for publication in Neural Computation, entitled ``Heterogeneity increases information transmission of neuronal populations''. This article examines the effects of heterogeneity on information transmission in populations of simulated neurons, and finds that it has a role to play in increasing the amount of information a population is able to encode and transmit about an input signal. We compare these results with previous results in the stochastic resonance field detailing how noise can benefit information transmission in neural populations, and find that heterogeneity has similar effects to noise in this respect.

This research is important to the field because it advances the view that heterogeneity has a role to play in neural systems, and explains in detail why this is so. Research in stochastic resonance has done a great service by documenting the beneficial effects of noise in simulated, in vitro, and in vivo neural systems, but this same service has not yet been extended to heterogeneity. The goal of this manuscript is to provide theoretical foundations for understanding the mechanisms by which heterogeneity operates, and why it is important to neural systems. These foundations tie together the limited amount of existing research on the subject, and provide direction and motivation for future theoretical and experimental research into heterogeneity in neural systems.

Given the general interest of this paper and its potential to inspire future research in the area, we ask that you consider the enclosed manuscript for publication in Neural Computation. None of the material in the article has been published or is under consideration for publication elsewhere. We thank you for your time, and look forward to your reply.

\closing{Sincerely,}

\end{letter}
\end{document}

### 2014-01-20 paper/manuscript.tex
First version of the paper

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{times}
\usepackage{graphicx}
\usepackage{color}
\usepackage{multirow}
%
\usepackage[authoryear]{natbib}
%
\usepackage{rotating}
\usepackage{bbm}
\usepackage{latexsym}
%\DeclareGraphicsExtensions{.eps,.png}

%%% margins 
\textheight 23.4cm
\textwidth 14.65cm
\oddsidemargin 0.375in
\evensidemargin 0.375in
\topmargin  -0.55in
%
\renewcommand{\baselinestretch}{2}
%
\interfootnotelinepenalty=10000
%
\renewcommand{\thesubsubsection}{\arabic{section}.\arabic{subsubsection}}
\newcommand{\myparagraph}[1]{\ \\{\em #1}.\ \ }
\newcommand{\citealtt}[1]{\citeauthor{#1},\citeyear{#1}}
\newcommand{\myycite}[1]{\citep{#1}}

% Different font in captions
\newcommand{\captionfonts}{\normalsize}

\makeatletter  
\long\def\@makecaption#1#2{%
  \renewcommand{\baselinestretch}{1}% EH: stop captions running off the bottom of the page
  \vskip\abovecaptionskip
  \sbox\@tempboxa{{\captionfonts #1: #2}}%
  \ifdim \wd\@tempboxa >\hsize
    {\captionfonts #1: #2\par}
  \else
    \hbox to\hsize{\hfil\box\@tempboxa\hfil}%
  \fi
  \vskip\belowcaptionskip}
\makeatother   
%%%%%

\newcommand{\ie}{\emph{i.e., }}
\newcommand{\eg}{\emph{e.g., }}
%% \newcommand{\cf}{\emph{cf. }}
%% \newcommand{\etc}{\emph{etc.}}
%% \newcommand{\wrt}{w.r.t. }

\usepackage{nameref}
\newcommand{\fig}[1]{Figure \ref{fig:#1}}
\newcommand{\figlabel}[1]{\label{fig:#1}}
\newcommand{\eqn}[1]{Equation \ref{eqn:#1}}
\newcommand{\eqnlabel}[1]{\label{eqn:#1}}
\newcommand{\scn}[1]{\textsc{\nameref{scn:#1}}}
\newcommand{\scnlabel}[1]{\label{scn:#1}}

\renewcommand{\thefootnote}{\normalsize \arabic{footnote}}

\begin{document}
\hspace{13.9cm}1

\ \vspace{20mm}\\

{\LARGE Heterogeneity increases information transmission of neuronal populations}

\ \\
{\bf \large Eric G. Hunsberger$^{\displaystyle 1}$, Matthew Scott$^{\displaystyle 2}$, Chris Eliasmith$^{\displaystyle 1}$}\\
{$^{\displaystyle 1}$Centre for Theoretical Neuroscience, University of Waterloo, Waterloo, Ontario, Canada.}\\
{$^{\displaystyle 2}$Department of Applied Mathematics, University of Waterloo, Waterloo, Ontario, Canada.}\\
%

%\ \\[-2mm]
{\bf Keywords:} Heterogeneity; stochastic resonance; population coding

\thispagestyle{empty}
\markboth{}{NC instructions}
%
\ \vspace{-0mm}\\
%
%Abstract
\begin{center} {\bf Abstract} \end{center}
Noise---specifically random fluctuations added to the membrane voltages of a neuronal population---has been shown to increase the amount of information a neuronal population can encode about an input signal; this phenomenon is one type of stochastic resonance. We demonstrate that heterogeneity within a population of simulated neurons---specifically the random variation of membrane bias currents across a population---produces similar benefits as additive noise in terms of increasing information encoding within the population. Furthermore, we find that both noise and heterogeneity use a common set of three mechanisms to increase information transmission: 1) they temporally desynchronize the firing of neurons within the population, 2) they decrease the response time of the population of neurons to sudden changes in the input signal, and 3) they linearize the response of a population to a stimulus. We examine each mechanism in detail, describing how these mechanisms operate in neuronal populations, and demonstrating how they operate in populations of simulated FitzHugh-Nagumo and leaky integrate-and-fire neurons. These three mechanisms work to distribute information equally across all neurons in the population, both in terms of signal timing (desynchronization and response time) and in terms of signal amplitude (linearization). Without noise or heterogeneity, all neurons in the population encode the same aspects of the input signal; adding noise or heterogeneity allows all neurons to encode \emph{complementary} aspects of the input signal, thereby preserving more information. Overall, the simulations detailed in this paper highlight the importance of heterogeneity in neuronal population coding, and demonstrate that it can be equally effective as stochastic resonance in increasing information transmission. 
%%%%%%%%%%%

\section{Introduction}
\scnlabel{introduction}

The term stochastic resonance (SR) refers to the beneficial effects of noise, specifically the phenomenon whereby a non-zero amount of stochastic fluctuation over time in the system improves some measure of performance \citep{McDonnell2009}. Since its conception, the idea of stochastic resonance has developed into a field of its own, a field which has found increasing relevance for neuroscience.

The early stochastic resonance literature dealt mainly with \emph{subthreshold} stochastic resonance \citep{Gammaitoni1998}. In a deterministic setting, systems which only respond to inputs above a certain threshold---such as neurons---cannot transmit any information about subthreshold input signals, \ie input signals which are entirely below the system's threshold. In contrast, a noisy input signal (\ie a signal with an added stochastic offset) will occasionally cross the system's threshold, and thereby allows the system to transmit more information about the input signal. Too much noise, though, causes the input signal to become lost in random fluctuations. A moderate amount of noise---not too much, and not too little---is needed to optimize information transmission by the system about the input signal \citep{Wiesenfeld1994,Longtin1998}.

A new form of stochastic resonance---called \emph{suprathreshold} stochastic resonance (SSR)---has emerged in the past decade \citep{McDonnell2009}. In neuroscience, suprathreshold SR looks at the benefits of noise to populations of nonlinear neurons receiving a common signal of which a significant portion is above the systems' firing threshold(s). This concept was first introduced by \cite{Stocks2000}, who looked at groups of binary threshold units, and found that noise can increase information transmission in such systems. Subsequent research has examined the SSR effect in populations of more sophisticated simulated neurons \citep{Stocks2001}.

Heterogeneity among neurons---the fact that not all neurons are identical, and vary across a wide range of parameters---is a topic that has largely been ignored in the stochastic resonance literature, and neuroscience literature in general \citep{Machens2010}. It is well known that there are large differences between neurons in different parts of the nervous system, and that there are many different types of neurons, but research into the role of heterogeneity within a group of neurons of the same type is limited (\cite{Eliasmith2003} is a notable exception). In the field of stochastic resonance, the effects of noise in a heterogeneous population have been examined, but generally in specific cases such as for input signals much larger than the range of heterogeneity \citep{Stocks2000}, or for very large levels of noise \citep{McDonnell2006}. Furthermore, research into stochastic resonance with heterogeneity has focused on populations of binary threshold units, which fail to capture the dynamics of actual neurons.

The purpose of this paper is to compare the benefits of noise and the benefits of heterogeneity on information transmission by a population of simulated neurons. While previous studies have looked at the effects of noise, often claiming that noise is the central mechanism of SR and SSR effects \citep{Collins1995,Stocks2001} (see \cite{McDonnell2011} for a review), we are not aware of any work that has examined the effects of heterogeneity in similar circumstances. As we show, many effects of noise can also be achieved with heterogeneity. We argue that this is at least partly because of several shared effects on neural coding. We use both the FitzHugh-Nagumo (FHN) neuron model, a common model in SR research \citep{Wiesenfeld1994,Collins1995,Longtin1998,Stocks2001}, as well as the leaky integrate-and-fire (LIF) neuron model, a typical model of cortical neurons \citep{Koch1999}. We use a common aperiodic random signal $s(t)$ to drive populations of FHN and LIF neurons with varying levels of internal noise and heterogeneity, and compare the similarity of the decoded output signal with the input signal using mutual information. Stochastic offset is modeled as zero-mean Gaussian white noise with standard deviation $\sigma_\eta$, independently and identically distributed across all neurons (see \scn{methods}). Heterogeneity was added by randomly selecting the bias voltages $b_i$ from a uniform interval of varying width. We report how noise and heterogeneity both act to increase information transmission through these populations. They achieve this using three common mechanisms---linearizing the response of a population of neurons to a stimulus, desynchronizing neuronal firing, and decreasing the reaction time of a population of neurons to a stimulus---which we examine in detail.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\scnlabel{results}

We performed numerical experiments analyzing how noise and heterogeneity affect the information transmission capacity of populations of both FitzHugh-Nagumo (FHN) neurons and leaky integrate-and-fire (LIF) neurons, to test whether heterogeneity has similar effects to noise in increasing information transmission. Specifically, trials were performed using four distinct levels of random heterogeneity across the bias voltages $b_i$, the levels being none, low, medium, and high, corresponding to 0, 0.1, 0.5, and 2 times the input signal standard deviation $\sigma_s = 0.1$ (where the highest level spans the expected range of the input signal). At each level of heterogeneity, the standard deviation $\sigma_\eta$ of a stochastic offset---applied to the voltage parameter of the neuron model---was varied logarithmically from an insignificant level ($\sigma_\eta = 10^{-9}$ for FHN neurons and $\sigma_\eta = 10^{-6}$ for LIF neurons) to levels 10 times the magnitude of the input signal ($\sigma_\eta = 10^0$), in increments of $\log_{10}(\Delta\sigma_\eta) = 0.1$. For the remainder of this paper, we will use the term noise to refer specifically to this stochastic offset. At each level of heterogeneity and noise, 25 trials were performed to achieve an accurate estimate of the mean information transmission, with each trial using a unique set of random biases and a unique set of stochastic offsets added to the membrane voltage.

Each trial consisted of presenting a population of $N = 64$ FHN or LIF neurons with 4.5 seconds of band-limited aperiodic random signal with zero mean and standard deviation $\sigma_s = 0.1$. This signal has equal power below 5 Hz and zero power above 5 Hz, where the threshold of 5 Hz was chosen because it is significantly slower (approximately $\frac{1}{5}$) of the maximum firing rate of the FHN and LIF neurons (25--30 Hz). The population was divided into 32 ``on'' neurons and 32 ``off'' neurons, where each ``on'' (``off'') neuron received a positive (negative) version of the aperiodic input signal as input. The FHN or LIF neurons were simulated using their respective differential equations, and the output was decoded using a simple summing neuron that summed and low-pass filtered the outputs of the individual neurons to obtain the population output (outputs from ``off'' neurons were multiplied by -1 before summing). Finally, the mutual information between the aperiodic input signal and the decoded output signal was calculated. The mutual information for each level of heterogeneity and noise was computed as the average mutual information from all 25 trials at those levels. See \scn{methods} for the full details of the experimental setup.

\fig{info} shows the results of these experiments. In the homogeneous populations, a strong stochastic resonance effect is present, in which the transmitted information (\ie mutual information between the input and decoded output signals) was optimized for noise levels of $\sigma_\eta \approx 3 \times 10^{-3}$ (FHN neurons) and $\sigma_\eta \approx 1 \times 10^{-2}$ (LIF neurons). These noise levels are about 30 times (FHN neurons) or 10 times (LIF neurons) smaller than the input signal standard deviation of $\sigma_s = 0.1$. This makes them small enough not to significantly degrade the input signal, while being large enough to elicit stochastic resonance effects.

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/figure1_info.eps}
  \fi
  \caption{
    \textbf{Heterogeneity increases information transfer in populations of FHN and LIF neurons.} We simulated populations of 64 FHN neurons and 64 LIF neurons, varying the level of noise ($x$-axis) and heterogeneity (individual traces), and measured the mutual information between the aperiodic input signal $s(t)$ (where $\sigma_s = 0.1$) and the output signal $r(t)$ ($y$-axis). Moderate to high levels of heterogeneity resulted in an increase in information transmission at low to moderate noise levels in both FHN and LIF neurons. The increases in FHN neurons were modest, with a strong stochastic resonance effect still exhibited at all levels of heterogeneity. The increases in LIF neurons were much more pronounced, with moderate to high levels of heterogeneity surpassing the results achieved by stochastic resonance.
  }
  \figlabel{info}
\end{figure}

For both FHN and LIF neurons, heterogeneity also increases the information transmission of the population, specifically for noise levels below the optimal stochastic resonance noise level. In FHN neurons, the benefits of adding heterogeneity to neuron biases are modest (see \scn{phase}), resulting in a significant increase in information transmission, but not achieving the same levels of information transmission as stochastic resonance (\fig{info}A). For LIF neurons, the benefits of heterogeneity in neuron biases are pronounced, with moderate to high levels of heterogeneity achieving the same or even better levels of information transmission---across a wide range of noise values---as the optimal level of noise (\fig{info}B).

In the remaining experiments, we consider three mechanisms that allow both noise and heterogeneity to improve signal transmission in populations of neurons: 1) putting neurons out of phase with one another, 2) decreasing the reaction time of a population of neurons to a stimulus, and 3) linearizing the response of a population of neurons to a stimulus. Though noise and heterogeneity affect neurons differently, they both act to achieve these three goals. None of the three factors is sufficient in and of itself to explain the effects of stochastic resonance and heterogeneity on nonlinear dynamical systems, such as the populations of FHN and LIF neurons studied. Nevertheless, by examining each factor individually, we gain a deeper understanding of how both stochastic resonance and heterogeneity operate in nonlinear dynamical systems, and ultimately show that the two concepts are fundamentally connected (see \scn{discussion}).

\subsection{Phase}
\scnlabel{phase}

Consider the following: a group of perfectly homogeneous neurons with no noise, such that their dynamics are completely deterministic. Begun with the same initial conditions, they will follow exactly the same trajectory in phase space, since they all receive the same input signal and obey the same set of differential equations. Such a population of simulated neurons will exhibit perfectly synchronized firing events, and therefore the neurons contribute redundant information: any signal that could be decoded from the firing events of this population of neurons could also be decoded from the spikes of a single neuron. These neurons are perfectly in phase: at any time, they are all in the same phase of their firing cycle, and in the same position in phase space.

Adding independent noise to each neuron desynchronizes the firing events of the population over time by causing diffusion of the neuronal phases. A small random perturbation causes a slightly advanced or delayed firing of a neuron, creating a small shift in the phase of that neuron. Specifically, a firing neuron follows a stable limit cycle attractor in phase space; perturbations tangent to the limit cycle result in diffusion along the limit cycle, while perturbations perpendicular to the limit cycle rapidly return the the stable attractor \citep{Tomita1974}. Over time, this diffusion along the limit cycle results in the population of neurons having a wide distribution of phases (\fig{phase}), and the neurons fire asynchronously.

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/figure2_phase.eps}
  \fi
  \caption{
    \textbf{Noise and heterogeneity both desynchronize neurons.} To measure whether noise and heterogeneity have a desynchronizing effect on populations of neurons, we measured the standard deviation of neuron phases across the population (see \scn{methods}), where a smaller standard deviation indicates increased synchronization. Heterogeneity is seen to significantly desynchronize both FHN and LIF neurons, specifically across low to moderate levels of noise. At high levels of noise, the noise itself is sufficient to desynchronize neurons, and additional heterogeneity has no effect.
  }
  \figlabel{phase}
\end{figure}

This previous example is, of course, artificial: it requires that all neurons are \emph{exactly} the same, and start with \emph{exactly} the same initial conditions. In a more realistic scenario, one expects that small differences in neuronal parameters or small amounts of noise will cause neurons to desynchronize over a long period of time. This is not the case, because signals below the firing thresholds of a group of neurons have a synchronizing effect on the group. By stopping neuronal firing, these signals push all the neurons in the group towards their stable equilibrium points. When the signal returns to an amplitude above the firing thresholds of neurons in the group, all neurons will begin firing at approximately the same time, and will be once again in phase. Signals with higher frequency components are able to cross larger ranges of firing thresholds in quick succession, and have a stronger synchronizing effect than purely low frequency signals. Thus, more negative (positive) portions of the input signal repeatedly act to synchronize ``on'' (``off'') neurons, and groups of neurons must be able to desynchronize quickly after these synchronization events in order to remain out of phase with one another.

Similar to noise, heterogeneity acts to desynchronize neurons. Adding heterogeneity to a population significantly increases the range of distribution of neuron phases, thereby desynchronizing populations of both FHN and LIF neurons (\fig{phase}). Compared with a noiseless homogeneous population, where changes in an input signal will cross the common bias of all neurons at the same time and thereby put all neurons in phase with one another, changes in an input signal in a heterogeneous population will sequentially cross the biases of different neurons, causing some to begin firing before others. For example, an increasing input signal will cause neurons with lower biases to begin firing before those with higher biases, resulting in desynchronized firing within the population of neurons.

High frequency input signals may cross the biases of many neurons in a heterogeneous population at the same time, causing a large group of neurons to begin firing simultaneously. This effect does not have a significant synchronizing effect on LIF neurons, since the LIF tuning curves used have significantly different firing rates for different sizes of suprathreshold signal (\fig{tuning}). Specifically, a high-frequency increase in signal amplitude starts neurons firing at approximately the same time, but starts them firing at different rates, allowing them to quickly desynchronize. FHN neurons, however, all have around the same firing rate regardless of the strength of the suprathreshold signal (\fig{tuning}), and heterogeneity with higher-frequency signals has a reduced effect, since these signals cause many neurons to begin firing at once, unable to desynchronize due to their common firing rate. This discrepancy between LIF and FHN neurons results in heterogeneity having a large effect on phase in LIF neurons (\fig{phase}B), whereas the desynchronizing effect of heterogeneity in FHN neurons is much more modest (\fig{phase}A).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Response time}

Both noise and heterogeneity benefit the response time of a population of neurons, \ie how quickly the population responds to a rapidly changing signal. To measure this, we presented populations of FHN and LIF neurons with a step function and observed how well the population can follow the step function (\fig{step}). Though noise and heterogeneity both offer decreased response time, they achieve this using different mechanisms.

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/figure3_step.eps}
  \fi
  \caption{
    \textbf{Noise and heterogeneity both improve the step response of neuronal populations.} We excited populations of 100 FHN neurons and 100 LIF neurons using a step function, and measured the responses of populations without noise or heterogeneity ($b_i = 0, \sigma_\eta = 0$), with noise ($b_i = 0, \sigma_\eta = 10^{-3}$), and with heterogeneity ($b_i \in [-0.01,0.01], \sigma_\eta = 0$). Each line depicts the average population response over 100 trials, and the surrounding region shows one standard deviation. FHN populations without noise or heterogeneity show a considerable delay in responding to the step function, because all neurons are close to their stable equilibrium. Both noise and heterogeneity significantly reduce this delay, and also reduce oscillations by desynchronizing neurons (\fig{phase}). LIF populations only exhibit a mild initial delay, which noise and heterogeneity both correct; in LIF neurons, the main benefit to the step response is reducing oscillations through desynchronization.
  }
  \figlabel{step}
\end{figure}

Noise increases the response time of neurons by keeping them away from their stable equilibria during low points in the input signal. By presenting a population of 100 FHN neurons with a step input function, we found that a homogeneous population with no noise showed a considerable delay between the leading edge of the input step function and the response of the population (\fig{step}). Adding noise to the population decreases the response such that it is negligible, with a high proportion of the neurons responding immediately to the leading edge of the step. Noise produces this benefit by keeping neurons away from their stable equilibria during the below-threshold portions of the input signal (\ie before the leading edge of the step), allowing them to respond quickly to above-threshold portions of the input signal (\ie after the leading edge of the step). Neurons not subject to noise, on the other hand, are brought very close to their stable equilibria during the subthreshold portion of the input signal, and it takes a mildly suprathreshold signal---as provided by the step function---a long time to bring the neuron away from this equilibrium and into the tonic firing regime.

Heterogeneity also acts to increase the response time of a population of neurons. It does not directly keep neurons away from their equilibrium points, but by spreading neuron biases across a wide range, it ensures that neurons remain in different states of activation. Upon receiving a low input signal, most neurons will be pushed towards their stable equilibria, but neurons with higher firing thresholds will be pushed there more quickly than neurons with lower thresholds. Specifically, at the rising edge of the signal, neurons with low thresholds will be less settled to their equilibria than those with high thresholds, allowing them to respond more quickly. Furthermore, the high input signal will push neurons with low thresholds away from their equilibria more quickly than those with high thresholds, again resulting in quick activation of low threshold neurons. This effect can be observed in the response of the FHN population to a step function (\fig{step}), where the heterogeneous population responds more quickly than the noiseless homogeneous population, but still less quickly than the noisy homogeneous population.

\subsection{Linearization}

Intrinsic noise makes neuron tuning curves more linear, using a mechanism similar to subthreshold stochastic resonance. For input signals below the neuron's threshold, noise operates in a manner identical to subthreshold stochastic resonance, occasionally raising the neuron's membrane voltage high enough to cross the threshold and trigger an action potential. For input signals above the neuron's threshold, noise causes the membrane voltage to occasionally fall below the threshold, slowing the firing rate slightly compared with the noise-free case. For signals well below the neuron's threshold, the noise is not sufficient to cause any spikes, and for signals well above the threshold, the noise is insufficient to stop any spikes: these portions of the tuning curve remain unchanged between the noise-free and noisy cases. Overall, the addition of noise has the effect of linearizing the response of the neuron near its firing threshold, resulting in a smoother transition from the silent to tonic firing regimes (\fig{tuning}). This result has been previously observed and extensively studied \citep{Stocks1996,Chialvo1997}.

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/figure4_tuningnoisy.eps}
  \fi
  \caption{
    \textbf{Noise linearizes neuron tuning curves.} We probed FHN neurons and LIF neurons using constant inputs at various voltages, using varying levels of noise, and measured the output firing rate. Noise has the effect of making the tuning curves more linear, specifically eliminating the sharp threshold in both FHN and LIF neurons. The main mechanism is that of subthreshold stochastic resonance: noise allows subthreshold inputs to cause spikes by occasionally pushing the membrane voltage above the threshold. This occurs in both FHN and LIF neurons, allowing them to represent subthreshold signal components. In FHN neurons, small levels of noise can also cause firing rates for small suprathreshold signals to drop slightly by briefly pushing the neuron away from its firing limit cycle in state space, though interestingly this does not happen for larger noise levels. Large levels of noise in FHN neurons can cause large fluctuations which jump from one point on the limit cycle to another point much farther ahead, causing elevated firing rates. In LIF neurons, noise added to small suprathreshold signals causes the neuron to fire slightly faster, because the noise can cause the rising membrane voltage to prematurely cross the threshold, causing the spike to occasionally occur sooner than it would in the noiseless case.
  }
  \figlabel{tuning}
\end{figure}

Neuron heterogeneity has a linearizing effect similar to adding noise to a population, but unlike the noisy case, where it made sense to talk about noise linearizing an individual neuron's tuning curve, here we will discuss how heterogeneity linearizes the \emph{population's} tuning curve. The population's tuning curve is the set of points connecting the input signal strength to the decoded output. Since our model uses a simple summing decoder, the population tuning curve is equal to the sum of the individual neuron tuning curves. For a population of neurons with uniformly varied random biases, the population tuning curve is much more linear than a single neuron's tuning curve, whereas in the homogeneous case the population's tuning curve is the same as the neurons' tuning curve (\fig{tuninghetero}). This linearizing effect of heterogeneity allows for better representation of an input signal.

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/figure5_tuninghetero.eps}
  \fi
  \caption{
    \textbf{Heterogeneity linearizes the response of a neuronal population to a stimulus.} Panel A shows the tuning curves for a heterogeneous population of $N = 9$ neurons, where one neuron has been chosen to have a bias of $-0.15$ and the other neurons have random biases uniformly sampled from the interval $[-0.15,0.15]$. Panel B shows how the population response (heavy solid line) is much closer to a linear response (heavy dot-dashed line) than a single LIF tuning curve (heavy dashed line), even though the heterogeneous biases are far from evenly spaced. The light traces indicate how the population response is the average of the individual neuron tuning curves, where each colored curve is a scaled version of the tuning curves in panel A (with scaling factor $1/N$), and the population response is the sum of these curves. Using a larger population of neurons (\ie increasing $N$) results, on average, in a more linear population response.
  }
  \figlabel{tuninghetero}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
\scnlabel{discussion}

Our results show that noise and heterogeneity in populations of neurons can improve the ability of a neuronal population to encode and transmit a signal. Both noise and heterogeneity can be understood as using similar mechanisms to achieve this increase in information transmission, specifically by 1) desynchronizing the neurons within the population, 2) improving the reaction time of the population to a rapidly changing signal, and 3) linearizing the response of the population. Consequently, both noise and heterogeneity may be important mechanisms for optimizing information flow in neural systems.

To characterize the benefits of both noise and heterogeneity on signal transmission in populations of neurons, we can separate them into two categories: \emph{temporal} properties of neurons, and \emph{rate} properties of neurons. The temporal category is concerned with the desynchronizing and reaction time effects of noise and heterogeneity, and the rate category is concerned with the linearizing property of both noise and heterogeneity on a neuronal population.

The temporal benefits of noise and heterogeneity can only be observed in spiking neuron models, such as the FHN and LIF models examined in this paper. The output of these models depends not only on the the output firing rate for a given input signal, as determined by the neuron tuning curve, but also by the timing of individual spike events. When nonlinear dynamical models of neurons are examined in a stochastic resonance simulation, as in \cite{Stocks2001}, the results are a combination of the temporally-based desynchronizing and response-time effects of noise, as well as the rate-based linearizing effects of noise.

The rate view of neurons treats them as being no more than input-output (or tuning) curves. That is, all temporal properties are ignored, and a neuron is reduced to a nonlinear function which turns an input signal into an output signal. Suprathreshold stochastic resonance effects observed in binary threshold units deal only with the rate benefits of noise \citep{Stocks2000,Stocks2001a,McDonnell2006}. This is because binary threshold units have no temporal properties---they are either on or off depending on the current input signal and noise---and cannot be affected by the temporal benefits of noise. The linearization effects of noise on nonlinear systems have been previously studied (see \cite{Stocks1996} for a review). We have observed that heterogeneity can also achieve these linearization effects across a population of neurons, which has previously been reported by \cite{Eliasmith2003}.

One important difference between the LIF and FHN neuron models is that the LIF model is a type~I neuron model, and the FHN model is a type~II neuron model (though LIF neurons are not prototypical type~I neurons \citep{Mato2008}, they are qualitatively more similar to this category). The difference between type~I and type~II neuron models is that type~I models undergo a saddle-node bifurcation as the input current (the bifurcation parameter) increases, whereas type~II models undergo a subcritical Hopf bifurcation \citep{Mato2008}. Qualitatively, this means that type~I models have continuous tuning curves, \ie they can have arbitrarily small firing rates, whereas type~II models have a discontinuity at their firing threshold, and cannot fire slower than a given rate. The fact that the LIF model is a type~I neuron model, in addition to other factors, makes it a better model of cortical neurons \citep{Koch1999}. The LIF model is therefore more relevant when examining population encoding as proposed to occur in cortex, though other neuron models may be preferable for certain sensory coding situations. Previous numerical experiments that have examined stochastic resonance and heterogeneity together, such as \cite{Stocks2000,McDonnell2006}, have used binary threshold units in their simulations, which are qualitatively very similar to type~II neurons (but without the temporal component); the results of these experiments may have limited applicability to cortical neural systems.

The heterogeneity of biases added to neuronal populations had a much more beneficial effect on information transmission in LIF neurons than in FHN neurons (\fig{info}). We hypothesize that this is because of key differences in the shape of the neuron tuning curves, combined with the fact that heterogeneity was only added to one neuron parameter, namely the bias voltage $b_i$. The FHN model, being a type~II neuron model, has a tuning curve that is qualitatively similar to a binary threshold unit, with a discontinuous increase in firing rate at the firing threshold, and offering little change in firing rate for signals above the threshold (\fig{tuning}A). On the other hand, the LIF model, being a type~I neuron model, has a continuous and monotonically increasing tuning curve, capable of having arbitrarily small firing rates, and producing significantly higher firing rates for input signals that are well above the firing threshold than for those slightly above the firing threshold (\fig{tuning}B). If a high-frequency signal crosses the firing thresholds of many FHN neurons in quick succession, all these neurons will begin firing both at approximately the same time and with approximately the same rate, taking a long time to desynchronize. With LIF neurons, the high-frequency signal can still cross the firing thresholds of many neurons simultaneously, causing them to begin firing at approximately the same time. However, the signal will result in significantly higher firing rates for neurons with lower firing thresholds than for those with higher firing thresholds, allowing the group of neurons to desynchronize quickly. This key difference between the tuning curves of FHN and LIF neurons results in better desynchronization for LIF neurons when heterogeneity is only added to the bias voltage. We hypothesize that if heterogeneity were added to other neuron parameters, for example giving different FHN neurons different maximum firing rates, then FHN neurons would be able to desynchronize as quickly as LIF neurons and heterogeneity would offer a similar benefit in information transmission to stochastic resonance in FHN neurons. A system with this level of heterogeneity would require a more complex decoding system than the simple summing neuron model used in this paper, perhaps using the tuning curves to solve for the optimal linear decoders as described by \cite{Eliasmith2003}, and we leave this for future work.

Noise and heterogeneity have a fundamental similarity which goes deeper than their mutual ability to increase information transmission in neuronal populations: noise and heterogeneity are both a form of random variation. Specifically, noise corresponds to variation within the phase space (state space) of a system over time, and heterogeneity corresponds to variation within the parameter space of a system across a population. It is important to emphasize that the heterogeneity used in this paper was completely random; the heterogeneous biases did not have to be specifically chosen or fine-tuned in any way, indicating that random heterogeneity could be implemented in a messy biological system like the brain. Finally, in analogy to stochastic resonance with noise, there is a critical level of heterogeneity that optimizes information transmission (\fig{infohetero}). Thus noise and heterogeneity should be thought of as analogous mechanisms, but in different parts of the neuronal model (the state space versus the parameter space, respectively).

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics{../figures/bw/figure6_infohetero.eps}
  \fi
  \caption{
    \textbf{Heterogeneity exhibits resonance effects in information transfer.} In an experiment similar to that shown in \fig{info}, the mutual information was measured between the input signal $s(t)$ and output signal $r(t)$ of populations of 64 LIF neurons with varying levels of noise and heterogeneity. The key difference is that in this figure, the radius of heterogeneity $b_r$ (where $b_i \in [-b_r, b_r]$) is varied along the $x$-axis, and each individual trace depicts a different noise level $\sigma_\eta$. At lower levels of noise ($\sigma_\eta = 1.0 \times 10^{-6}, 3.0 \times 10^{-3}$), there is a critical level of heterogeneity that optimizes information transmission. Specifically, this level of heterogeneity is between one and two times the standard deviation of the input signal $\sigma_s = 0.1$. At the optimal level of noise ($\sigma_\eta = 1.0 \times 10^{-2}$, as shown in \fig{info}), heterogeneity has no additional benefit: the noise itself is enough to allow for good information transmission. Finally, at high levels of noise ($\sigma_\eta = 1.0 \times 10^{-1}$), the noise overpowers the input signal, and no amount of heterogeneity can improve information transmission. This same experiment was performed with FHN neurons, but the change in information transmission due to heterogeneity was minimal, likely because the FHN tuning curves have little variation above the input stimulus level (see \scn{discussion}).
  }
  \figlabel{infohetero}
\end{figure}

The extent to which actual biological neural systems exploit either heterogeneity or stochastic resonance for improved information encoding and transmission remains unclear, and may vary significantly from system to system. One recent study shows that heterogeneity can be very important in encoding sensory information for certain electric fishes \citep{Marsat2010}. Another recent study shows SR effects in human auditory cortex, specifically that noise can increase correlation both within and between some cortical regions \citep{Ward2010}. More studies of this nature are required to determine both what neural systems show evidence of heterogeneity and stochastic resonance, and more importantly, whether these systems actually take advantage of heterogeneity and stochastic resonance for improved population coding. Our contribution is demonstrating \emph{how} a neural system could exploit either heterogeneity or stochastic resonance, or both, for improved information processing characteristics.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Models and Methods}
\scnlabel{methods}

Our model consists of an input signal $s(t)$ encoded by a population of neurons, specifically having the same signal injected into the soma of all cells in the population (see \eqn{fhn} and \eqn{lif}). We ran simulations using both the FitzHugh-Nagumo (FHN) neuron model and the leaky integrate-and-fire (LIF) neuron model. Signal decoding is modeled by summing the spikes of each neuron in the population and low-pass filtering the result to obtain the output signal $r(t)$. This can be thought of as analogous to determining the somatic current of a single receiving neuron, where the filtering is performed by synapses. The input and output signals are then compared using mutual information, to measure how well the output signal encodes the input signal. Independent and identically distributed (i.i.d.) additive Gaussian noise processes are added to each neuron, with standard deviation $\sigma_\eta$ used to modulate the noise level in each experiment. Heterogeneity is also added by choosing neuron biases $b_i$ from a random uniform distribution, where the width of the distribution controls the degree of heterogeneity.

\textbf{FitzHugh-Nagumo (FHN) model.} The FHN model is a simplification of the Hodgkin-Huxley model \citep{FitzHugh1961}. We chose to use the standard formulation presented by \cite{Izhikevich2006}, given for the $i$-th neuron by two coupled ordinary differential equations:
\begin{equation}
  \begin{array}{rl}
    \dot{v_i} & = v_i - \frac{1}{3}v_i^3 - w_i + \beta_i + e_i s(t) + \eta_i(t) \vspace{1em}\\
    \dot{w_i} & = 0.08(v_i - 0.8w_i + 0.7),
  \end{array}
  \eqnlabel{fhn}
\end{equation}
where $v_i(t)$ is a fast (voltage) variable, $w_i(t)$ is a slow (recovery) variable, $\beta_i$ is a constant bias voltage, $e_i$ is a simple encoder equal to 1 for ``on'' neurons and $-1$ for ``off'' neurons, $s(t)$ is an input signal common to all neurons, and $\eta_i(t)$ is a Gaussian white noise process with autocorrelation $\left<\eta_i(t)\eta_j(t + \tau)\right> = \sigma^2 \delta_{ij}\delta(\tau)$. We chose $\beta_i = 0.3216 - b_i$, with the offset 0.3216 chosen such that the bias $b_i$ corresponds to the threshold at which the neuron begins firing.

\textbf{Leaky integrate-and-fire (LIF) model.} The LIF model describes a neuron as a membrane with resistance $R$ and capacitance $C$, which fires spikes when the membrane voltage $v(t)$ crosses the threshold $V_{th}$. It is governed by the differential equation
\begin{equation}
  \tau_{RC} \dot{v_i} = -v_i + \beta_i + \alpha \left( e_i s(t) + \eta_i(t) \right),
  \eqnlabel{lif}
\end{equation}
where $v_i(t)$ is the membrane voltage, $\tau_{RC} = RC$ is the membrane time constant, $\alpha$ is the input signal gain, $\beta_i$ is a constant bias voltage, $\eta_i(t)$ is a delta-correlated Gaussian white noise process (as with the FHN neuron model), and $s(t)$ is an input signal common to all neurons. When a neuron voltage reaches the threshold voltage $V_{th} = 1$, the neuron fires a spike, and begins a refractory period. The length of the refractory period is given by the refractory time constant $\tau_{ref}$, and during this time the neuron voltage is held at zero. Once the refractory period is finished, the neuron obeys \eqn{lif} until another spike occurs.

We chose $\tau_{RC}$ to be 20 ms, which is typical for cortical neurons \citep{McCormick1985}. We then chose $\tau_{ref}$ to be 33 ms, thus giving a maximum firing rate (30 Hz) approximately the same as the FHN model. We chose $\alpha$ to be 15, to achieve a similar maximum firing rate to the FHN model over the range of inputs used. We chose $\beta_i = V_{th} - \alpha b_i$, such that $b_i$ is a bias that corresponds to the threshold at which the neuron begins firing.

\textbf{Input signal generation.} For the input signal $s(t)$, we used an aperiodic random signal with zero mean and equal power below a nominal maximum frequency. The signal had a standard deviation of $\sigma_s = 0.1$. It had a maximum frequency of 5Hz for the simulations, roughly corresponding to one fifth of the maximum firing frequency of the neurons over the range of inputs.

\textbf{Output signal decoding.} To determine the output signal $r(t)$, we used the basic model of a summing neuron similar to the one proposed by \cite{Stocks2001}. Specifically, the output signal $r$ is given by
\begin{equation}
  \begin{array}{rcl}
    \tau \dot{r} & = & -r + \sum\limits_{i=1}^N q_i \vspace{0.5em}\\
    q_i & = & \begin{cases} v_i & \text{if } v_i > V_{th} \\ 0 & \text{otherwise} \end{cases},
  \end{array}
\end{equation}
where $N$ is the number of neurons in the encoding population, and $V_{th}$ is the firing threshold ($V_{th} = 0$ for the FHN model, and $V_{th} = 2$ for the LIF model). This model simulates a summing neuron which only receives the spikes (\ie parts of the neuron voltage $v_i$ greater than the threshold $V_{th}$, and then low-pass filters the sum of these spikes, a basic description of how a summing neuron membrane sums and filters input voltages.

\textbf{Mutual information.} We use Shannon mutual-information as the measure of similarity between the input and output signal \citep{Heneghan1996,Stocks2001}, where the mutual information $I$ (in bits) between the output signal $r$ and input signal $s$ is given by
\begin{equation}
  I(s,r) = \sum\limits_{s \in S} \sum\limits_{r \in R} p(s,r) \log_2\left(\frac{p(s,r)}{p(s)p(r)}\right),
\end{equation}
where $S$ is the domain of the input signal $s(t)$ and $R$ is the domain of the output signal $r(t)$. To compute the joint probability distribution $p(s,r)$ for discrete signals $s(t)$ and $r(t)$, the domains $S$ and $R$ are each divided into equal bins, and a joint histogram of the points $[s(t_i),r(t_i)]$ is formed. This histogram is used to approximate the joint probability function $p(s,r)$ by scaling it so the sum across all $s$ and $r$ is unity (\ie dividing the histogram by the total number of points $N$). For all simulations, we divided the domains $S$ and $R$ into 19 bins each.

One shortfall of this comparison method is that it does not account for the delay between the input and output signals, due to the non-zero reaction time of the neurons. In neural systems, some delay is to be expected, and we should not penalize our simulated networks for not being able to encode and decode an input signal instantaneously. The difficulty in comparing a delayed version of the input signal with the output signal is that it raises the question of what length of delay should be used. Each pair of noise and heterogeneity levels may conceivably have a unique optimal delay, and in order to not favor either noise or heterogeneity, the optimal delay (\ie the delay that produces the highest mutual information) should be calculated and used at each pair of noise and heterogeneity levels. In the interest of simplicity, we decided to instead compare the input and output signals with no delay, as is consistent with previous experiments \citep{Stocks2001}, and allows the benefits of noise and heterogeneity on response time to be measured.

\textbf{Heterogeneity.} To introduce heterogeneity into the population, we varied the bias voltages of the neurons. In homogeneous populations, all neurons had bias voltages equal to zero, the mean of the input signal. Heterogeneous populations had neurons with uniformly randomly distributed biases chosen from a range centered around the mean. In particular, three ranges were used: $[-0.01,0.01]$, $[-0.05,0.05]$, and $[-0.2,0.2]$, corresponding to small, medium and large levels of heterogeneity. The ranges were chosen to be roughly logarithmically distributed, with the largest range corresponding to biases that covered two standard deviations of the input signal.

\textbf{``On'' and ``off'' neurons.} To help minimize subthreshold stochastic resonance effects from significantly contributing to our results, we used populations that contained an equal number of ``on'' and ``off'' neurons, \ie neurons that both increase firing rates as the input signal increases, and increase firing rates as the signal decreases. In terms of our model, this means that half the neurons of each population had $e_i = 1$, and the other half had $e_i = -1$. This means that half the population represents the positive and negative parts of the signal, respectively, and only a small amount of information is gained by the subthreshold SR effect, which allows each half of the population to represent both positive and negative signal components.

\textbf{Population phase.} To determine how synchronous a population of neurons was, we measured the standard deviation of the neuron phases around the mean population phase. FHN neuron phases were determined by measuring the angle of the $(V(t),W(t))$ point in phase space around a central point, chosen to be $(-0.22,0.60)$ since this lies roughly in the center of the FHN limit cycle. For LIF neurons, the refractory period was scaled to correspond to the first $180^\circ$ of phase, and if the neuron was out of the refractory period, the phase was computed as $180^\circ ( 1 + \max(0,\min(1,V(t))) )$. The mean phase $\bar\theta$ was computed as
\begin{equation}
  \bar\theta(t) = \tan^{-1} \left( \frac{1}{N}\sum\limits_{i=1}^N \cos\theta_i(t), 
                                  \frac{1}{N}\sum\limits_{i=1}^N \sin\theta_i(t) \right),
\end{equation}
where $\theta_i(t)$ is the phase of neuron $i$ at time $t$, $N$ is the number of neurons in the population, and $\alpha = \tan^{-1}(x,y)$ is the two-argument arctangent function such that $\tan\alpha = y/x$. The standard deviation $\theta_\sigma(t)$ of the population phase is then given by
\begin{equation}
  \theta_\sigma(t) = \sqrt{ \frac{1}{N} \sum\limits_{i=1}^N (\mathrm{wrap}(\theta_i - \bar\theta))^2 },
\end{equation}
where the function $\mathrm{wrap}(\alpha) = (\alpha + \pi \mod 2\pi) - \pi$ converts the angle $\alpha$ to its equivalent angle in the range $[-\pi,\pi)$. \fig{phase} presents the mean over all simulation time of the standard deviation of the phase.

\textbf{Simulation details.} The information transmission (\fig{info}) and phase (\fig{phase}) plots were computed using noise levels in the range $\sigma_\eta \in [10^{-9},1]$ for FHN neurons and in the range $\sigma_\eta \in [10^{-6},1]$ for LIF neurons, with the levels logarithmically spaced in steps of $10^{0.1}$. We chose the low end of these ranges to be an insignificant level of noise (\ie with information transmission qualitatively the same as the no-noise case) and increasing until the noise completely overpowered the signal (\ie only a negligible amount of signal information could be recovered). At each noise level, and each of the four levels of heterogeneity, 25 simulations were performed, each having a simulation time of 4.5 seconds, with the first 0.5 seconds of each simulation being discarded to allow any transients created by the initial conditions to settle. The time step used varied depending on the noise level: larger levels of noise are more unstable and require a smaller time step, so a step of $\Delta t = 10^{-5}$ was used for $\sigma_\eta >= 0.1$, $\Delta t = 2.5 \times 10^{-5}$ for $\sigma_\eta >= 0.01$, and $\Delta t = 10^{-4}$ for $\sigma_\eta < 0.01$. For each simulation, new sets of both random biases $b_i$ and noise $\eta_i(t)$ were generated for each neuron. The final traces in the figures show the mean over all 25 simulations. The standard error of the mean at each point was not plotted because it was small in all cases, with statistically significant differences between the traces.

The step response plots (\fig{step}) represent the average population step response of a $N = 100$ neuron population over 100 trials. Each trial consists of presenting a population of $N = 100$ neurons with a step input $S_{step}(t)$ of 2 second duration, where the step input is equal to $-0.01$ for $t < 1.0$ and $0.01$ for $t >= 1.0$, with a time step $\Delta t = 10^{-4}$. The time shown in the plots corresponds to simulation time $t \in [0.9, 1.4]$, that is 0.1 seconds before the leading edge of the step to 0.4 seconds afterward. Output responses are scaled so that the average response for simulation time $t > 1.5$ corresponds to $0.01$, the final value of the step function.

The tuning curve plots (\fig{tuning}) were all calculated empirically by measuring the firing rate of simulated neurons. The range of inputs measured was $[-0.2, 0.2]$, divided into 51 equally spaced points (intervals of $0.008$). At each point, a constant signal at that input value was presented to $N = 30$ neurons, with a duration of 5.5 seconds. The first 0.5 seconds of neuron output was discarded, to reduce the effects of initial condition transients, and the average spike rate for each neuron over the remaining time period was calculated. FHN neurons were considered to spike when they maintained a voltage level $V(t) > 0$ for at least 0.01 seconds, and LIF neurons were considered to spike whenever the voltage crossed the threshold $V_{th}$, as given by the neuron model. The tuning curves presented in the plots show the average firing rate of all $N = 30$ neurons for each input value.

\subsection*{Acknowledgments}

We thank Jeff Orchard for feedback on an earlier draft of this manuscript.

\begin{thebibliography}{100}
\providecommand{\natexlab}[1]{#1}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
  \urlstyle{rm}\Url}\fi

\bibitem[Chialvo et~al., 1997]{Chialvo1997}
Chialvo, D., Longtin, A., and M\"{u}ller-Gerking, J. (1997).
\newblock {Stochastic resonance in models of neuronal ensembles}.
\newblock {\em Physical review E}, 55(2):1798--1808.

\bibitem[Collins et~al., 1995]{Collins1995}
Collins, J.~J., Chow, C.~C., and Imhoff, T.~T. (1995).
\newblock {Stochastic resonance without tuning}.
\newblock {\em Nature}, 376:236--238.

\bibitem[Eliasmith and Anderson, 2003]{Eliasmith2003}
Eliasmith, C. and Anderson, C.~H. (2003).
\newblock {\em {Neural Engineering: Computation, Representation, and Dynamics
  in Neurobiological Systems}}.
\newblock MIT Press, Cambridge, MA.

\bibitem[Fitzhugh, 1961]{FitzHugh1961}
Fitzhugh, R. (1961).
\newblock {Impulses and physiological states in theoretical models of nerve
  membrane}.
\newblock {\em Biophysical journal}, 1(1948):445--466.

\bibitem[Gammaitoni et~al., 1998]{Gammaitoni1998}
Gammaitoni, L., H\"{a}nggi, P., Jung, P., and Marchesoni, F. (1998).
\newblock {Stochastic resonance}.
\newblock {\em Reviews of Modern Physics}, 70(1):223--287.

\bibitem[Heneghan et~al., 1996]{Heneghan1996}
Heneghan, C., Chow, C., Collins, J.~J., Imhoff, T.~T., Lowen, S.~B., and Teich,
  M.~C. (1996).
\newblock {Information measures quantifying aperiodic stochastic resonance.}
\newblock {\em Physical Review E}, 54(3):R2228--R2231.

\bibitem[Izhikevich and FitzHugh, 2006]{Izhikevich2006}
Izhikevich, E.~M. and FitzHugh, R. (2006).
\newblock {FitzHugh-Nagumo model}.
\newblock {\em Scholarpedia}, 1(9):1349.

\bibitem[Koch, 1999]{Koch1999}
Koch, C. (1999).
\newblock {\em {Biophysics of computation: Information processing in single
  neurons}}.
\newblock Oxford University Press, New York, NY.

\bibitem[Longtin and Chialvo, 1998]{Longtin1998}
Longtin, A. and Chialvo, D.~R. (1998).
\newblock {Stochastic and Deterministic Resonances for Excitable Systems}.
\newblock {\em Physical Review Letters}, 81(18):4012--4015.

\bibitem[Machens et~al., 2010]{Machens2010}
Machens, C.~K., Romo, R., and Brody, C.~D. (2010).
\newblock {Functional, but not anatomical, separation of "what" and "when" in
  prefrontal cortex.}
\newblock {\em The Journal of neuroscience : the official journal of the
  Society for Neuroscience}, 30(1):350--60.

\bibitem[Marsat and Maler, 2010]{Marsat2010}
Marsat, G. and Maler, L. (2010).
\newblock {Neural heterogeneity and efficient population codes for
  communication signals.}
\newblock {\em Journal of neurophysiology}, 104(5):2543--55.

\bibitem[Mato and Samengo, 2008]{Mato2008}
Mato, G. and Samengo, I. (2008).
\newblock {Type I and type II neuron models are selectively driven by
  differential stimulus features}.
\newblock {\em Neural computation}, 2440:2418--2440.

\bibitem[McCormick et~al., 1985]{McCormick1985}
McCormick, D.~A., Connors, B.~W., Lighthall, J.~W., and Prince, D.~A. (1985).
\newblock {Comparative electrophysiology of pyramidal and sparsely spiny
  stellate neurons of the neocortex}.
\newblock {\em Journal of Neurophysiology}, 54:782--806.

\bibitem[McDonnell and Abbott, 2009]{McDonnell2009}
McDonnell, M.~D. and Abbott, D. (2009).
\newblock {What is stochastic resonance? Definitions, misconceptions, debates,
  and its relevance to biology.}
\newblock {\em PLoS Computational Biology}, 5(5):e1000348.

\bibitem[McDonnell et~al., 2006]{McDonnell2006}
McDonnell, M.~D., Stocks, N.~G., Pearce, C.~E., and Abbott, D. (2006).
\newblock {Optimal information transmission in nonlinear arrays through
  suprathreshold stochastic resonance}.
\newblock {\em Physics Letters A}, 352(3):183--189.

\bibitem[McDonnell and Ward, 2011]{McDonnell2011}
McDonnell, M.~D. and Ward, L.~M. (2011).
\newblock {The benefits of noise in neural systems: bridging theory and
  experiment}.
\newblock {\em Nature Reviews Neuroscience}, 12(7):415--426.

\bibitem[Stocks, 2000]{Stocks2000}
Stocks, N.~G. (2000).
\newblock {Suprathreshold stochastic resonance in multilevel threshold
  systems}.
\newblock {\em Physical review letters}, 84(11):2310--3.

\bibitem[Stocks, 2001]{Stocks2001a}
Stocks, N.~G. (2001).
\newblock {Information transmission in parallel threshold arrays:
  Suprathreshold stochastic resonance}.
\newblock {\em Physical Review E}, 63(4):1--9.

\bibitem[Stocks and Mannella, 2001]{Stocks2001}
Stocks, N.~G. and Mannella, R. (2001).
\newblock {Generic noise-enhanced coding in neuronal arrays}.
\newblock {\em Physical Review E}, 64:030902.

\bibitem[Stocks et~al., 1996]{Stocks1996}
Stocks, N.~G., Stein, N., Short, H., Mannella, R., Luchinsky, D., McClintock,
  P., and Dykman, M. (1996).
\newblock {Noise-induced linearization and delinearization}.
\newblock In {\em Fluctuations and Order}, pages 53--67.

\bibitem[Tomita et~al., 1974]{Tomita1974}
Tomita, K., Ohta, T., and Tomita, H. (1974).
\newblock {Irreversible circulation and orbital revolution}.
\newblock {\em Prog. Theor. Phys}, 52(6):1744--1765.

\bibitem[Ward et~al., 2010]{Ward2010}
Ward, L.~M., MacLean, S.~E., and Kirschner, A. (2010).
\newblock {Stochastic Resonance Modulates Neural Synchronization within and
  between Cortical Sources}.
\newblock {\em PLoS ONE}, 5(12):e14371.

\bibitem[Wiesenfeld et~al., 1994]{Wiesenfeld1994}
Wiesenfeld, K., Pierson, D., and Pantazelou, E. (1994).
\newblock {Stochastic resonance on a circle}.
\newblock {\em Physical Review Letters}, 72(14):2125--2129.

%% \bibitem[{LastName(2009)}]{Ref2009}
%% LastName, A. (2009).
%% \newblock Title for the first reference.
%% \newblock \emph{Journal of the first reference}, \emph{3}, 18 -- 88.

%% \bibitem[{Authors et~al.(2008)Author1, Author2, \&
%%   Author3}]{Ref2008}
%% Author1, A., Author2, A., \& Author3, A. (2008).
%% \newblock Title for the second reference.
%% \newblock \emph{Journal for the second reference}, \emph{5}, 188 -- 200.

\end{thebibliography}

\end{document}

### 2014-01-20 paper/letter.tex
Second version of the paper, with figures.

@@
 \usepackage{graphicx}
+\usepackage{enumitem}
 
-\usepackage[margin=1.375in]{geometry}
+\usepackage[margin=1.25in]{geometry}
 
-%% \topmargin 0.0cm
-%% \oddsidemargin 0.0cm
-%% \evensidemargin 0.0cm
-%% \textwidth 5.75in
-%% \textheight 9.25in
+\hyphenation{man-u-script}
 
-%% \usepackage{hyperref}
-
-\signature{Eric G. Hunsberger}
+\signature{Eric Hunsberger\\Matthew Scott\\Chris Eliasmith}
 \address{Centre for Theoretical Neuroscience\\200 University Ave. W.\\Waterloo, ON, Canada N2L 3G1}
@@
 \begin{document}
-\begin{letter}{Editoral Board\\Neural Computation}
-\opening{Dear Sir or Madam:}
+\begin{letter}{Editorial Board\\Neural Computation}
+\opening{Dear Editor:}
 
-I am writing on behalf of myself and my co-authors, Matthew Scott and Chris Eliasmith, to present you with a manuscript that we have prepared for publication in Neural Computation, entitled ``Heterogeneity increases information transmission of neuronal populations''.
-This article examines the effects of heterogeneity on information transmission in populations of simulated neurons, and finds that it has a role to play in increasing the amount of information a population is able to encode and transmit about an input signal.
-We compare these results with previous results in the stochastic resonance field detailing how noise can benefit information transmission in neural populations, and find that heterogeneity has similar effects to noise in this respect.
+Please find enclosed a revised version of our paper, now entitled ``The competing benefits of noise and heterogeneity in neural coding''.
+The comments and critiques from both reviewers have been helpful in improving the content and presentation of the manuscript.
+This letter details our changes to the manuscript since its previous submission, and describes how these changes address the criticisms of the reviewers.
+We respond to each reviewer comment individually, with the quoted comment presented in an indented paragraph, followed by our response.
 
-This research is important to the field because it advances the view that heterogeneity has a role to play in neural systems, and explains in detail why this is so.
-Research in stochastic resonance has done a great service by documenting the beneficial effects of noise in simulated, in vitro, and in vivo neural systems, but this same service has not yet been extended to heterogeneity.
-The goal of this manuscript is to provide theoretical foundations for understanding the mechanisms by which heterogeneity operates, and why it is important to neural systems.
-These foundations tie together the limited amount of existing research on the subject, and provide direction and motivation for future theoretical and experimental research into heterogeneity in neural systems.
+\ \newline
+{\large \bf Reviewer~1}
 
-Given the general interest of this paper and its potential to inspire future research in the area, we ask that you consider the enclosed manuscript for publication in Neural Computation.
-None of the material in the article has been published or is under consideration for publication elsewhere.
-We thank you for your time, and look forward to your reply.
+\begin{quotation}
+  The paper "Heterogeneity increases information transmission of neuronal populations" examines the role of heterogeneity in neuronal coding among groups of neurons of the same type using simulations of FitzHugh-Nagumo and Leaky Integrate and Fire (LIF) neurons.
+Many of the central findings of this paper, including the results that heterogeneity improves information transmission have been published previously in one version or another, for example:
+
+  \begin{enumerate}[label=\arabic*)]
+  \item The neural basis for combinatorial coding in a cortical population response.
+    Osborne LC, Palmer SE, Lisberger SG, Bialek W.
+  \item Intrinsic biophysical diversity decorrelates neuronal firing while increasing information content. Padmanabhan K, Urban NN.
+  \item The effect of noise correlations in populations of diversely tuned neurons. Ecker AS, Berens P, Tolias AS, Bethge M.
+  \item Implications of neuronal diversity on population coding. Shamir M, Sompolinsky H.
+  \item Intrinsic heterogeneity in oscillatory dynamics limits correlation-induced neural synchronization. Burton SD, Ermentrout GB, Urban NN.
+  \item Efficient coding in heterogeneous neuronal populations. Chelaru MI, Dragoi V.
+  \end{enumerate}
+
+  This is just a short list of recent papers.
+The work by Padmanabhan and Urban (2010) for instance, has a nearly identical title to this manuscript and reports the central finding of this paper, demonstrating this not only theoretically by experimentally.
+Additionally, work on the connection between heterogeneity and stochastic resonance and synchrony has been previously published by a number of authors.
+
+  For instance, the results in section 2.1 Phase, are described in part by Burton et al (2012).
+In addition to the ones described above, nearly a dozen previous papers exploring the question of heterogeneity and information coding have either been overlooked or not correctly cited by the authors.
+Consequently, a more complete literature review, both of neuroscience results in general, and computational neuroscience results specifically is warranted to determine what parts in this manuscript have not been previously published.
+\end{quotation}
+
+As Reviewer~1 observed, many areas of our paper reproduce results previously observed to some extent in the literature, and we have included additional references to document this.
+Specifically, we have added all the references suggested by Reviewer~1.
+We also performed further literature review, and added additional references to related heterogeneity studies, including a recent study by Mejias and Longtin (2012).
+These references, in addition to the references suggested by Reviewer~2, constitute a more complete review of the field.
+
+\begin{quotation}
+  The results in section 2.2 Response time require considerable development to be interesting, and may be an artifact of the parameters chosen.
+The authors are interested in describing the latency to firing of a population of neurons to a step input.
+They draw attention to the importance of heterogeneity to spread the latency times across the population, thus improving the mean response time.
+However, if the speed of response is the key metric by which information is conveyed, then having a population where the threshold is very low, or the bias (b(i)) is high (b(i) =0.1) for the population would be most optimal.
+In their simulations, would not a population where b(i) = 0.01 be the ones that respond fastest (particularly in the  FHN) neurons)and have a shorted latency as compared to a population where b(i) = [-0.01 to 0.01]?
+If that is correct, then their argument that diversity improves response time as compared to a homogeneous population with a low threshold is a ``straw man'' argument, achieved as a result of idiosyncratic parameter choices that produce their desired outcome.
+Would it not be correct that a homogeneous population would have a shorter latency if the threshold of that population was sufficiently low as compared to a diverse population with the different thresholds?
+\end{quotation}
+
+We have removed all results and discussion related to the benefits of noise and heterogeneity on response time.
+As Reviewer~1 observed, our presented results could have simply been the result of ``idiosyncratic parameter choices.'' While we do not believe this is the case, we do agree that the claim would have required significantly more development to sufficiently justify this conclusion.
+These results were not essential to our main thesis, and we therefore elected to remove them entirely.
+
+\begin{quotation}
+  Although the main result of the manuscript is one of interest to the theoretical and experimental neuroscience community, the subject has been explored in detail previously (a simple Pubmed search of ``heterogeneity and neuronal coding'' pulls up many of the papers that the authors appear not to be aware of) and is therefore not novel.
+\end{quotation}
+
+The main concern of Reviewer~1 is that the work presented in our paper is not novel.
+In our most extensive change to the paper, we added additional figures and refocused the presentation of the material to highlight the complex interaction between noise and heterogeneity in neuronal populations.
+This is the main contribution of our paper to the field, and we have revised the paper to emphasize this novel result.
+
+\begin{quotation}
+  Minor Comments:
+  \begin{itemize}
+  \item The authors should distinguish how heterogeneity in the bias is conceptually different from heterogeneity in threshold (which has been described previously), which would produce many of the same results described here.
+  \end{itemize}
+\end{quotation}
+
+Throughout the manuscript, we have replaced references to bias currents with references to firing threshold.
+In our experimental setup, these values are one and the same, but we wanted to emphasize that our results corroborate previous results related to the benefits of heterogeneity in firing threshold.
+
+\begin{quotation}
+  \ \vspace{-1.3em}
+  \begin{itemize}
+  \item There are a number of formatting and typographical errors through the manuscript.
+These seem to appear in the original uploaded file that made the manuscript difficult to read.
+For instance in the first line of the abstract, ``Noise- speci cally random uctuations added to the membrane voltages of a neuronal population.'' A few lines down ``Furthermore, we nd that both noise and heterogeneity.''
+  \end{itemize}
+\end{quotation}
+
+It is unfortunate that one reviewer had difficulty reading the manuscript.
+None of the errors described appeared in our copy of the manuscript.
+They may have been caused by not all fonts being embedded in the PDF version of the document; we will ensure that the revised version has all fonts embedded.
+
+Our revised manuscript highlights the novel identification and characterization of a nonlinear interaction between noise and heterogeneity; this subject has not been explored in detail in the current literature.
+The added literature references make clear where our research overlaps with previous results.
+
+\ \newline
+{\large \bf Reviewer~2}
+\begin{quotation}
+  %% \itshape
+  Overall comments: This manuscript addresses an interesting and timely topic and nicely illustrates a number of specific computational advantages of populations containing neurons with heterogeneous response properties over populations whose neurons are more homogenous.
+Given the growing interest in the computational role of biological heterogeneity in both experimental and theoretical contexts, the work is very timely.
+The manuscript would be improved by making better links to biological data and systems and 2) having detailed figures showing raw data like spike rasters rather than depending almost exclusively on summary figures.
+\end{quotation}
+
+Reviewer~2 had concerns about the connection to biological data and systems, and the omission of figures showing raw results from individual trials.
+We have added a significant number of references to both computational and experimental neuroscientific results, which provides a stronger connection between our research and the biological data.
+We have also added spike rasters to better illustrate our results on desynchronization; these spike rasters also help connect our results to biological data.
+
+\begin{quotation}
+  Specific comments:
+  The authors note that ``research into the role of heterogeneity within a group of neurons of the same type is limited (Eliasmith and Anderson (2003) is a notable exception.''  In fact the fail to mention several recent papers providing experimental evidence for heterogeneity.
+Beyond the Marsat and Maler 2010 paper cited by the authors considerable work on the computational (Brody and Hopfield 2003 Neuron;  Tkacik et al 2010 PNAS; Urban and Padmanabhan 2010 nature Neuroscience) and experimental (Angelo and Margrie 2012; Nature; Urban and Padmanabhan 2010 Nature Neuroscience  Burton et al 2012 J Neurophysiol) aspects of this issue have been published.
+Some of these even go beyond the current manuscript in some respects.
+Effects of heterogeneity on synchrony, for example are directly addressed by (Burton et al 2012).
+\end{quotation}
+
+As previously mentioned in our response to Reviewer~1, we have added many additional references to document the connections between our work and previous work on heterogeneity.
+Of the above references suggested by Reviewer~2, all but one were added to the revised manuscript.
+A reference to Tkacik et al.
+(2010, PNAS) was omitted because it analyzes heterogeneity in neural connections using Ising models, and is only marginally related to our examination of heterogeneity.
+
+\begin{quotation}
+  The idea of an optimal amount of heterogeneity (at least in the low sigma case) is quite interesting (Figure 6) and this figure and discussion should be moved to the results from the discussion.
+Do the authors have an explanation for why this optimum exists and why only in the low noise case?
+\end{quotation}
+
+As suggested, we moved the figure showing resonance in heterogeneity to the front of the results section.
+This result feeds nicely into our presentation of the conflicting benefits of noise and heterogeneity, and both results are now discussed early in the paper, including an explanation of why resonance occurs in heterogeneity, and only in the absence of noise.
+
+\begin{quotation}
+  The authors should better motivate their decision to provide all the neurons homogeneous inputs for decoding.  This will strike experimentalists as highly artificial and should be discussed.
+\end{quotation}
+
+We added a paragraph to the discussion section to motivate the use of a common input signal for all neurons.
+We both describe how noise and heterogeneity act to provide each neuron with a unique input current, and note several previous experiments in computational and experimental neuroscience that use a similar experimental setup.
+
+\begin{quotation}
+  The authors emphasize the ways in which noise and heterogeneity work to produce similar effects.
+However, one important way in which the effects differ should be considered.
+Noise can change the average activity (or response time or phase) across a population, but on a given trial, which neuron fires first or at a particular phase is random.
+Heterogeneity across cells that generates the same distribution of activity values as does noise does so in a way such that the activity of a given neuron is reliable from trial to trial.
+Thus downstream decoding may be based on the pattern of activity (or phase or response time) across neurons rather than only on the average value.
+\end{quotation}
+
+We added discussion of the idea that heterogeneity produces variability that remains constant from trial to trial, as opposed to variability from noise which differs between trials.
+We mention that more complex decoding methods could take advantage of this reproducibility, and that this is an area for future work.
+
+\begin{quotation}
+  The description desynchronizing as ``putting neurons out of phase'' with each other could be clarified.
+Phase implies a degree of periodicity that, while it may be present in these simulations, using these models, seems unnecessary for the point that the authors want to make.
+I would recommend describing this as ``desynchronizing'' or even calculating correlations and discussing the ``decorrelating'' effects of heterogeneity.
+Given the known effects of correlation on coding and information  (e.g.
+Zohary and Newsome) this may be the best approach.
+This point also would be helped by showing some more raw results, such as simulated spike rasters.
+\end{quotation}
+
+We modified our section on the desynchronizing effects of noise and heterogeneity to emphasize desynchronization instead of phase.
+We did leave some discussion of phase, because we think that viewing neurons as dynamical systems operating on a limit cycle in a phase space---with noise causing diffusion along this limit cycle---is beneficial to understanding how noise desynchronizes neurons.
+We also added spike rasters to these results to better illustrate the desynchronization.
+
+\begin{quotation}
+  The discussion of improved response time should consider or at least note that this effect is dependent on the sign of the noise.
+That is, noise with an inhibitory bias will not decrease response time or do so only minimally.
+\end{quotation}
+
+We have removed the discussion of improved response time, as suggested by Reviewer~1.
+
+\begin{quotation}
+  The discussion of noise linearizing F-I curves should address similar previous work of several groups, including Reyes Abbott and Chance Nature Neuroscience; Kuman...
+Doiron PloS Comp Bio; and Prescott PNAS.
+\end{quotation}
+
+We added two of the three suggested references to previous work on the linearization of F-I curves.
+A reference to a paper by Litwin-Kumar and Brent Doiron (with unspecified coauthors) appearing in PLoS Computational Biology was omitted because two such papers exist (Litwin-Kumar A, Oswald A-MM, Urban NN, Doiron B (2011) and Litwin-Kumar A, Chacron MJ, Doiron B (2012)), and neither one appeared to explicitly discuss the effects of noise on stimulus-response curves.
+
+\begin{quotation}
+  The authors frequently use the phrase ``information transmission'' even when it is not clear that information is being transmitted anywhere.
+I would suggest that the clarity of the MS could be improved by replacing this with other phrases such as information ``content'' or ``representation''.
+\end{quotation}
+
+References to information transmission have been replaced with references to information content or representation, to help reduce ambiguity.
+
+\begin{quotation}
+  p.
+18 Discussion.
+The authors point out that FHN show type II dynamics whereas LIF are type I and argue that therefore LIF are more comparable to cortical neurons.
+Previous work (e.g.
+Tsubo et al 2007 and others) has concluded that some cortical neurons show type II phase resetting curves.
+The apparent lack of a discontinuous increase in firing rate in these neurons may be due to the linearizing effects of noise around threshold.
+\end{quotation}
+
+We now mention the results of Tsubo et al. (2007), as well as similar results from Tateno et al. (2004), which may make our results in the type II FHN model applicable to cortex.
+
+\begin{quotation}
+  p. 11 Authors say that ``Noise increases response time...'' But I think that they mean decrease.
+\end{quotation}
+
+Response time is no longer discussed, so this sentence was removed.
+
+We hope that the changes detailed here address the concerns of the reviewers.
+We would like to thank them for their time, and for their insightful comments which have helped us improve the quality of the manuscript.
 

### 2014-01-20 paper/manuscript.tex
Second version of the paper, with figures.

@@
 \usepackage{multirow}
-%
+
 \usepackage[authoryear]{natbib}
-%
+
 \usepackage{rotating}
@@
 \usepackage{latexsym}
-%\DeclareGraphicsExtensions{.eps,.png}
 
@@
 \topmargin  -0.55in
-%
+
 \renewcommand{\baselinestretch}{2}
-%
+
 \interfootnotelinepenalty=10000
-%
-\renewcommand{\thesubsubsection}{\arabic{section}.\arabic{subsubsection}}
-\newcommand{\myparagraph}[1]{\ \\{\em #1}.\ \ }
-\newcommand{\citealtt}[1]{\citeauthor{#1},\citeyear{#1}}
-\newcommand{\myycite}[1]{\citep{#1}}
 
@@
 
-\newcommand{\ie}{\emph{i.e., }}
-\newcommand{\eg}{\emph{e.g., }}
-%% \newcommand{\cf}{\emph{cf. }}
-%% \newcommand{\etc}{\emph{etc.}}
-%% \newcommand{\wrt}{w.r.t. }
+\usepackage{nameref}
 
-\usepackage{nameref}
-\newcommand{\fig}[1]{Figure \ref{fig:#1}}
-\newcommand{\figlabel}[1]{\label{fig:#1}}
-\newcommand{\eqn}[1]{Equation \ref{eqn:#1}}
-\newcommand{\eqnlabel}[1]{\label{eqn:#1}}
-\newcommand{\scn}[1]{\textsc{\nameref{scn:#1}}}
-\newcommand{\scnlabel}[1]{\label{scn:#1}}
-
-\renewcommand{\thefootnote}{\normalsize \arabic{footnote}}
+%% \newcommand{\hidefigures}{\relax}
 
 \begin{document}
-\hspace{13.9cm}1
+\hspace{13.9cm}
 
@@
 
-{\LARGE Heterogeneity increases information transmission of neuronal populations}
+{\LARGE The competing benefits of noise and heterogeneity in neural coding}
 
 \ \\
-{\bf \large Eric G. Hunsberger$^{\displaystyle 1}$, Matthew Scott$^{\displaystyle 2}$, Chris Eliasmith$^{\displaystyle 1}$}\\
+{\bf \large Eric Hunsberger$^{\displaystyle 1}$, Matthew Scott$^{\displaystyle 2}$, Chris Eliasmith$^{\displaystyle 1}$}\\
 {$^{\displaystyle 1}$Centre for Theoretical Neuroscience, University of Waterloo, Waterloo, Ontario, Canada.}\\
@@
 \begin{center} {\bf Abstract} \end{center}
-Noise---specifically random fluctuations added to the membrane voltages of a neuronal population---has been shown to increase the amount of information a neuronal population can encode about an input signal; this phenomenon is one type of stochastic resonance.
-We demonstrate that heterogeneity within a population of simulated neurons---specifically the random variation of membrane bias currents across a population---produces similar benefits as additive noise in terms of increasing information encoding within the population.
-Furthermore, we find that both noise and heterogeneity use a common set of three mechanisms to increase information transmission: 1) they temporally desynchronize the firing of neurons within the population, 2) they decrease the response time of the population of neurons to sudden changes in the input signal, and 3) they linearize the response of a population to a stimulus.
-We examine each mechanism in detail, describing how these mechanisms operate in neuronal populations, and demonstrating how they operate in populations of simulated FitzHugh-Nagumo and leaky integrate-and-fire neurons.
-These three mechanisms work to distribute information equally across all neurons in the population, both in terms of signal timing (desynchronization and response time) and in terms of signal amplitude (linearization).
-Without noise or heterogeneity, all neurons in the population encode the same aspects of the input signal; adding noise or heterogeneity allows all neurons to encode \emph{complementary} aspects of the input signal, thereby preserving more information.
-Overall, the simulations detailed in this paper highlight the importance of heterogeneity in neuronal population coding, and demonstrate that it can be equally effective as stochastic resonance in increasing information transmission.
+
+Noise and heterogeneity are both known to benefit neural coding.
+Stochastic resonance describes how noise, in the form of random fluctuations in a neuron's membrane voltage, can improve neural representations of an input signal.
+Neuronal heterogeneity refers to variation in any one of a number of neuron parameters, and is also known to increase the information content of a population.
+We explore the interaction between noise and heterogeneity, and find that their benefits to neural coding are not independent.
+Specifically, a neuronal population better represents an input signal when either noise or heterogeneity is added, but adding both does not always improve representation further.
+To explain this phenomenon, we propose noise and heterogeneity both operate using two shared mechanisms: 1) temporally desynchronizing the firing of neurons in the population, and 2) linearizing the response of a population to a stimulus.
+We first characterize the effects of noise and heterogeneity on the information content of populations of either leaky integrate-and-fire or FitzHugh-Nagumo neurons.
+We then examine how the mechanisms of desynchronization and linearization produce these effects, and find that they work to distribute information equally across all neurons in the population, both in terms of signal timing (desynchronization) and in terms of signal amplitude (linearization).
+Without noise or heterogeneity, all neurons in the population encode the same aspects of the input signal; adding noise or heterogeneity allows neuron to encode \emph{complementary} aspects of the input signal, thereby increasing information content.
+The simulations detailed in this paper highlight the importance of heterogeneity and noise in population coding, demonstrate their complex interactions in terms of the information content of neurons, and explain these effects in terms of underlying mechanisms.
+
 %%%%%%%%%%%
@@
 \section{Introduction}
-\scnlabel{introduction}
+\label{scn:introduction}
 
@@
 The early stochastic resonance literature dealt mainly with \emph{subthreshold} stochastic resonance \citep{Gammaitoni1998}.
-In a deterministic setting, systems which only respond to inputs above a certain threshold---such as neurons---cannot transmit any information about subthreshold input signals, \ie input signals which are entirely below the system's threshold.
-In contrast, a noisy input signal (\ie a signal with an added stochastic offset) will occasionally cross the system's threshold, and thereby allows the system to transmit more information about the input signal.
+In a deterministic setting, systems which only respond to inputs above a certain threshold---such as neurons---cannot transmit any information about subthreshold input signals, \emph{i.e.,} input signals which are entirely below the system's threshold.
+In contrast, a noisy input signal (\emph{i.e,} a signal with an added stochastic offset) will occasionally cross the system's threshold, and thereby allows the system to transmit more information about the input signal.
 Too much noise, though, causes the input signal to become lost in random fluctuations.
-A moderate amount of noise---not too much, and not too little---is needed to optimize information transmission by the system about the input signal \citep{Wiesenfeld1994,Longtin1998}.
+A moderate amount of noise---not too much, and not too little---is needed to optimize the information about the input signal contained in the system \citep{Wiesenfeld1994,Longtin1998}.
 
-A new form of stochastic resonance---called \emph{suprathreshold} stochastic resonance (SSR)---has emerged in the past decade \citep{McDonnell2009}.
+A new form of stochastic resonance---called \emph{suprathreshold} stochastic resonance (SSR)---has emerged in the past decade \citep{McDonnell2009,McDonnell2011}.
 In neuroscience, suprathreshold SR looks at the benefits of noise to populations of nonlinear neurons receiving a common signal of which a significant portion is above the systems' firing threshold(s).
-This concept was first introduced by \cite{Stocks2000}, who looked at groups of binary threshold units, and found that noise can increase information transmission in such systems.
+This concept was first introduced by \cite{Stocks2000}, who looked at groups of binary threshold units, and found that noise increases information transmission in such systems.
 Subsequent research has examined the SSR effect in populations of more sophisticated simulated neurons \citep{Stocks2001}.
 
-Heterogeneity among neurons---the fact that not all neurons are identical, and vary across a wide range of parameters---is a topic that has largely been ignored in the stochastic resonance literature, and neuroscience literature in general \citep{Machens2010}.
-It is well known that there are large differences between neurons in different parts of the nervous system, and that there are many different types of neurons, but research into the role of heterogeneity within a group of neurons of the same type is limited (\cite{Eliasmith2003} is a notable exception).
+The fact that neurons are heterogeneous---with variation across a wide range of parameters---has largely been ignored in the stochastic resonance literature.
+It is well known that neurons vary widely across the nervous system.
+Even neurons in the same brain area and of the same qualitative type vary across numerous parameters.
+Though this diversity in neural systems was documented by some of the first researchers in the field, it is only recently that neuroscientists have begun to illuminate the functional role of heterogeneity.
+Early work by \cite{Eliasmith2003} demonstrated that heterogeneity increases both the ability of a population to represent an input signal, and the range of functions computable by the population.
+More recent work showed that heterogeneity allows neurons to use combinatorial coding schemes \citep{Osborne2008} and in general increase the information encoded by a population \citep{Shamir2006,Chelaru2008,Padmanabhan2010,Ecker2011}.
 In the field of stochastic resonance, the effects of noise in a heterogeneous population have been examined, but generally in specific cases such as for input signals much larger than the range of heterogeneity \citep{Stocks2000}, or for very large levels of noise \citep{McDonnell2006}.
@@
 
-The purpose of this paper is to compare the benefits of noise and the benefits of heterogeneity on information transmission by a population of simulated neurons.
-While previous studies have looked at the effects of noise, often claiming that noise is the central mechanism of SR and SSR effects \citep{Collins1995,Stocks2001} (see \cite{McDonnell2011} for a review), we are not aware of any work that has examined the effects of heterogeneity in similar circumstances.
-As we show, many effects of noise can also be achieved with heterogeneity.
-We argue that this is at least partly because of several shared effects on neural coding.
-We use both the FitzHugh-Nagumo (FHN) neuron model, a common model in SR research \citep{Wiesenfeld1994,Collins1995,Longtin1998,Stocks2001}, as well as the leaky integrate-and-fire (LIF) neuron model, a typical model of cortical neurons \citep{Koch1999}.
-We use a common aperiodic random signal $s(t)$ to drive populations of FHN and LIF neurons with varying levels of internal noise and heterogeneity, and compare the similarity of the decoded output signal with the input signal using mutual information.
-Stochastic offset is modeled as zero-mean Gaussian white noise with standard deviation $\sigma_\eta$, independently and identically distributed across all neurons (see \scn{methods}).
+The purpose of this paper is to compare how noise and heterogeneity help a neuronal population to represent an input signal.
+Previous studies have looked at the individual benefits of noise and heterogeneity in population coding, but the interplay of these two factors has not been examined in detail.
+We find that noise and heterogeneity exhibit complex interactions, and though they both help to increase the information contained in a population, adding both does not always increase information content further, as compared with adding only one or the other.
+We argue that the similar benefits of noise and heterogeneity, as well as their complex interactions, occur because they operate on the dynamics of a neuronal population through shared mechanisms.
+
+We explore this hypothesis in the context of two neuron models: the FitzHugh-Nagumo (FHN) neuron model, a common model in SR research \citep{Wiesenfeld1994,Collins1995,Longtin1998,Stocks2001}, and the leaky integrate-and-fire (LIF) neuron model, a typical model of cortical neurons \citep{Koch1999}.
+We drive the population of FHN or LIF neurons with a common aperiodic random signal $s(t)$, and vary the levels of internal noise and heterogeneity in the population.
+Finally, we decode an output signal from the population by summing and filtering the spikes from individual neurons, and compare the similarity of the decoded output signal with the input signal using mutual information.
+Stochastic offset is modeled as zero-mean Gaussian white noise with standard deviation $\sigma_\eta$, independently and identically distributed across all neurons (see \textsc{\nameref{scn:methods}}).
 Heterogeneity was added by randomly selecting the bias voltages $b_i$ from a uniform interval of varying width.
-We report how noise and heterogeneity both act to increase information transmission through these populations.
-They achieve this using three common mechanisms---linearizing the response of a population of neurons to a stimulus, desynchronizing neuronal firing, and decreasing the reaction time of a population of neurons to a stimulus---which we examine in detail.
+
+We report how noise and heterogeneity both act to increase information content in these populations, but in a complex, non-independent manner.
+They achieve this using two common mechanisms---linearizing the response of a population of neurons to a stimulus, and desynchronizing neuronal firing---that we examine in detail.
 
@@
 \section{Results}
-\scnlabel{results}
+\label{scn:results}
 
-We performed numerical experiments analyzing how noise and heterogeneity affect the information transmission capacity of populations of both FitzHugh-Nagumo (FHN) neurons and leaky integrate-and-fire (LIF) neurons, to test whether heterogeneity has similar effects to noise in increasing information transmission.
-Specifically, trials were performed using four distinct levels of random heterogeneity across the bias voltages $b_i$, the levels being none, low, medium, and high, corresponding to 0, 0.1, 0.5, and 2 times the input signal standard deviation $\sigma_s = 0.1$ (where the highest level spans the expected range of the input signal).
-At each level of heterogeneity, the standard deviation $\sigma_\eta$ of a stochastic offset---applied to the voltage parameter of the neuron model---was varied logarithmically from an insignificant level ($\sigma_\eta = 10^{-9}$ for FHN neurons and $\sigma_\eta = 10^{-6}$ for LIF neurons) to levels 10 times the magnitude of the input signal ($\sigma_\eta = 10^0$), in increments of $\log_{10}(\Delta\sigma_\eta) = 0.1$.
+We performed numerical experiments analyzing how noise and heterogeneity affect the information capacity of populations of both FitzHugh-Nagumo (FHN) neurons and leaky integrate-and-fire (LIF) neurons.
+Heterogeneity was added to a population by selecting the firing thresholds ${b_i}$ from a uniform distribution $\mathcal{U}(-b_r, b_r)$, where $b_r$ controls the degree of heterogeneity, with larger $b_r$ corresponding to a more heterogeneous population (on average).
+Noise was added to the voltage parameter of the neuron model via a stochastic offset $\eta(t)$, Gaussian distributed with zero-mean and standard deviation $\sigma_\eta$, and independent and identically distributed for each neuron.
 For the remainder of this paper, we will use the term noise to refer specifically to this stochastic offset.
-At each level of heterogeneity and noise, 25 trials were performed to achieve an accurate estimate of the mean information transmission, with each trial using a unique set of random biases and a unique set of stochastic offsets added to the membrane voltage.
+At each level of heterogeneity and noise, 25 trials were performed to achieve an accurate estimate of the mean information content, with each trial using a unique set of random biases and a unique set of stochastic offsets added to the membrane voltage.
 
@@
 The mutual information for each level of heterogeneity and noise was computed as the average mutual information from all 25 trials at those levels.
-See \scn{methods} for the full details of the experimental setup.
+See \textsc{\nameref{scn:methods}} for the full details of the experimental setup.
 
-\fig{info} shows the results of these experiments.
-In the homogeneous populations, a strong stochastic resonance effect is present, in which the transmitted information (\ie mutual information between the input and decoded output signals) was optimized for noise levels of $\sigma_\eta \approx 3 \times 10^{-3}$ (FHN neurons) and $\sigma_\eta \approx 1 \times 10^{-2}$ (LIF neurons).
-These noise levels are about 30 times (FHN neurons) or 10 times (LIF neurons) smaller than the input signal standard deviation of $\sigma_s = 0.1$.
-This makes them small enough not to significantly degrade the input signal, while being large enough to elicit stochastic resonance effects.
+Figure~\ref{fig:infohetero} summarizes the effect of neuronal heterogeneity on the information content of a population of LIF neurons.
+At low noise levels, we observe a resonance effect in the heterogeneity parameter---there is an optimal level of heterogeneity that maximizes the information content of the population.
+Furthermore, this optimal level of heterogeneity is around one to two times the root mean square amplitude of the input signal.
+This indicates that optimal heterogeneity corresponds to having neuron firing thresholds that span the full range of input signal values.
+Too little heterogeneity means that some input signal components cannot be uniquely encoded by the population, since these values will be either well below or well above the firing thresholds of all neurons in the population, and all neurons will fire approximately constantly for these values.
+Too much heterogeneity results in some neurons that have firing thresholds well below or well above all signal values, causing them to either fire tonically independent of the input signal value, or to fire for no part of the input signal, respectively.
+This type of resonance in heterogeneity was recently observed by \cite{Mejias2012}, both theoretically and in numerical experiments.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/bw/figure1_info.eps}
+    \includegraphics{../figures/bw/infohetero.eps}
   \fi
   \caption{
-    \textbf{Heterogeneity increases information transfer in populations of FHN and LIF neurons.} We simulated populations of 64 FHN neurons and 64 LIF neurons, varying the level of noise ($x$-axis) and heterogeneity (individual traces), and measured the mutual information between the aperiodic input signal $s(t)$ (where $\sigma_s = 0.1$) and the output signal $r(t)$ ($y$-axis).
-Moderate to high levels of heterogeneity resulted in an increase in information transmission at low to moderate noise levels in both FHN and LIF neurons.
-The increases in FHN neurons were modest, with a strong stochastic resonance effect still exhibited at all levels of heterogeneity.
-The increases in LIF neurons were much more pronounced, with moderate to high levels of heterogeneity surpassing the results achieved by stochastic resonance.
+    \textbf{Heterogeneity increases information content of a population of LIF neurons, and exhibits a resonance effect.}
+    Adding heterogeneity to a population of LIF neurons increases their ability to represent an input signal, but only for lower levels of noise.
+At higher levels of noise, the noise alone is able to fully desynchronize neuronal firing and linearize neuron responses, so adding heterogeneity offers no additional benefit.
   }
-  \figlabel{info}
+  \label{fig:infohetero}
 \end{figure}
 
-For both FHN and LIF neurons, heterogeneity also increases the information transmission of the population, specifically for noise levels below the optimal stochastic resonance noise level.
-In FHN neurons, the benefits of adding heterogeneity to neuron biases are modest (see \scn{phase}), resulting in a significant increase in information transmission, but not achieving the same levels of information transmission as stochastic resonance (\fig{info}A).
-For LIF neurons, the benefits of heterogeneity in neuron biases are pronounced, with moderate to high levels of heterogeneity achieving the same or even better levels of information transmission---across a wide range of noise values---as the optimal level of noise (\fig{info}B).
-
-In the remaining experiments, we consider three mechanisms that allow both noise and heterogeneity to improve signal transmission in populations of neurons: 1) putting neurons out of phase with one another, 2) decreasing the reaction time of a population of neurons to a stimulus, and 3) linearizing the response of a population of neurons to a stimulus.
-Though noise and heterogeneity affect neurons differently, they both act to achieve these three goals.
-None of the three factors is sufficient in and of itself to explain the effects of stochastic resonance and heterogeneity on nonlinear dynamical systems, such as the populations of FHN and LIF neurons studied.
-Nevertheless, by examining each factor individually, we gain a deeper understanding of how both stochastic resonance and heterogeneity operate in nonlinear dynamical systems, and ultimately show that the two concepts are fundamentally connected (see \scn{discussion}).
-
-\subsection{Phase}
-\scnlabel{phase}
-
-Consider the following: a group of perfectly homogeneous neurons with no noise, such that their dynamics are completely deterministic.
-Begun with the same initial conditions, they will follow exactly the same trajectory in phase space, since they all receive the same input signal and obey the same set of differential equations.
-Such a population of simulated neurons will exhibit perfectly synchronized firing events, and therefore the neurons contribute redundant information: any signal that could be decoded from the firing events of this population of neurons could also be decoded from the spikes of a single neuron.
-These neurons are perfectly in phase: at any time, they are all in the same phase of their firing cycle, and in the same position in phase space.
-
-Adding independent noise to each neuron desynchronizes the firing events of the population over time by causing diffusion of the neuronal phases.
-A small random perturbation causes a slightly advanced or delayed firing of a neuron, creating a small shift in the phase of that neuron.
-Specifically, a firing neuron follows a stable limit cycle attractor in phase space; perturbations tangent to the limit cycle result in diffusion along the limit cycle, while perturbations perpendicular to the limit cycle rapidly return the the stable attractor \citep{Tomita1974}.
-Over time, this diffusion along the limit cycle results in the population of neurons having a wide distribution of phases (\fig{phase}), and the neurons fire asynchronously.
+At higher levels of noise, heterogeneity offers no benefit to LIF neuron populations (Figure~\ref{fig:infohetero}).
+To examine the interactions between noise and heterogeneity further, we conducted the same experiment across a wide range of noise and heterogeneity values.
+Figure~\ref{fig:infocontour} illustrates the effects of these interactions on the information content of a population of LIF or FHN neurons.
+In LIF neurons, the population is best able to represent the input signal with a moderate amount of heterogeneity ($b_r \approx 1.5 \times 10^{-1}$) and no noise ($\sigma_\eta = 1.0 \times 10^{-4}$).
+Given this optimal level of heterogeneity, adding noise only serves to decrease the information content of the system.
+But if no heterogeneity is present in the system, then adding the right amount of noise ($\sigma_\eta \approx 1.0 \times 10^{-2}$) optimizes information content (Figure~\ref{fig:infonoise}).
+This shows that the effects of noise and heterogeneity are not independent.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/bw/figure2_phase.eps}
+    \includegraphics[width=\textwidth]{../figures/bw/infocontour.eps}
   \fi
   \caption{
-    \textbf{Noise and heterogeneity both desynchronize neurons.} To measure whether noise and heterogeneity have a desynchronizing effect on populations of neurons, we measured the standard deviation of neuron phases across the population (see \scn{methods}), where a smaller standard deviation indicates increased synchronization.
+    \textbf{Noise and heterogeneity show nonlinear interactions in their benefits to information representation.} We measured the mutual information (contours every 0.15 bits) across many levels of noise and heterogeneity.
+In LIF neurons, both noise and heterogeneity show independent resonance effects: given that only one of noise or heterogeneity is present, there is a non-zero and finite level of noise or heterogeneity that optimizes information content.
+Setting both noise and heterogeneity to their optimal levels fails to further improve the representational ability of the population.
+Both parameters use the same underlying mechanisms of desynchronization and linearization, so once the population has been desynchronized and linearized by one parameter, adding the other cannot further improve information content.
+Heterogeneity is unable to fully desynchronize FHN neurons (see \textsc{\nameref{scn:desync}}), so information content in these neurons is maximized for a particular level of noise.
+  }
+  \label{fig:infocontour}
+\end{figure}
+
+\begin{figure}
+  \ifx\hidefigures\undefined
+    \centering
+    \includegraphics[width=\textwidth]{../figures/bw/infonoise.eps}
+  \fi
+  \caption{
+    \textbf{Noise increases information content of homogeneous populations of FHN and LIF neurons.} In both LIF (left) and FHN (right) populations, noise exhibits a stochastic resonance effect, where adding noise to the homogeneous population (black line) improves the population's information content.
+This stochastic resonance effect disappears in heterogeneous populations of LIF neurons, but still occurs in heterogeneous populations of FHN neurons.
+  }
+  \label{fig:infonoise}
+\end{figure}
+
+We propose that noise and heterogeneity both improve the information representation ability of a population of neurons by 1) desynchronizing the firing of neurons in the population, and 2) linearizing the response of the population to an input signal.
+These shared mechanisms result in the nonlinear interaction between noise and heterogeneity in terms of information content: once either noise or heterogeneity has adequately desynchronized a population, the addition of the other cannot further desynchronize or linearize, and therefore provides no additional increase in information representation.
+In the following two sections, we present additional experimental results to investigate and quantify the mechanisms of desynchronization and linearization.
+
+We also note that there is a difference between FHN and LIF neurons in terms of noise and heterogeneity, and their effects on information content.
+Heterogeneity is much less effective at increasing information content in FHN neurons than LIF neurons (Figures~\ref{fig:infocontour}~and~\ref{fig:infonoise}).
+This is explained by the inability of heterogeneity to fully desynchronize firing in FHN neurons (see \textsc{\nameref{scn:desync}}), which is in turn due to the nature of the FHN tuning curves.
+We revisit this result in the \textsc{\nameref{scn:discussion}}.
+
+\subsection{Desynchronization}
+\label{scn:desync}
+
+Consider the following: a group of perfectly homogeneous neurons with no noise, such that their dynamics are completely deterministic.
+Begun with the same initial conditions, they will follow exactly the same trajectory in phase space because they all receive the same input signal and obey the same set of differential equations.
+Such a population of simulated neurons will exhibit perfectly synchronized firing events, and therefore the neurons contribute redundant information.
+Any signal that could be decoded from the firing events of this population of neurons could also be decoded from the spikes of a single neuron.
+These neurons are perfectly synchronized; at any time, they are all in the same phase of their firing cycle, and in the same position in phase space.
+
+Adding independent noise to each neuron desynchronizes the firing events of the population by causing diffusion of the neuronal phases \citep{Stocks2001a}.
+A small random perturbation causes a slightly advanced or delayed firing of a neuron, creating a small shift in the phase of that neuron.
+Specifically, a firing neuron follows a stable limit cycle attractor in phase space; perturbations tangent to the limit cycle result in diffusion along the limit cycle, while perturbations perpendicular to the limit cycle rapidly return the the stable attractor \citep{Tomita1974}.
+Over time, this diffusion along the limit cycle results in the population of neurons having a wide distribution of phases (Figure~\ref{fig:phase}), and the neurons fire asynchronously (Figure~\ref{fig:syncraster}).
+
+\begin{figure}
+  \ifx\hidefigures\undefined
+    \centering
+    \includegraphics[width=\textwidth]{../figures/bw/phase.eps}
+  \fi
+  \caption{
+    \textbf{Noise and heterogeneity both desynchronize neurons.} To measure whether noise and heterogeneity have a desynchronizing effect on populations of neurons, we measured the standard deviation of neuron phases across the population (see \textsc{\nameref{scn:methods}}), where a smaller standard deviation indicates increased synchronization.
 Heterogeneity is seen to significantly desynchronize both FHN and LIF neurons, specifically across low to moderate levels of noise.
@@
   }
-  \figlabel{phase}
+  \label{fig:phase}
 \end{figure}
 
-This previous example is, of course, artificial: it requires that all neurons are \emph{exactly} the same, and start with \emph{exactly} the same initial conditions.
+\begin{figure}
+  \ifx\hidefigures\undefined
+    \centering
+    \includegraphics[width=\textwidth]{../figures/bw/syncraster.eps}
+  \fi
+  \caption{
+    \textbf{The effects of noise and heterogeneity on LIF neuron firing.}
+    Spike rasters illustrate the effects of noise and heterogeneity parameters on neuronal firing in a population of 64 LIF neurons.
+With low noise and heterogeneity, neuronal firing is highly synchronized.
+Adding a moderate amount of noise or heterogeneity desynchronizes this firing.
+Adding a moderate level of both parameters cannot provide further desynchronization, since firing is already quite desynchronized.
+High levels of noise cause rapid firing in almost all neurons, degrading the ability of the population to represent the input signal.
+High levels of heterogeneity cause the signal to be entirely above or below the firing thresholds of many neurons, again degrading representational ability.
+Noise: $\sigma_\eta = $$10^{-3}$ (low), $10^{-2}$ (moderate), $10^{-1}$ (high).
+Heterogeneity: $b_r = $$10^{-2}$ (low), $10^{-1}$ (moderate), $10^{0}$ (high).
+  }
+  \label{fig:syncraster}
+\end{figure}
+
+This previous example is, of course, artificial.
+It requires that all neurons are \emph{exactly} the same, and start with \emph{exactly} the same initial conditions.
 In a more realistic scenario, one expects that small differences in neuronal parameters or small amounts of noise will cause neurons to desynchronize over a long period of time.
@@
 
-Similar to noise, heterogeneity acts to desynchronize neurons.
-Adding heterogeneity to a population significantly increases the range of distribution of neuron phases, thereby desynchronizing populations of both FHN and LIF neurons (\fig{phase}).
-Compared with a noiseless homogeneous population, where changes in an input signal will cross the common bias of all neurons at the same time and thereby put all neurons in phase with one another, changes in an input signal in a heterogeneous population will sequentially cross the biases of different neurons, causing some to begin firing before others.
-For example, an increasing input signal will cause neurons with lower biases to begin firing before those with higher biases, resulting in desynchronized firing within the population of neurons.
+Similar to noise, heterogeneity acts to desynchronize neurons \citep{Burton2012}.
+Adding heterogeneity to a population increases the range of distribution of neuron phases (Figure~\ref{fig:phase}), thereby desynchronizing populations of both FHN and LIF neurons (Figure~\ref{fig:syncraster}).
+In a homogeneous population, changes in the input signal cross the common firing threshold of all neurons at the same time, resulting in synchronous firing.
+Contrast this with a heterogeneous population, where changes in the input signal sequentially cross the thresholds of neurons, causing some to begin firing before others.
+For example, an increasing input signal will cause neurons with lower thresholds to begin firing before those with higher thresholds, resulting in desynchronized firing within the population of neurons.
 
-High frequency input signals may cross the biases of many neurons in a heterogeneous population at the same time, causing a large group of neurons to begin firing simultaneously.
-This effect does not have a significant synchronizing effect on LIF neurons, since the LIF tuning curves used have significantly different firing rates for different sizes of suprathreshold signal (\fig{tuning}).
+High frequency input signals may cross the thresholds of many neurons in a heterogeneous population at the same time, causing a large group of neurons to begin firing simultaneously.
+LIF neurons are still able to desynchronize quickly after such a stimulus, since the LIF tuning curves used have different firing rates for different sizes of suprathreshold signal (Figure~\ref{fig:tuning}).
 Specifically, a high-frequency increase in signal amplitude starts neurons firing at approximately the same time, but starts them firing at different rates, allowing them to quickly desynchronize.
-FHN neurons, however, all have around the same firing rate regardless of the strength of the suprathreshold signal (\fig{tuning}), and heterogeneity with higher-frequency signals has a reduced effect, since these signals cause many neurons to begin firing at once, unable to desynchronize due to their common firing rate.
-This discrepancy between LIF and FHN neurons results in heterogeneity having a large effect on phase in LIF neurons (\fig{phase}B), whereas the desynchronizing effect of heterogeneity in FHN neurons is much more modest (\fig{phase}A).
+FHN neurons, however, all have around the same firing rate regardless of the strength of the suprathreshold signal (Figure~\ref{fig:tuning}), and heterogeneity with higher-frequency signals has a reduced effect, since these signals cause many neurons to begin firing at once, unable to desynchronize due to their common firing rate.
+This discrepancy between LIF and FHN neurons results in heterogeneity having a large effect on phase in LIF neurons (Figure~\ref{fig:phase}A), whereas the desynchronizing effect of heterogeneity in FHN neurons is much more modest (Figure~\ref{fig:phase}B).
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Response time}
+\subsection{Linearization}
 
-Both noise and heterogeneity benefit the response time of a population of neurons, \ie how quickly the population responds to a rapidly changing signal.
-To measure this, we presented populations of FHN and LIF neurons with a step function and observed how well the population can follow the step function (\fig{step}).
-Though noise and heterogeneity both offer decreased response time, they achieve this using different mechanisms.
+Intrinsic noise makes neuron tuning curves more linear;
+this phenomenon underlies many results in the stochastic resonance literature
+\citep{Chialvo1997}.
+Noise allows a neuron to respond to input signals below its firing threshold,
+by occasionally raising the neuron's membrane voltage high enough
+to cross the threshold and trigger an action potential.
+This results in a non-zero firing rate for subthreshold signals,
+with the firing rate proportional to the strength of the signal.
+In type~II neuron models (\emph{e.g.} the FHN model),
+noise can also slow firing for signals slightly above the firing threshold,
+by momentarily pushing the neuron away from its limit cycle and back to near its unstable equilibrium.
+Overall, the addition of noise has the effect
+of linearizing the response of the neuron near its firing threshold,
+resulting in a smoother transition from the silent to tonic firing regimes (Figure~\ref{fig:tuning}).
+This result has been previously observed and extensively studied \citep{Stocks1996,Chialvo1997,Chance2002,Prescott2002,Shu2003}.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/bw/figure3_step.eps}
+    \includegraphics[width=\textwidth]{../figures/bw/tuningnoisy.eps}
   \fi
   \caption{
-    \textbf{Noise and heterogeneity both improve the step response of neuronal populations.} We excited populations of 100 FHN neurons and 100 LIF neurons using a step function, and measured the responses of populations without noise or heterogeneity ($b_i = 0, \sigma_\eta = 0$), with noise ($b_i = 0, \sigma_\eta = 10^{-3}$), and with heterogeneity ($b_i \in [-0.01,0.01], \sigma_\eta = 0$).
-Each line depicts the average population response over 100 trials, and the surrounding region shows one standard deviation.
-FHN populations without noise or heterogeneity show a considerable delay in responding to the step function, because all neurons are close to their stable equilibrium.
-Both noise and heterogeneity significantly reduce this delay, and also reduce oscillations by desynchronizing neurons (\fig{phase}).
-LIF populations only exhibit a mild initial delay, which noise and heterogeneity both correct; in LIF neurons, the main benefit to the step response is reducing oscillations through desynchronization.
-  }
-  \figlabel{step}
-\end{figure}
-
-Noise increases the response time of neurons by keeping them away from their stable equilibria during low points in the input signal.
-By presenting a population of 100 FHN neurons with a step input function, we found that a homogeneous population with no noise showed a considerable delay between the leading edge of the input step function and the response of the population (\fig{step}).
-Adding noise to the population decreases the response such that it is negligible, with a high proportion of the neurons responding immediately to the leading edge of the step.
-Noise produces this benefit by keeping neurons away from their stable equilibria during the below-threshold portions of the input signal (\ie before the leading edge of the step), allowing them to respond quickly to above-threshold portions of the input signal (\ie after the leading edge of the step).
-Neurons not subject to noise, on the other hand, are brought very close to their stable equilibria during the subthreshold portion of the input signal, and it takes a mildly suprathreshold signal---as provided by the step function---a long time to bring the neuron away from this equilibrium and into the tonic firing regime.
-
-Heterogeneity also acts to increase the response time of a population of neurons.
-It does not directly keep neurons away from their equilibrium points, but by spreading neuron biases across a wide range, it ensures that neurons remain in different states of activation.
-Upon receiving a low input signal, most neurons will be pushed towards their stable equilibria, but neurons with higher firing thresholds will be pushed there more quickly than neurons with lower thresholds.
-Specifically, at the rising edge of the signal, neurons with low thresholds will be less settled to their equilibria than those with high thresholds, allowing them to respond more quickly.
-Furthermore, the high input signal will push neurons with low thresholds away from their equilibria more quickly than those with high thresholds, again resulting in quick activation of low threshold neurons.
-This effect can be observed in the response of the FHN population to a step function (\fig{step}), where the heterogeneous population responds more quickly than the noiseless homogeneous population, but still less quickly than the noisy homogeneous population.
-
-\subsection{Linearization}
-
-Intrinsic noise makes neuron tuning curves more linear, using a mechanism similar to subthreshold stochastic resonance.
-For input signals below the neuron's threshold, noise operates in a manner identical to subthreshold stochastic resonance, occasionally raising the neuron's membrane voltage high enough to cross the threshold and trigger an action potential.
-For input signals above the neuron's threshold, noise causes the membrane voltage to occasionally fall below the threshold, slowing the firing rate slightly compared with the noise-free case.
-For signals well below the neuron's threshold, the noise is not sufficient to cause any spikes, and for signals well above the threshold, the noise is insufficient to stop any spikes: these portions of the tuning curve remain unchanged between the noise-free and noisy cases.
-Overall, the addition of noise has the effect of linearizing the response of the neuron near its firing threshold, resulting in a smoother transition from the silent to tonic firing regimes (\fig{tuning}).
-This result has been previously observed and extensively studied \citep{Stocks1996,Chialvo1997}.
-
-\begin{figure}
-  \ifx\hidefigures\undefined
-    \centering
-    \includegraphics[width=\textwidth]{../figures/bw/figure4_tuningnoisy.eps}
-  \fi
-  \caption{
-    \textbf{Noise linearizes neuron tuning curves.} We probed FHN neurons and LIF neurons using constant inputs at various voltages, using varying levels of noise, and measured the output firing rate.
-Noise has the effect of making the tuning curves more linear, specifically eliminating the sharp threshold in both FHN and LIF neurons.
+    \textbf{Noise linearizes neuron tuning curves.} We probed LIF and FHN neurons using constant inputs at various voltages, using varying levels of noise, and measured the output firing rate.
+Noise has the effect of making the tuning curves more linear, specifically eliminating the sharp threshold in both LIF and FHN neurons.
 The main mechanism is that of subthreshold stochastic resonance: noise allows subthreshold inputs to cause spikes by occasionally pushing the membrane voltage above the threshold.
-This occurs in both FHN and LIF neurons, allowing them to represent subthreshold signal components.
+This occurs in both both neuron models, allowing them to represent subthreshold signal components.
 In FHN neurons, small levels of noise can also cause firing rates for small suprathreshold signals to drop slightly by briefly pushing the neuron away from its firing limit cycle in state space, though interestingly this does not happen for larger noise levels.
-Large levels of noise in FHN neurons can cause large fluctuations which jump from one point on the limit cycle to another point much farther ahead, causing elevated firing rates.
+Large levels of noise in FHN neurons cause large membrane voltage fluctuations, resulting in jumps from one point on the limit cycle to another point much farther ahead, causing elevated firing rates.
 In LIF neurons, noise added to small suprathreshold signals causes the neuron to fire slightly faster, because the noise can cause the rising membrane voltage to prematurely cross the threshold, causing the spike to occasionally occur sooner than it would in the noiseless case.
   }
-  \figlabel{tuning}
+  \label{fig:tuning}
 \end{figure}
@@
 Since our model uses a simple summing decoder, the population tuning curve is equal to the sum of the individual neuron tuning curves.
-For a population of neurons with uniformly varied random biases, the population tuning curve is much more linear than a single neuron's tuning curve, whereas in the homogeneous case the population's tuning curve is the same as the neurons' tuning curve (\fig{tuninghetero}).
-This linearizing effect of heterogeneity allows for better representation of an input signal.
+For a population of neurons with uniformly varied random firing thresholds, the population tuning curve is much more linear than a single neuron's tuning curve, whereas in the homogeneous case the population's tuning curve is the same as the neurons' tuning curve (Figure~\ref{fig:tuninghetero}).
+The linearizing effect of heterogeneity thereby allows for better representation of an input signal.
+This result is related to the original idea from \cite{Hubel1962} that variation in neural receptive fields provides a better basis for representing an input signal; this idea reappears in the recent heterogeneity literature \citep{Angelo2012}.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/bw/figure5_tuninghetero.eps}
+    \includegraphics[width=\textwidth]{../figures/bw/tuninghetero.eps}
   \fi
   \caption{
-    \textbf{Heterogeneity linearizes the response of a neuronal population to a stimulus.} Panel A shows the tuning curves for a heterogeneous population of $N = 9$ neurons, where one neuron has been chosen to have a bias of $-0.15$ and the other neurons have random biases uniformly sampled from the interval $[-0.15,0.15]$.
-Panel B shows how the population response (heavy solid line) is much closer to a linear response (heavy dot-dashed line) than a single LIF tuning curve (heavy dashed line), even though the heterogeneous biases are far from evenly spaced.
+    \textbf{Heterogeneity linearizes the response of a neuronal population to a stimulus.} Panel A shows the tuning curves for a heterogeneous population of $N = 9$ neurons, where one neuron has been chosen to have a firing threshold of $-0.15$ and the other neurons have random thresholds uniformly sampled from the interval $[-0.15,0.15]$.
+Panel B shows how the population response (heavy solid line) is much closer to a linear response (heavy dot-dashed line) than a single LIF tuning curve (heavy dashed line), even though the heterogeneous thresholds are far from evenly spaced.
 The light traces indicate how the population response is the average of the individual neuron tuning curves, where each colored curve is a scaled version of the tuning curves in panel A (with scaling factor $1/N$), and the population response is the sum of these curves.
-Using a larger population of neurons (\ie increasing $N$) results, on average, in a more linear population response.
+Using a larger population of neurons (\emph{i.e.,} increasing $N$) results, on average, in a more linear population response.
   }
-  \figlabel{tuninghetero}
+  \label{fig:tuninghetero}
 \end{figure}
@@
 \section{Discussion}
-\scnlabel{discussion}
+\label{scn:discussion}
 
-Our results show that noise and heterogeneity in populations of neurons can improve the ability of a neuronal population to encode and transmit a signal.
-Both noise and heterogeneity can be understood as using similar mechanisms to achieve this increase in information transmission, specifically by 1) desynchronizing the neurons within the population, 2) improving the reaction time of the population to a rapidly changing signal, and 3) linearizing the response of the population.
-Consequently, both noise and heterogeneity may be important mechanisms for optimizing information flow in neural systems.
+Our results show that noise and heterogeneity both improve neuronal representation of an input signal, and that these benefits are not independent.
+Adding noise to a homogeneous neuronal population increases information content, while the same amount of noise degrades information contained by a heterogeneous population (Figures~\ref{fig:infocontour}~and~\ref{fig:infonoise}).
+This suggests that the benefits of noise and heterogeneity to information content can be explained in terms of shared mechanisms, which we identify as 1) desynchronizing the neurons within a population, and 2) linearizing the response of the population.
 
-To characterize the benefits of both noise and heterogeneity on signal transmission in populations of neurons, we can separate them into two categories: \emph{temporal} properties of neurons, and \emph{rate} properties of neurons.
-The temporal category is concerned with the desynchronizing and reaction time effects of noise and heterogeneity, and the rate category is concerned with the linearizing property of both noise and heterogeneity on a neuronal population.
+The dynamics of a neuron can be viewed as a combination of temporal properties and rate properties.
+The rate view of neurons treats them as being no more than input-output (or tuning) curves.
+That is, all temporal dynamics are ignored, and a neuron is reduced to a nonlinear function which turns an input signal into an output signal.
+The mechanism of linearization affects only the rate properties of neurons, and it is these linearization benefits that are observed in suprathreshold stochastic resonance studies that examine the benefits of noise in populations of binary threshold units \citep{Stocks2000,Stocks2001a,McDonnell2006}.
+This is because binary threshold units have no temporal properties---they are either on or off depending on the current input signal and noise---and cannot be affected by the temporal benefits of noise.
+The linearization effects of noise on nonlinear systems have been previously studied (see \cite{Stocks1996} for a review).
+We have observed that heterogeneity also achieves these linearization effects across a population of neurons, as previously reported by \cite{Eliasmith2003}.
 
-The temporal benefits of noise and heterogeneity can only be observed in spiking neuron models, such as the FHN and LIF models examined in this paper.
+Desynchronization operates by affecting the temporal properties of neuronal firing, and is only observed in spiking neuron models, such as the FHN and LIF models examined in this paper.
 The output of these models depends not only on the the output firing rate for a given input signal, as determined by the neuron tuning curve, but also by the timing of individual spike events.
 When nonlinear dynamical models of neurons are examined in a stochastic resonance simulation, as in \cite{Stocks2001}, the results are a combination of the temporally-based desynchronizing and response-time effects of noise, as well as the rate-based linearizing effects of noise.
+Desynchronized neurons provide a better temporal basis for representing a signal, since each neuron is able to fire independently and code for a different temporal component of the signal.
+Even mild correlations between neurons are known to impede a population's representational ability \citep{Zohary1994}.
 
-The rate view of neurons treats them as being no more than input-output (or tuning) curves.
-That is, all temporal properties are ignored, and a neuron is reduced to a nonlinear function which turns an input signal into an output signal.
-Suprathreshold stochastic resonance effects observed in binary threshold units deal only with the rate benefits of noise \citep{Stocks2000,Stocks2001a,McDonnell2006}.
-This is because binary threshold units have no temporal properties---they are either on or off depending on the current input signal and noise---and cannot be affected by the temporal benefits of noise.
-The linearization effects of noise on nonlinear systems have been previously studied (see \cite{Stocks1996} for a review).
-We have observed that heterogeneity can also achieve these linearization effects across a population of neurons, which has previously been reported by \cite{Eliasmith2003}.
+Our experimental setup was motivated by stochastic resonance experiments such as \cite{Stocks2001}, as well as computational \citep{Brody2003} and experimental \citep{Padmanabhan2010} work looking at heterogeneity.
+All these experiments use a common input signal presented to a number of neurons, with added noise or heterogeneity.
+The common input signal may seem artificial, but it allows us to examine how the properties of individual neurons---including sign of the input signal (``on'' and ``off'' neurons), bias current, and membrane voltage---affect the ability of the neuronal population to encode an input signal.
+These heterogeneities in the neurons can be viewed as modulating the input signal presented to each neuron, such that each neuron receives a unique input current.
+In the case of homogeneous populations, all neurons do receive the same input current.
+This is clearly artificial, and emphasizes why theoreticians should include heterogeneity when modeling neural behaviour.
 
-One important difference between the LIF and FHN neuron models is that the LIF model is a type~I neuron model, and the FHN model is a type~II neuron model (though LIF neurons are not prototypical type~I neurons \citep{Mato2008}, they are qualitatively more similar to this category).
-The difference between type~I and type~II neuron models is that type~I models undergo a saddle-node bifurcation as the input current (the bifurcation parameter) increases, whereas type~II models undergo a subcritical Hopf bifurcation \citep{Mato2008}.
-Qualitatively, this means that type~I models have continuous tuning curves, \ie they can have arbitrarily small firing rates, whereas type~II models have a discontinuity at their firing threshold, and cannot fire slower than a given rate.
-The fact that the LIF model is a type~I neuron model, in addition to other factors, makes it a better model of cortical neurons \citep{Koch1999}.
-The LIF model is therefore more relevant when examining population encoding as proposed to occur in cortex, though other neuron models may be preferable for certain sensory coding situations.
+In our experiments, we saw significantly different results for the LIF neuron model versus the FHN neuron model.
+This raises two questions: why is there such a large difference between the two models, and does either model better describe the neural systems we wish to study?
+One way to differentiate between the two models is using bifurcation theory, where the LIF model is a type~I neuron model, and the FHN model is a type~II neuron model (though LIF neurons are not prototypical type~I neurons \citep{Mato2008}, they are qualitatively more similar to this category).
+The difference is that type~I models undergo a saddle-node bifurcation as the input current (the bifurcation parameter) increases, whereas type~II models undergo a subcritical Hopf bifurcation \citep{Mato2008}.
+Qualitatively, this means that type~I models have continuous tuning curves, \emph{i.e.,} they can have arbitrarily small firing rates, whereas type~II models have a discontinuity at their firing threshold, and cannot fire slower than a given rate.
+The fact that the LIF model is a type~I neuron model, in addition to other factors, has traditionally made it a better model of cortical neurons \citep{Koch1999}, though studies indicate that some neurons in mammalian (rat) cortex exhibit both type~I and type~II properties \citep{Tateno2004,Tsubo2007}.
 Previous numerical experiments that have examined stochastic resonance and heterogeneity together, such as \cite{Stocks2000,McDonnell2006}, have used binary threshold units in their simulations, which are qualitatively very similar to type~II neurons (but without the temporal component); the results of these experiments may have limited applicability to cortical neural systems.
 
-The heterogeneity of biases added to neuronal populations had a much more beneficial effect on information transmission in LIF neurons than in FHN neurons (\fig{info}).
-We hypothesize that this is because of key differences in the shape of the neuron tuning curves, combined with the fact that heterogeneity was only added to one neuron parameter, namely the bias voltage $b_i$.
-The FHN model, being a type~II neuron model, has a tuning curve that is qualitatively similar to a binary threshold unit, with a discontinuous increase in firing rate at the firing threshold, and offering little change in firing rate for signals above the threshold (\fig{tuning}A).
-On the other hand, the LIF model, being a type~I neuron model, has a continuous and monotonically increasing tuning curve, capable of having arbitrarily small firing rates, and producing significantly higher firing rates for input signals that are well above the firing threshold than for those slightly above the firing threshold (\fig{tuning}B).
+We found that heterogeneity was more beneficial to LIF neurons than in FHN neurons (Figure~\ref{fig:infocontour}).
+We hypothesize that this is because of key differences in the shape of the neuron tuning curves, combined with the fact that heterogeneity was only added to one neuron parameter, namely the firing threshold $b_i$.
+The FHN model, being a type~II neuron model, has a tuning curve that is qualitatively similar to a binary threshold unit, with a discontinuous increase in firing rate at the firing threshold, and offering little change in firing rate for signals above the threshold (Figure~\ref{fig:tuning}A).
+On the other hand, the LIF model, being a type~I neuron model, has a continuous and monotonically increasing tuning curve, capable of having arbitrarily small firing rates, and producing significantly higher firing rates for input signals that are well above the firing threshold than for those slightly above the firing threshold (Figure~\ref{fig:tuning}B).
 If a high-frequency signal crosses the firing thresholds of many FHN neurons in quick succession, all these neurons will begin firing both at approximately the same time and with approximately the same rate, taking a long time to desynchronize.
@@
 However, the signal will result in significantly higher firing rates for neurons with lower firing thresholds than for those with higher firing thresholds, allowing the group of neurons to desynchronize quickly.
-This key difference between the tuning curves of FHN and LIF neurons results in better desynchronization for LIF neurons when heterogeneity is only added to the bias voltage.
-We hypothesize that if heterogeneity were added to other neuron parameters, for example giving different FHN neurons different maximum firing rates, then FHN neurons would be able to desynchronize as quickly as LIF neurons and heterogeneity would offer a similar benefit in information transmission to stochastic resonance in FHN neurons.
-A system with this level of heterogeneity would require a more complex decoding system than the simple summing neuron model used in this paper, perhaps using the tuning curves to solve for the optimal linear decoders as described by \cite{Eliasmith2003}, and we leave this for future work.
+This key difference between the tuning curves of FHN and LIF neurons results in better desynchronization for LIF neurons when heterogeneity is only added to the firing threshold.
+
+We hypothesize that if heterogeneity were added to other neuron parameters, then heterogeneous populations of FHN neurons would be able to contain the same amount of information as noisy homogeneous FHN populations.
+For example, adding variation to the maximum firing rates of FHN neurons would allow them to desynchronize as quickly as LIF neurons.
+A system with this level of heterogeneity would require a more complex decoding system than the simple summing neuron model used in this paper, perhaps using the tuning curves to solve for the optimal linear decoders as described by \cite{Salinas1994} or \cite{Eliasmith2003}.
+Such a decoding scheme could take advantage of a key difference between noise and heterogeneity: the variability produced by heterogeneity is reproducible from trial-to-trial, whereas the variability due to noise is not.
+Thus heterogeneity allows for more complex decoding schemes, such as the combinatorial decoding scheme examined in \cite{Osborne2008}.
+We leave the extension of heterogeneity to additional neural parameters, and the requisite implementation of more complex decoding schemes, for future work.
 
@@
 Specifically, noise corresponds to variation within the phase space (state space) of a system over time, and heterogeneity corresponds to variation within the parameter space of a system across a population.
-It is important to emphasize that the heterogeneity used in this paper was completely random; the heterogeneous biases did not have to be specifically chosen or fine-tuned in any way, indicating that random heterogeneity could be implemented in a messy biological system like the brain.
-Finally, in analogy to stochastic resonance with noise, there is a critical level of heterogeneity that optimizes information transmission (\fig{infohetero}).
+It is important to emphasize that the heterogeneity used in this paper was completely random; the heterogeneous firing thresholds did not have to be specifically chosen or fine-tuned in any way, suggesting that random heterogeneity may be a natural implementation scheme adopted by a biological system like the brain.
+Finally, in analogy to stochastic resonance with noise, there is a critical level of heterogeneity that optimizes information transmission (Figure~\ref{fig:infohetero}).
 Thus noise and heterogeneity should be thought of as analogous mechanisms, but in different parts of the neuronal model (the state space versus the parameter space, respectively).
 
-\begin{figure}
-  \ifx\hidefigures\undefined
-    \centering
-    \includegraphics{../figures/bw/figure6_infohetero.eps}
-  \fi
-  \caption{
-    \textbf{Heterogeneity exhibits resonance effects in information transfer.} In an experiment similar to that shown in \fig{info}, the mutual information was measured between the input signal $s(t)$ and output signal $r(t)$ of populations of 64 LIF neurons with varying levels of noise and heterogeneity.
-The key difference is that in this figure, the radius of heterogeneity $b_r$ (where $b_i \in [-b_r, b_r]$) is varied along the $x$-axis, and each individual trace depicts a different noise level $\sigma_\eta$.
-At lower levels of noise ($\sigma_\eta = 1.0 \times 10^{-6}, 3.0 \times 10^{-3}$), there is a critical level of heterogeneity that optimizes information transmission.
-Specifically, this level of heterogeneity is between one and two times the standard deviation of the input signal $\sigma_s = 0.1$.
-At the optimal level of noise ($\sigma_\eta = 1.0 \times 10^{-2}$, as shown in \fig{info}), heterogeneity has no additional benefit: the noise itself is enough to allow for good information transmission.
-Finally, at high levels of noise ($\sigma_\eta = 1.0 \times 10^{-1}$), the noise overpowers the input signal, and no amount of heterogeneity can improve information transmission.
-This same experiment was performed with FHN neurons, but the change in information transmission due to heterogeneity was minimal, likely because the FHN tuning curves have little variation above the input stimulus level (see \scn{discussion}).
-  }
-  \figlabel{infohetero}
-\end{figure}
-
-The extent to which actual biological neural systems exploit either heterogeneity or stochastic resonance for improved information encoding and transmission remains unclear, and may vary significantly from system to system.
-One recent study shows that heterogeneity can be very important in encoding sensory information for certain electric fishes \citep{Marsat2010}.
+Both our results here and the results of numerous other researchers suggest that noise and heterogeneity may be important vehicles used by neural systems to optimize information flow.
+Yet the extent to which actual biological neural systems exploit either heterogeneity or stochastic resonance for improved information encoding and transmission remains unclear, and may vary significantly from system to system.
+A number of recent studies show that heterogeneity can be very important in encoding sensory information, such as for certain electric fishes \citep{Marsat2010} or in the rodent olfactory system \citep{Padmanabhan2010,Burton2012}.
 Another recent study shows SR effects in human auditory cortex, specifically that noise can increase correlation both within and between some cortical regions \citep{Ward2010}.
 More studies of this nature are required to determine both what neural systems show evidence of heterogeneity and stochastic resonance, and more importantly, whether these systems actually take advantage of heterogeneity and stochastic resonance for improved population coding.
-Our contribution is demonstrating \emph{how} a neural system could exploit either heterogeneity or stochastic resonance, or both, for improved information processing characteristics.
+The work presented here demonstrates the complex interaction between noise and heterogeneity due to the shared underlying mechanisms of desynchronization and linearization, and explains how these mechanisms lead to improved information processing characteristics.
 
@@
 \section{Models and Methods}
-\scnlabel{methods}
+\label{scn:methods}
 
-Our model consists of an input signal $s(t)$ encoded by a population of neurons, specifically having the same signal injected into the soma of all cells in the population (see \eqn{fhn} and \eqn{lif}).
+Our model consists of an input signal $s(t)$ encoded by a population of neurons, specifically having the same signal injected into the soma of all cells in the population (see Equations~\ref{eqn:fhn}~and~\ref{eqn:lif}).
 We ran simulations using both the FitzHugh-Nagumo (FHN) neuron model and the leaky integrate-and-fire (LIF) neuron model.
 Signal decoding is modeled by summing the spikes of each neuron in the population and low-pass filtering the result to obtain the output signal $r(t)$.
-This can be thought of as analogous to determining the somatic current of a single receiving neuron, where the filtering is performed by synapses.
+This is analogous to determining the somatic current of a single receiving neuron, where the filtering is performed by synapses.
 The input and output signals are then compared using mutual information, to measure how well the output signal encodes the input signal.
@@
   \end{array}
-  \eqnlabel{fhn}
+  \label{eqn:fhn}
 \end{equation}
@@
   \tau_{RC} \dot{v_i} = -v_i + \beta_i + \alpha \left( e_i s(t) + \eta_i(t) \right),
-  \eqnlabel{lif}
+  \label{eqn:lif}
 \end{equation}
@@
 The length of the refractory period is given by the refractory time constant $\tau_{ref}$, and during this time the neuron voltage is held at zero.
-Once the refractory period is finished, the neuron obeys \eqn{lif} until another spike occurs.
+Once the refractory period is finished, the neuron obeys Equation~\ref{eqn:lif} until another spike occurs.
 
@@
 where $N$ is the number of neurons in the encoding population, and $V_{th}$ is the firing threshold ($V_{th} = 0$ for the FHN model, and $V_{th} = 2$ for the LIF model).
-This model simulates a summing neuron which only receives the spikes (\ie parts of the neuron voltage $v_i$ greater than the threshold $V_{th}$, and then low-pass filters the sum of these spikes, a basic description of how a summing neuron membrane sums and filters input voltages.
+This model simulates a summing neuron which only receives the spikes (\emph{i.e.,} parts of the neuron voltage $v_i$ greater than the threshold $V_{th}$, and then low-pass filters the sum of these spikes, a basic description of how a summing neuron membrane sums and filters input voltages.
 
@@
 To compute the joint probability distribution $p(s,r)$ for discrete signals $s(t)$ and $r(t)$, the domains $S$ and $R$ are each divided into equal bins, and a joint histogram of the points $[s(t_i),r(t_i)]$ is formed.
-This histogram is used to approximate the joint probability function $p(s,r)$ by scaling it so the sum across all $s$ and $r$ is unity (\ie dividing the histogram by the total number of points $N$).
+This histogram is used to approximate the joint probability function $p(s,r)$ by scaling it so the sum across all $s$ and $r$ is unity (\emph{i.e.,} dividing the histogram by the total number of points $N$).
 For all simulations, we divided the domains $S$ and $R$ into 19 bins each.
@@
 The difficulty in comparing a delayed version of the input signal with the output signal is that it raises the question of what length of delay should be used.
-Each pair of noise and heterogeneity levels may conceivably have a unique optimal delay, and in order to not favor either noise or heterogeneity, the optimal delay (\ie the delay that produces the highest mutual information) should be calculated and used at each pair of noise and heterogeneity levels.
-In the interest of simplicity, we decided to instead compare the input and output signals with no delay, as is consistent with previous experiments \citep{Stocks2001}, and allows the benefits of noise and heterogeneity on response time to be measured.
+Each pair of noise and heterogeneity levels may conceivably have a unique optimal delay, and in order to not favor either noise or heterogeneity, the optimal delay (\emph{i.e.,} the delay that produces the highest mutual information) should be calculated and used at each pair of noise and heterogeneity levels.
+In the interest of simplicity, we decided to instead compare the input and output signals with no delay, as is consistent with previous experiments \citep{Stocks2001}.
 
@@
 
-\textbf{``On'' and ``off'' neurons.} To help minimize subthreshold stochastic resonance effects from significantly contributing to our results, we used populations that contained an equal number of ``on'' and ``off'' neurons, \ie neurons that both increase firing rates as the input signal increases, and increase firing rates as the signal decreases.
+\textbf{``On'' and ``off'' neurons.} To help minimize subthreshold stochastic resonance effects from significantly contributing to our results, we used populations that contained an equal number of ``on'' and ``off'' neurons, \emph{i.e.,} neurons that both increase firing rates as the input signal increases, and increase firing rates as the signal decreases.
 In terms of our model, this means that half the neurons of each population had $e_i = 1$, and the other half had $e_i = -1$.
@@
 where the function $\mathrm{wrap}(\alpha) = (\alpha + \pi \mod 2\pi) - \pi$ converts the angle $\alpha$ to its equivalent angle in the range $[-\pi,\pi)$.
-\fig{phase} presents the mean over all simulation time of the standard deviation of the phase.
+Figure~\ref{fig:phase} presents the mean over all simulation time of the standard deviation of the phase.
 
-\textbf{Simulation details.} The information transmission (\fig{info}) and phase (\fig{phase}) plots were computed using noise levels in the range $\sigma_\eta \in [10^{-9},1]$ for FHN neurons and in the range $\sigma_\eta \in [10^{-6},1]$ for LIF neurons, with the levels logarithmically spaced in steps of $10^{0.1}$.
-We chose the low end of these ranges to be an insignificant level of noise (\ie with information transmission qualitatively the same as the no-noise case) and increasing until the noise completely overpowered the signal (\ie only a negligible amount of signal information could be recovered).
-At each noise level, and each of the four levels of heterogeneity, 25 simulations were performed, each having a simulation time of 4.5 seconds, with the first 0.5 seconds of each simulation being discarded to allow any transients created by the initial conditions to settle.
+\textbf{Simulation details.} The information transmission plots with a noise axis (Figures~\ref{fig:infocontour}~and~\ref{fig:infonoise}) were computed using noise levels in the range $\sigma_\eta \in [10^{-6},1]$ for FHN neurons and in the range $\sigma_\eta \in [10^{-4},1]$ for LIF neurons, with the levels logarithmically spaced in steps of $10^{0.1}$.
+We chose the low end of these ranges to be an insignificant level of noise---where information content qualitatively the same as the noise-free case---and increasing until the noise completely overpowered the signal such that only a negligible amount of signal information could be recovered.
+Phase plots (Figure~\ref{fig:phase}) used a wider range of noise values, since desynchronization effects were observable starting at very small noise levels.
+For information plots with a heterogeneity axis (Figures~\ref{fig:infohetero}~and~\ref{fig:infocontour}), we chose $b_r \in [10^{-3}, 1]$, with points logarithmically spaced in steps of $10^{0.1}$.
+Again, the low end of this range corresponds to an insignificant level of heterogeneity (qualitatively the same as the homogeneous case), and the high end corresponds to the case where either information content reduced to insignificant levels (LIF neurons), or firing thresholds were so extreme that increasing them further would push the neuron model outside its valid range of operation (FHN neurons).
+
+At each combination of noise and heterogeneity levels, 25 simulations were performed, each having a simulation time of 4.5 seconds, with the first 0.5 seconds of each simulation being discarded to allow any transients created by the initial conditions to settle.
 The time step used varied depending on the noise level: larger levels of noise are more unstable and require a smaller time step, so a step of $\Delta t = 10^{-5}$ was used for $\sigma_\eta >= 0.1$, $\Delta t = 2.5 \times 10^{-5}$ for $\sigma_\eta >= 0.01$, and $\Delta t = 10^{-4}$ for $\sigma_\eta < 0.01$.
@@
 
-The step response plots (\fig{step}) represent the average population step response of a $N = 100$ neuron population over 100 trials.
-Each trial consists of presenting a population of $N = 100$ neurons with a step input $S_{step}(t)$ of 2 second duration, where the step input is equal to $-0.01$ for $t < 1.0$ and $0.01$ for $t >= 1.0$, with a time step $\Delta t = 10^{-4}$.
-The time shown in the plots corresponds to simulation time $t \in [0.9, 1.4]$, that is 0.1 seconds before the leading edge of the step to 0.4 seconds afterward.
-Output responses are scaled so that the average response for simulation time $t > 1.5$ corresponds to $0.01$, the final value of the step function.
-
-The tuning curve plots (\fig{tuning}) were all calculated empirically by measuring the firing rate of simulated neurons.
+The tuning curve plots (Figure~\ref{fig:tuning}) were all calculated empirically by measuring the firing rate of simulated neurons.
 The range of inputs measured was $[-0.2, 0.2]$, divided into 51 equally spaced points (intervals of $0.008$).
@@
 
-%% \bibitem[{LastName(2009)}]{Ref2009}
-%% LastName, A. (2009).
-%% \newblock Title for the first reference.
-%% \newblock \emph{Journal of the first reference}, \emph{3}, 18 -- 88.
-
-%% \bibitem[{Authors et~al.(2008)Author1, Author2, \&
-%%   Author3}]{Ref2008}
-%% Author1, A., Author2, A., \& Author3, A. (2008).
-%% \newblock Title for the second reference.
-%% \newblock \emph{Journal for the second reference}, \emph{5}, 188 -- 200.
-
 \end{thebibliography}

### 2014-01-20 paper/letter.tex
Third version of the paper, with figures.

--- 2014-01-20 paper/letter.tex
@@
 \begin{document}
-\begin{letter}{Editoral Board\\Neural Computation}
-\opening{Dear Sir or Madam:}
+\begin{letter}{Editorial Board\\Neural Computation}
+\opening{Dear Editor:}
 
-I am writing on behalf of myself and my co-authors, Matthew Scott and Chris Eliasmith, to present you with a manuscript that we have prepared for publication in Neural Computation, entitled ``Heterogeneity increases information transmission of neuronal populations''.
-This article examines the effects of heterogeneity on information transmission in populations of simulated neurons, and finds that it has a role to play in increasing the amount of information a population is able to encode and transmit about an input signal.
-We compare these results with previous results in the stochastic resonance field detailing how noise can benefit information transmission in neural populations, and find that heterogeneity has similar effects to noise in this respect.
+Please find enclosed a revised version of our paper
+entitled ``The competing benefits of noise and heterogeneity in neural coding''.
+In this revision, we address the further concerns of reviewer one.
+We respond to each comment individually,
+with the quoted comment presented in an indented paragraph,
+followed by our response.
 
-This research is important to the field because it advances the view that heterogeneity has a role to play in neural systems, and explains in detail why this is so.
-Research in stochastic resonance has done a great service by documenting the beneficial effects of noise in simulated, in vitro, and in vivo neural systems, but this same service has not yet been extended to heterogeneity.
-The goal of this manuscript is to provide theoretical foundations for understanding the mechanisms by which heterogeneity operates, and why it is important to neural systems.
-These foundations tie together the limited amount of existing research on the subject, and provide direction and motivation for future theoretical and experimental research into heterogeneity in neural systems.
+\ \newline
+{\large \bf Reviewer~1}
 
-Given the general interest of this paper and its potential to inspire future research in the area, we ask that you consider the enclosed manuscript for publication in Neural Computation.
-None of the material in the article has been published or is under consideration for publication elsewhere.
-We thank you for your time, and look forward to your reply.
+\begin{quotation}
+  The revised manuscript is substantially improved and may be of interest to the computational and experimental neuroscience community.
+The authors have done an especially good job reviewing the literature, and the greater clarity with which the relationship between stochastic resonance and heterogeneity in the manuscript is described improves the readability greatly.
+The authors should be commended for this.
+In addition, eliminating some of the other components of the initial submission makes the conclusions of the manuscript much more compelling.
+Addressing one major comment and the minor points of clarification below would help to drive home the results in this paper.
+
+\vspace{1em}
+\noindent Major point:
+\vspace{0.5em}
+
+The authors use the bias (the threshold of the neuron) as their metric for heterogeneity.
+The biological motivation needs to be addressed more clearly.
+Specifically, this form of diversity may be perceived as somewhat artificial, but if the authors could justify it with references or a few examples, this would go a long way.
+An alternative way to generate diversity, at least in the leaky integrate and fire neuron, would be to vary the time constants (tau) for the cells would it not?
+As the authors cite McCormick et al 1985, they should note that while the mean tau in that paper was 20 ms, the variance was +/-14 ms for regular spiking cells.
+Across neurons of different types the tau varies even more.
+Or vary tau ref, which would correspond to differential expression of channels with varied kinetics that contribute to differences in diversity of refractory times.
+Such a metric of diversity would be more realistic and distinguish this result from others wherein the use of the variable threshold has been employed, as the authors themselves concede with ``research into stochastic resonance with heterogeneity has focused on populations of binary threshold units, which fail to capture the dynamics of actual neurons''.
+This is especially important to really distinguish the novelty of this work from other in the field of heterogeneity where a threshold diversity is used.
+\end{quotation}
+
+We have added a few sentences to both the introduction and discussion to better motivate the use of heterogeneity in the bias (threshold) as the metric for heterogeneity in our paper.
+We agree that other forms of diversity are important, but as we note in the discussion, using forms of heterogeneity that affect the maximum firing rate of the neurons would require a more complex decoding scheme than what we have used (see, for example, Eliasmith 2003).
+We therefore leave this as future work.
+
+\begin{quotation}
+\noindent Minor points:
+\vspace{0.5em}
+
+The authors should motivate the choice of the stimulus better.
+In other examples, both experimental (see Mainen and Sejnowski, 1995) and theoretical (Galan et al 2008), the power spectrum of the input signal seems to be qualitatively different that the ones being used by the authors, or if this is a misunderstanding on my part, then the authors should clearly state this.
+As the statistics of the input shape reliability (Galan et al 2008) and encoding fidelity (RR de Ruyter van Steveninck et al 1997), it would be important to ensure that the stimulus parameters match those seen in real neurons and that changing the stimulus parameters within other domains of plausibility does not affect their result.
+Even if it does, it would be ok because it would highlight that different combinations of heterogeneity and noise that are optimal depending on the stimulus statistics.
+Either way this should be considered.
+\end{quotation}
+
+We have re-run all simulations using the alpha function-filtered white noise used by Mainen and Sejnowski (1995) and Gal\'an et al.
+(2008).
+This significantly changed some results, specifically the information landscape for FHN neurons, so we revised our discussion to reflect this.
+The results for FHN and LIF neurons are actually much more similar now than previously.
+
+\begin{quotation}
+In figure 5, the y axis (64 different neurons) should be labelled as such to ensure that neurons are not confused with trials.
+
+Also for figure 5, it would be helpful to readers to see the stimulus that drives the neurons.  Are they the same in all conditions?
+\end{quotation}
+
+We have made both these changes; they both improve the figure. The stimulus now appears at the top of the figure, and is the same for all conditions.
+
+\begin{quotation}
+In most of the figures, the authors report I (mutual information) in units of bits.
+Given the increase in firing rate in the high noise case (Fig.
+5), it may be worth reporting information in bits/spike - a qualitative look at figures 2 and 5 suggests that the central result should not be affected, and the shape of the landscape in figure 2 may be even become more peaky, further highlighting the result of the paper.
+If the authors are reporting their result in bits per spike and I missed this, then they should update the axes on the figures to clarify this.
+\end{quotation}
+
+We ran simulations measuring the information per spike, and found that the results are significantly different than when measuring the information content of the system, especially for FHN neurons.
+We have made significant revisions to the manuscript to discuss this new result.
+
+\begin{quotation}
+The authors should also make note that the information metric is calculated on the summing neuron (Stocks and Mannella, 2001).
+They do so in the methods, but a reminder would be useful for readers as a schematic in figure 1 - it could be as simple as a box diagram with stimulus going into 64 neurons, then going into a box for the decoder neuron.
+This would also help clarify and distinguish their information calculation method and subsequent result from those of Osborne et.
+al., 2008 and Padmanabhan and Urban, 2010 where the information metric is calculated across the population spike train.
+\end{quotation}
+
+We added an introductory figure depicting the experimental setup.
+
+\begin{quotation}
+A motivation should be given for the choice of 64 neurons and the authors should state in the text that the population size did not affect their central result (assuming of course that it does not).
+Population size effects can be significant when talking about information content, and while the Osborne at al 2008 result suggests that information grows with population size, it would be important to show that this holds true in the formulation selected by the authors.
+Simply put, does the landscape in Fig.
+2 change if the population is 5 neurons vs.
+50 neurons vs.
+100 cells?
+\end{quotation}
+
+We ran simulations with other numbers of neurons, and found that it does not affect the qualitative shape of the information landscapes.
+As might be expected, using fewer neurons lowers the overall information content of the system.
+We added two sentences to this regard to the experiment description at the beginning of the results section.
+
+{\large \bf Reviewer~2}
+\begin{quotation}
+  The manuscript is significantly improved and the authors have addressed all my major concerns.
+\end{quotation}
+
+{\large \bf Additional Changes}
+
+Previously, we had considered the FHN neuron model to be spiking whenever the membrane voltage was greater than zero.
+This allowed information not only to be transmitted in the occurrence and timing of spikes, but also in their amplitude and exact structure.
+This particularly became a problem at large levels of heterogeneity, which were able to push the model into a regime where it spent long periods with a positive membrane voltage, allowing direct transmission of the input signal to the output via these neurons, and ultimately resulting in inflated information transmission rates.
+We therefore revised our method of spike transmission for the FHN neuron model, so that spikes are transmitted as instantaneous stereotyped events, as with the LIF neuron model.
+This resulted in significantly different, and in our opinion more sensible, results for the FHN neuron model (see Figure 3).
+
+We hope that the changes detailed here address the concerns of the reviewers.
+We would like to thank them for their time, and for their insightful comments which have helped us improve the quality of the manuscript.
 

### 2014-01-20 paper/manuscript.tex
Third version of the paper, with figures.

@@
 
+\graphicspath{{../figures/bw/}}
 %% \newcommand{\hidefigures}{\relax}
@@
 {\bf \large Eric Hunsberger$^{\displaystyle 1}$, Matthew Scott$^{\displaystyle 2}$, Chris Eliasmith$^{\displaystyle 1}$}\\
-{$^{\displaystyle 1}$Centre for Theoretical Neuroscience, University of Waterloo, Waterloo, Ontario, Canada.}\\
-{$^{\displaystyle 2}$Department of Applied Mathematics, University of Waterloo, Waterloo, Ontario, Canada.}\\
+{$^{\displaystyle 1}$Centre for Theoretical Neuroscience, University of Waterloo, Waterloo, Canada.}\\
+{$^{\displaystyle 2}$Department of Applied Mathematics, University of Waterloo, Waterloo, Canada.}\\
 %
@@
 Noise and heterogeneity are both known to benefit neural coding.
-Stochastic resonance describes how noise, in the form of random fluctuations in a neuron's membrane voltage, can improve neural representations of an input signal.
-Neuronal heterogeneity refers to variation in any one of a number of neuron parameters, and is also known to increase the information content of a population.
-We explore the interaction between noise and heterogeneity, and find that their benefits to neural coding are not independent.
-Specifically, a neuronal population better represents an input signal when either noise or heterogeneity is added, but adding both does not always improve representation further.
-To explain this phenomenon, we propose noise and heterogeneity both operate using two shared mechanisms: 1) temporally desynchronizing the firing of neurons in the population, and 2) linearizing the response of a population to a stimulus.
-We first characterize the effects of noise and heterogeneity on the information content of populations of either leaky integrate-and-fire or FitzHugh-Nagumo neurons.
-We then examine how the mechanisms of desynchronization and linearization produce these effects, and find that they work to distribute information equally across all neurons in the population, both in terms of signal timing (desynchronization) and in terms of signal amplitude (linearization).
-Without noise or heterogeneity, all neurons in the population encode the same aspects of the input signal; adding noise or heterogeneity allows neuron to encode \emph{complementary} aspects of the input signal, thereby increasing information content.
-The simulations detailed in this paper highlight the importance of heterogeneity and noise in population coding, demonstrate their complex interactions in terms of the information content of neurons, and explain these effects in terms of underlying mechanisms.
+Stochastic resonance describes how noise,
+in the form of random fluctuations in a neuron's membrane voltage,
+can improve neural representations of an input signal.
+Neuronal heterogeneity refers to variation in any one of a number of neuron parameters,
+and is also known to increase the information content of a population.
+We explore the interaction between noise and heterogeneity
+and find that their benefits to neural coding are not independent.
+Specifically, a neuronal population better represents an input signal
+when either noise or heterogeneity is added,
+but adding both does not always improve representation further.
+To explain this phenomenon, we propose
+that noise and heterogeneity both operate using two shared mechanisms:
+1) temporally desynchronizing the firing of neurons in the population,
+and 2) linearizing the response of a population to a stimulus.
+We first characterize the effects of noise and heterogeneity
+on the information content of populations of
+either leaky integrate-and-fire or FitzHugh-Nagumo neurons.
+We then examine how the mechanisms
+of desynchronization and linearization produce these effects,
+and find that they work to distribute information equally across all neurons in the population,
+both in terms of signal timing (desynchronization) and in terms of signal amplitude (linearization).
+Without noise or heterogeneity,
+all neurons in the population encode the same aspects of the input signal;
+adding noise or heterogeneity allows neurons
+to encode \emph{complementary} aspects of the input signal,
+thereby increasing information content.
+The simulations detailed in this paper
+highlight the importance of heterogeneity and noise in population coding,
+demonstrate their complex interactions in terms of the information content of neurons,
+and explain these effects in terms of underlying mechanisms.
 
@@
 The early stochastic resonance literature dealt mainly with \emph{subthreshold} stochastic resonance \citep{Gammaitoni1998}.
-In a deterministic setting, systems which only respond to inputs above a certain threshold---such as neurons---cannot transmit any information about subthreshold input signals, \emph{i.e.,} input signals which are entirely below the system's threshold.
-In contrast, a noisy input signal (\emph{i.e,} a signal with an added stochastic offset) will occasionally cross the system's threshold, and thereby allows the system to transmit more information about the input signal.
-Too much noise, though, causes the input signal to become lost in random fluctuations.
+In a deterministic setting, systems which only respond to inputs above a certain threshold---such as neurons---cannot transmit any information about subthreshold input signals (\emph{i.e.,} input signals which are entirely below the system's threshold).
+In contrast, a noisy input signal (\emph{i.e.,} a signal with an added stochastic offset) will occasionally cross the system's threshold, and thereby allow the system to transmit more information about the input signal.
+Too much noise, however, causes the input signal to become lost in random fluctuations.
 A moderate amount of noise---not too much, and not too little---is needed to optimize the information about the input signal contained in the system \citep{Wiesenfeld1994,Longtin1998}.
@@
 In neuroscience, suprathreshold SR looks at the benefits of noise to populations of nonlinear neurons receiving a common signal of which a significant portion is above the systems' firing threshold(s).
-This concept was first introduced by \cite{Stocks2000}, who looked at groups of binary threshold units, and found that noise increases information transmission in such systems.
+This concept was first introduced by \cite{Stocks2000}, who looked at groups of binary threshold units and found that noise increases information transmission in such systems.
 Subsequent research has examined the SSR effect in populations of more sophisticated simulated neurons \citep{Stocks2001}.
@@
 Though this diversity in neural systems was documented by some of the first researchers in the field, it is only recently that neuroscientists have begun to illuminate the functional role of heterogeneity.
-Early work by \cite{Eliasmith2003} demonstrated that heterogeneity increases both the ability of a population to represent an input signal, and the range of functions computable by the population.
+Early work by \cite{Eliasmith2003} demonstrated that heterogeneity increases both the ability of a population to represent an input signal and the range of functions computable by the population.
 More recent work showed that heterogeneity allows neurons to use combinatorial coding schemes \citep{Osborne2008} and in general increase the information encoded by a population \citep{Shamir2006,Chelaru2008,Padmanabhan2010,Ecker2011}.
-In the field of stochastic resonance, the effects of noise in a heterogeneous population have been examined, but generally in specific cases such as for input signals much larger than the range of heterogeneity \citep{Stocks2000}, or for very large levels of noise \citep{McDonnell2006}.
+In the field of stochastic resonance, the effects of noise in a heterogeneous population have been examined, but only in specific cases such as for input signals much larger than the range of heterogeneity \citep{Stocks2000} or for very large levels of noise \citep{McDonnell2006}.
 Furthermore, research into stochastic resonance with heterogeneity has focused on populations of binary threshold units, which fail to capture the dynamics of actual neurons.
@@
 
-We explore this hypothesis in the context of two neuron models: the FitzHugh-Nagumo (FHN) neuron model, a common model in SR research \citep{Wiesenfeld1994,Collins1995,Longtin1998,Stocks2001}, and the leaky integrate-and-fire (LIF) neuron model, a typical model of cortical neurons \citep{Koch1999}.
-We drive the population of FHN or LIF neurons with a common aperiodic random signal $s(t)$, and vary the levels of internal noise and heterogeneity in the population.
-Finally, we decode an output signal from the population by summing and filtering the spikes from individual neurons, and compare the similarity of the decoded output signal with the input signal using mutual information.
+We explore this hypothesis in the context of two neuron models:
+the FitzHugh-Nagumo (FHN) neuron model,
+a common model in SR research \citep{Wiesenfeld1994,Collins1995,Longtin1998,Stocks2001},
+and the leaky integrate-and-fire (LIF) neuron model,
+a typical model of cortical neurons \citep{Koch1999}.
+We drive the population of FHN or LIF neurons
+with a common aperiodic random signal $s(t)$,
+and vary the levels of internal noise and heterogeneity in the population.
+Finally, we decode an output signal from the population
+by summing and filtering the spikes from individual neurons
+and compare the similarity of the decoded output signal
+with the input signal using mutual information.
 Stochastic offset is modeled as zero-mean Gaussian white noise with standard deviation $\sigma_\eta$, independently and identically distributed across all neurons (see \textsc{\nameref{scn:methods}}).
-Heterogeneity was added by randomly selecting the bias voltages $b_i$ from a uniform interval of varying width.
+Heterogeneity is added by randomly selecting
+the bias voltages $b_i$ from a uniform interval of varying width.
+This type of heterogeneity has been used in previous studies \citep[e.g.,][]{Brody2003},
+and captures the inherent variation in excitability among neurons in vivo \citep{Mejias2012} due to variations in neuron physiology or differences in the background activity of afferent neurons.
 
 We report how noise and heterogeneity both act to increase information content in these populations, but in a complex, non-independent manner.
-They achieve this using two common mechanisms---linearizing the response of a population of neurons to a stimulus, and desynchronizing neuronal firing---that we examine in detail.
+They achieve this using two common mechanisms---linearizing the response of a population of neurons to a stimulus and desynchronizing neuronal firing---that we examine in detail.
 
@@
 Noise was added to the voltage parameter of the neuron model via a stochastic offset $\eta(t)$, Gaussian distributed with zero-mean and standard deviation $\sigma_\eta$, and independent and identically distributed for each neuron.
-For the remainder of this paper, we will use the term noise to refer specifically to this stochastic offset.
-At each level of heterogeneity and noise, 25 trials were performed to achieve an accurate estimate of the mean information content, with each trial using a unique set of random biases and a unique set of stochastic offsets added to the membrane voltage.
+For the remainder of this paper, we will use the term ``noise'' to refer specifically to this stochastic offset.
+At each level of heterogeneity and noise, 100 trials were performed to achieve an accurate estimate of the mean information content, with each trial using a unique set of random biases and a unique set of stochastic offsets added to the membrane voltage.
 
-Each trial consisted of presenting a population of $N = 64$ FHN or LIF neurons with 4.5 seconds of band-limited aperiodic random signal with zero mean and standard deviation $\sigma_s = 0.1$.
-This signal has equal power below 5 Hz and zero power above 5 Hz, where the threshold of 5 Hz was chosen because it is significantly slower (approximately $\frac{1}{5}$) of the maximum firing rate of the FHN and LIF neurons (25--30 Hz).
+Each trial consisted of presenting a population of $N = 64$ FHN or LIF neurons with 4.5 seconds of aperiodic random signal with zero mean and standard deviation $\sigma_s = 0.1$.
+The aperiodic signal was created by convolving a Gaussian white noise signal with an $\alpha$ function ($\alpha(t) = (t / \tau) e^{-t / \tau}$), with the correlation time $\tau = 20$ ms \citep{Mainen1995}.
 The population was divided into 32 ``on'' neurons and 32 ``off'' neurons, where each ``on'' (``off'') neuron received a positive (negative) version of the aperiodic input signal as input.
@@
 Finally, the mutual information between the aperiodic input signal and the decoded output signal was calculated.
-The mutual information for each level of heterogeneity and noise was computed as the average mutual information from all 25 trials at those levels.
-See \textsc{\nameref{scn:methods}} for the full details of the experimental setup.
+The mutual information for each level of heterogeneity and noise was computed as the average mutual information from all 100 trials at those levels.
+We also performed experiments for different sizes of neuronal population, and found that population size does not significantly affect the results presented here.
+Using smaller populations simply lowers the overall information content of the system slightly and makes single-trial results more variable because there are fewer neurons to average across.
+The experimental setup is summarized in Figure~\ref{fig:exp}; see \textsc{\nameref{scn:methods}} for further details.
 
-Figure~\ref{fig:infohetero} summarizes the effect of neuronal heterogeneity on the information content of a population of LIF neurons.
+\begin{figure}
+  \ifx\hidefigures\undefined
+    \centering
+    \includegraphics{experiment.eps}
+  \fi
+  \caption{
+    \textbf{The experimental setup.} In our experiment, neurons are fed a common aperiodic input.
+Each neuron also has an independent white noise process and a randomly chosen bias that are added to the input.
+Varying the amplitude of the noise and the width of the interval from which the biases are randomly drawn allows us to control the amount of noise and heterogeneity in the experiment, respectively.
+Additionally, some ``off'' neurons invert their input (not shown).
+The output of each FHN or LIF neuron model is summed and filtered to produce the population output; this is compared with the input signal using mutual information as a metric.
+  }
+  \label{fig:exp}
+\end{figure}
+
+Figure~\ref{fig:infohetero} summarizes the effect of neuronal heterogeneity on the information content of a population of neurons.
 At low noise levels, we observe a resonance effect in the heterogeneity parameter---there is an optimal level of heterogeneity that maximizes the information content of the population.
-Furthermore, this optimal level of heterogeneity is around one to two times the root mean square amplitude of the input signal.
+Furthermore, this optimal level of heterogeneity is around one to two times the root-mean-square amplitude of the input signal.
 This indicates that optimal heterogeneity corresponds to having neuron firing thresholds that span the full range of input signal values.
@@
     \centering
-    \includegraphics{../figures/bw/infohetero.eps}
+    \includegraphics[width=\textwidth]{../figures/bw/infohetero.eps}
   \fi
   \caption{
-    \textbf{Heterogeneity increases information content of a population of LIF neurons, and exhibits a resonance effect.}
-    Adding heterogeneity to a population of LIF neurons increases their ability to represent an input signal, but only for lower levels of noise.
-At higher levels of noise, the noise alone is able to fully desynchronize neuronal firing and linearize neuron responses, so adding heterogeneity offers no additional benefit.
+    \textbf{Heterogeneity increases information content of a population of neurons, and exhibits a resonance effect.}
+    Adding some heterogeneity (x-axis) to a population of LIF or FHN neurons increases their ability to represent an input signal (y-axis), while too much heterogeneity decreases representational ability.
+This effect is only observed at lower noise levels; at higher levels of noise, the noise alone is able to fully desynchronize neuronal firing and linearize neuron responses, so adding heterogeneity offers no additional benefit.
   }
@@
 
-At higher levels of noise, heterogeneity offers no benefit to LIF neuron populations (Figure~\ref{fig:infohetero}).
+Higher levels of noise reduce or completely eliminate the benefits offered by heterogeneity (Figure~\ref{fig:infohetero}).
 To examine the interactions between noise and heterogeneity further, we conducted the same experiment across a wide range of noise and heterogeneity values.
 Figure~\ref{fig:infocontour} illustrates the effects of these interactions on the information content of a population of LIF or FHN neurons.
-In LIF neurons, the population is best able to represent the input signal with a moderate amount of heterogeneity ($b_r \approx 1.5 \times 10^{-1}$) and no noise ($\sigma_\eta = 1.0 \times 10^{-4}$).
-Given this optimal level of heterogeneity, adding noise only serves to decrease the information content of the system.
-But if no heterogeneity is present in the system, then adding the right amount of noise ($\sigma_\eta \approx 1.0 \times 10^{-2}$) optimizes information content (Figure~\ref{fig:infonoise}).
-This shows that the effects of noise and heterogeneity are not independent.
+We measured both the absolute information content (in bits) and the information content per spike, or information rate (in bits per spike).
+
+LIF neurons are best able to represent the input signal
+with a moderate amount of heterogeneity ($b_r \approx 1.5 \times 10^{-1}$)
+and no noise ($\sigma_\eta = 1.0 \times 10^{-4}$).
+This was true when measuring both the absolute information content and the information per spike.
+FHN neurons, on the other hand, required both heterogeneity and noise
+to achieve maximum information content,
+because heterogeneity alone is not able to fully desynchronize FHN neurons (see \textsc{\nameref{scn:desync}}).
+However, heterogeneity alone was able to maximize the information rate of the system.
+
+We explain this difference between the information content and information rate of FHN neurons
+with the fact that noise increases the firing rates of neurons.
+While noise offers some benefit to the information content of a population of FHN neurons
+(given an optimal level of heterogeneity),
+this effect disappears when examining information per spike,
+since noise also increases the number of spikes.
+In Figure~\ref{fig:infonoise}, we examine more closely the effects of noise on information content and rate in the absence of heterogeneity.
+Noise shows a clear resonance effect when measuring the absolute information content of the population, but this resonance effect is reduced when measuring the information per spike.
+We conclude that noise operates partially by increasing the firing rates of neurons, and indeed this effect can be observed when measuring the tuning curves of neurons under noise (see Figure~\ref{fig:tuning}).
+
+The information landscapes of Figure~\ref{fig:infocontour} demonstrate that the effects of noise and heterogeneity are not independent.
+This is most apparent when examining the absolute information content of LIF populations (top left), and the information rate of FHN populations (bottom right).
+For example, LIF neurons show a resonance in heterogeneity in the absence of noise,
+but they also show a resonance in noise in the absence of heterogeneity (see also Figure~\ref{fig:infonoise}).
+However, if we choose the optimal levels of noise and heterogeneity in the absence of the other factor ($\sigma_\eta \approx 10^{-2}$ and $b_r \approx 2 \times 10^{-1}$), this combined level of noise and heterogeneity does not further increase the information content of the system.
+This indicates that the effects of noise and heterogeneity are not independent: the optimal level of noise depends on the level of heterogeneity, and vice versa.
 
@@
   \caption{
-    \textbf{Noise and heterogeneity show nonlinear interactions in their benefits to information representation.} We measured the mutual information (contours every 0.15 bits) across many levels of noise and heterogeneity.
-In LIF neurons, both noise and heterogeneity show independent resonance effects: given that only one of noise or heterogeneity is present, there is a non-zero and finite level of noise or heterogeneity that optimizes information content.
-Setting both noise and heterogeneity to their optimal levels fails to further improve the representational ability of the population.
-Both parameters use the same underlying mechanisms of desynchronization and linearization, so once the population has been desynchronized and linearized by one parameter, adding the other cannot further improve information content.
-Heterogeneity is unable to fully desynchronize FHN neurons (see \textsc{\nameref{scn:desync}}), so information content in these neurons is maximized for a particular level of noise.
+    \textbf{The benefits of noise and heterogeneity to information representation are not independent.}
+We measured the information content (top, in bits) and information rate (bottom, in bits per spike),
+across many levels of noise and heterogeneity.
+In LIF neurons, both noise and heterogeneity show independent resonance effects:
+given that only one of noise or heterogeneity is present,
+there is a non-zero and finite level of noise or heterogeneity that optimizes information content.
+Setting both noise and heterogeneity to their (independent) optimal levels
+fails to further improve the representational ability of the population,
+showing that these effects are not independent.
+We propose that both parameters use
+the same underlying mechanisms of desynchronization and linearization,
+so once the population has been desynchronized and linearized by one parameter,
+adding the other cannot further improve information content.
+Heterogeneity is unable to fully desynchronize FHN neurons (see \textsc{\nameref{scn:desync}}),
+so information content in these neurons is maximized for a specific combination of heterogeneity and noise. However, this effect is not seen when examining the information rate,
+indicating that noise benefits neural systems partly through increasing firing rates.
   }
@@
   \caption{
-    \textbf{Noise increases information content of homogeneous populations of FHN and LIF neurons.} In both LIF (left) and FHN (right) populations, noise exhibits a stochastic resonance effect, where adding noise to the homogeneous population (black line) improves the population's information content.
-This stochastic resonance effect disappears in heterogeneous populations of LIF neurons, but still occurs in heterogeneous populations of FHN neurons.
+    \textbf{Noise increases information content of homogeneous populations of FHN and LIF neurons.} A) In homogeneous populations, noise exhibits a resonance effect on information content.
+B) The resonance effect of noise is less pronounced when measuring the information rate, indicating that noise operates partially by increasing firing rates.
   }
@@
 
-We propose that noise and heterogeneity both improve the information representation ability of a population of neurons by 1) desynchronizing the firing of neurons in the population, and 2) linearizing the response of the population to an input signal.
-These shared mechanisms result in the nonlinear interaction between noise and heterogeneity in terms of information content: once either noise or heterogeneity has adequately desynchronized a population, the addition of the other cannot further desynchronize or linearize, and therefore provides no additional increase in information representation.
+We propose that both noise and heterogeneity improve the information representation ability of a population of neurons by 1) desynchronizing the firing of neurons in the population and 2) linearizing the response of the population to an input signal.
+These shared mechanisms result in the non-independent interaction between noise and heterogeneity in terms of information content: once either noise or heterogeneity has adequately desynchronized a population, the addition of the other cannot further desynchronize or linearize, and therefore provides no additional increase in information representation.
 In the following two sections, we present additional experimental results to investigate and quantify the mechanisms of desynchronization and linearization.
-
-We also note that there is a difference between FHN and LIF neurons in terms of noise and heterogeneity, and their effects on information content.
-Heterogeneity is much less effective at increasing information content in FHN neurons than LIF neurons (Figures~\ref{fig:infocontour}~and~\ref{fig:infonoise}).
-This is explained by the inability of heterogeneity to fully desynchronize firing in FHN neurons (see \textsc{\nameref{scn:desync}}), which is in turn due to the nature of the FHN tuning curves.
-We revisit this result in the \textsc{\nameref{scn:discussion}}.
 
@@
     \textbf{The effects of noise and heterogeneity on LIF neuron firing.}
-    Spike rasters illustrate the effects of noise and heterogeneity parameters on neuronal firing in a population of 64 LIF neurons.
+    Spike rasters illustrate the effects of noise and heterogeneity parameters on neuronal firing in a population of 32 LIF neurons.
 With low noise and heterogeneity, neuronal firing is highly synchronized.
@@
 This previous example is, of course, artificial.
-It requires that all neurons are \emph{exactly} the same, and start with \emph{exactly} the same initial conditions.
+It requires that all neurons are \emph{exactly} the same and start with \emph{exactly} the same initial conditions.
 In a more realistic scenario, one expects that small differences in neuronal parameters or small amounts of noise will cause neurons to desynchronize over a long period of time.
@@
 FHN neurons, however, all have around the same firing rate regardless of the strength of the suprathreshold signal (Figure~\ref{fig:tuning}), and heterogeneity with higher-frequency signals has a reduced effect, since these signals cause many neurons to begin firing at once, unable to desynchronize due to their common firing rate.
-This discrepancy between LIF and FHN neurons results in heterogeneity having a large effect on phase in LIF neurons (Figure~\ref{fig:phase}A), whereas the desynchronizing effect of heterogeneity in FHN neurons is much more modest (Figure~\ref{fig:phase}B).
+This discrepancy between LIF and FHN neurons results in heterogeneity having a large effect on phase in LIF neurons (Figure~\ref{fig:phase}A), whereas the desynchronizing effect of heterogeneity in FHN neurons is more modest (Figure~\ref{fig:phase}B).
+
+This difference in the ability of heterogeneity to desynchronize LIF versus FHN neurons
+explains why LIF neurons achieve maximum information content with only heterogeneity,
+whereas FHN neurons require both heterogeneity and noise (Figure~\ref{fig:infocontour}).
+Since heterogeneity does not desynchronize FHN neurons as well as noise does,
+adding noise to a heterogeneous population of FHN neurons further desynchronizes them
+and increases information content.
 
@@
   \caption{
-    \textbf{Noise linearizes neuron tuning curves.} We probed LIF and FHN neurons using constant inputs at various voltages, using varying levels of noise, and measured the output firing rate.
-Noise has the effect of making the tuning curves more linear, specifically eliminating the sharp threshold in both LIF and FHN neurons.
-The main mechanism is that of subthreshold stochastic resonance: noise allows subthreshold inputs to cause spikes by occasionally pushing the membrane voltage above the threshold.
-This occurs in both both neuron models, allowing them to represent subthreshold signal components.
-In FHN neurons, small levels of noise can also cause firing rates for small suprathreshold signals to drop slightly by briefly pushing the neuron away from its firing limit cycle in state space, though interestingly this does not happen for larger noise levels.
-Large levels of noise in FHN neurons cause large membrane voltage fluctuations, resulting in jumps from one point on the limit cycle to another point much farther ahead, causing elevated firing rates.
-In LIF neurons, noise added to small suprathreshold signals causes the neuron to fire slightly faster, because the noise can cause the rising membrane voltage to prematurely cross the threshold, causing the spike to occasionally occur sooner than it would in the noiseless case.
+    \textbf{Noise linearizes neuron tuning curves.}
+    We measured the tuning (stimulus-response) curves of LIF and FHN neurons under various levels of noise.
+    We observe that noise linearizes the neuron tuning curves,
+    Specifically, noise eliminates the sharp threshold in the tuning curve by
+    occasionally pushing the neuron above its firing threshold
+    for stimuli that are below the threshold.
+    Noise also increases neuron firing rates across some or all stimuli.
   }
@@
 The linearizing effect of heterogeneity thereby allows for better representation of an input signal.
-This result is related to the original idea from \cite{Hubel1962} that variation in neural receptive fields provides a better basis for representing an input signal; this idea reappears in the recent heterogeneity literature \citep{Angelo2012}.
+This result is related to the original idea from \cite{Hubel1962} that variation in neural receptive fields provides a better basis for representing an input signal, and reappears in the recent heterogeneity literature \citep{Angelo2012}.
 
@@
   \caption{
-    \textbf{Heterogeneity linearizes the response of a neuronal population to a stimulus.} Panel A shows the tuning curves for a heterogeneous population of $N = 9$ neurons, where one neuron has been chosen to have a firing threshold of $-0.15$ and the other neurons have random thresholds uniformly sampled from the interval $[-0.15,0.15]$.
-Panel B shows how the population response (heavy solid line) is much closer to a linear response (heavy dot-dashed line) than a single LIF tuning curve (heavy dashed line), even though the heterogeneous thresholds are far from evenly spaced.
-The light traces indicate how the population response is the average of the individual neuron tuning curves, where each colored curve is a scaled version of the tuning curves in panel A (with scaling factor $1/N$), and the population response is the sum of these curves.
-Using a larger population of neurons (\emph{i.e.,} increasing $N$) results, on average, in a more linear population response.
+    \textbf{Heterogeneity linearizes the response of a neuronal population to a stimulus.} A) Tuning curves for a heterogeneous population of $N = 9$ neurons.
+B) The population response of the heterogeneous population (heavy solid line) is much closer to a linear response (heavy dot-dashed line) than a single LIF tuning curve (heavy dashed line), even though the heterogeneous thresholds are far from evenly spaced.
   }
@@
 Our experimental setup was motivated by stochastic resonance experiments such as \cite{Stocks2001}, as well as computational \citep{Brody2003} and experimental \citep{Padmanabhan2010} work looking at heterogeneity.
-All these experiments use a common input signal presented to a number of neurons, with added noise or heterogeneity.
-The common input signal may seem artificial, but it allows us to examine how the properties of individual neurons---including sign of the input signal (``on'' and ``off'' neurons), bias current, and membrane voltage---affect the ability of the neuronal population to encode an input signal.
+All these experiments present a common input signal to a number of neurons, with added noise or heterogeneity.
+The common input signal may seem artificial, but it allows us to examine how the properties of individual neurons---including the sign of the input signal (``on'' and ``off'' neurons), the bias current, and the membrane voltage---affect the ability of the neuronal population to encode an input signal.
 These heterogeneities in the neurons can be viewed as modulating the input signal presented to each neuron, such that each neuron receives a unique input current.
@@
 In our experiments, we saw significantly different results for the LIF neuron model versus the FHN neuron model.
-This raises two questions: why is there such a large difference between the two models, and does either model better describe the neural systems we wish to study?
-One way to differentiate between the two models is using bifurcation theory, where the LIF model is a type~I neuron model, and the FHN model is a type~II neuron model (though LIF neurons are not prototypical type~I neurons \citep{Mato2008}, they are qualitatively more similar to this category).
+Specifically, given an optimal level of heterogeneity, the FHN neurons still benefited from the addition of noise, whereas LIF neurons did not.
+This raises two questions: why is there a difference between the two models, and does either model better describe the neural systems we wish to study?
+One way to differentiate between the two models is with bifurcation theory, which classifies the LIF model is a type~I neuron model and the FHN model is a type~II neuron model (though LIF neurons are not prototypical type~I neurons \citep{Mato2008}, they are qualitatively more similar to this category).
 The difference is that type~I models undergo a saddle-node bifurcation as the input current (the bifurcation parameter) increases, whereas type~II models undergo a subcritical Hopf bifurcation \citep{Mato2008}.
-Qualitatively, this means that type~I models have continuous tuning curves, \emph{i.e.,} they can have arbitrarily small firing rates, whereas type~II models have a discontinuity at their firing threshold, and cannot fire slower than a given rate.
-The fact that the LIF model is a type~I neuron model, in addition to other factors, has traditionally made it a better model of cortical neurons \citep{Koch1999}, though studies indicate that some neurons in mammalian (rat) cortex exhibit both type~I and type~II properties \citep{Tateno2004,Tsubo2007}.
-Previous numerical experiments that have examined stochastic resonance and heterogeneity together, such as \cite{Stocks2000,McDonnell2006}, have used binary threshold units in their simulations, which are qualitatively very similar to type~II neurons (but without the temporal component); the results of these experiments may have limited applicability to cortical neural systems.
+Qualitatively, this means that type~I models have continuous tuning curves and can have arbitrarily small firing rates, whereas type~II models have a discontinuity at their firing threshold and when firing cannot fire slower than a given rate.
+The fact that the LIF model is a type~I neuron model has traditionally made it a better model of cortical neurons \citep{Koch1999}, though studies indicate that some neurons in mammalian (rat) cortex exhibit both type~I and type~II properties \citep{Tateno2004,Tsubo2007}.
+Previous numerical experiments that have examined stochastic resonance and heterogeneity together have used binary threshold units in their simulations \citep{Stocks2000,McDonnell2006}, which are qualitatively very similar to type~II neurons but without the temporal component.
+The results of these experiments may have limited applicability to cortical neural systems.
 
-We found that heterogeneity was more beneficial to LIF neurons than in FHN neurons (Figure~\ref{fig:infocontour}).
-We hypothesize that this is because of key differences in the shape of the neuron tuning curves, combined with the fact that heterogeneity was only added to one neuron parameter, namely the firing threshold $b_i$.
-The FHN model, being a type~II neuron model, has a tuning curve that is qualitatively similar to a binary threshold unit, with a discontinuous increase in firing rate at the firing threshold, and offering little change in firing rate for signals above the threshold (Figure~\ref{fig:tuning}A).
-On the other hand, the LIF model, being a type~I neuron model, has a continuous and monotonically increasing tuning curve, capable of having arbitrarily small firing rates, and producing significantly higher firing rates for input signals that are well above the firing threshold than for those slightly above the firing threshold (Figure~\ref{fig:tuning}B).
-If a high-frequency signal crosses the firing thresholds of many FHN neurons in quick succession, all these neurons will begin firing both at approximately the same time and with approximately the same rate, taking a long time to desynchronize.
+We found that heterogeneity was better able to desynchronize
+LIF neurons than FHN neurons (Figure~\ref{fig:phase}),
+given that heterogeneity was added only to the firing threshold $b_i$.
+We hypothesize that this is due to key differences in the shape of the neuron tuning curves.
+The FHN model, being a type~II neuron model,
+has a tuning curve that is qualitatively similar to a binary threshold unit,
+with a discontinuous increase in firing rate at the firing threshold
+and minimal change in firing rate for signals above the threshold (Figure~\ref{fig:tuning}A).
+On the other hand, the LIF model, being a type~I neuron model,
+has a continuous and monotonically increasing tuning curve,
+specifically one that increases smoothly with the strength of a
+suprathreshold signal (Figure~\ref{fig:tuning}B).
+If a high-frequency signal crosses the firing thresholds of many FHN neurons in quick succession, all the neurons will begin firing both at approximately the same time and with approximately the same rate, taking a long time to desynchronize.
 With LIF neurons, the high-frequency signal can still cross the firing thresholds of many neurons simultaneously, causing them to begin firing at approximately the same time.
-However, the signal will result in significantly higher firing rates for neurons with lower firing thresholds than for those with higher firing thresholds, allowing the group of neurons to desynchronize quickly.
-This key difference between the tuning curves of FHN and LIF neurons results in better desynchronization for LIF neurons when heterogeneity is only added to the firing threshold.
+However, the signal will result in higher firing rates for neurons with lower firing thresholds than for those with higher firing thresholds, allowing the group of neurons to desynchronize quickly.
+This key difference between the tuning curves of FHN and LIF neurons results in better desynchronization for LIF neurons.
 
-We hypothesize that if heterogeneity were added to other neuron parameters, then heterogeneous populations of FHN neurons would be able to contain the same amount of information as noisy homogeneous FHN populations.
+In our experiments, we added heterogeneity only to the firing threshold $b_i$.
+While this does capture the variable levels of excitability seen in in vivo neurons \citep{Mejias2012}, due to variations in neuron physiology or differences in the background activity of afferent neurons, there are clearly many other types of heterogeneity in neurons that also deserve attention.
+We hypothesize that if heterogeneity were added to other neuron parameters,
+the information content of a population of FHN neurons could be optimized without noise, as with LIF neurons.
 For example, adding variation to the maximum firing rates of FHN neurons would allow them to desynchronize as quickly as LIF neurons.
 A system with this level of heterogeneity would require a more complex decoding system than the simple summing neuron model used in this paper, perhaps using the tuning curves to solve for the optimal linear decoders as described by \cite{Salinas1994} or \cite{Eliasmith2003}.
-Such a decoding scheme could take advantage of a key difference between noise and heterogeneity: the variability produced by heterogeneity is reproducible from trial-to-trial, whereas the variability due to noise is not.
-Thus heterogeneity allows for more complex decoding schemes, such as the combinatorial decoding scheme examined in \cite{Osborne2008}.
+Such a decoding scheme could take advantage of a key difference between noise and heterogeneity: The variability produced by heterogeneity is reproducible from trial to trial, whereas the variability due to noise is not.
+Heterogeneity thereby allows for more complex decoding schemes, such as the combinatorial decoding scheme examined in \cite{Osborne2008}.
 We leave the extension of heterogeneity to additional neural parameters, and the requisite implementation of more complex decoding schemes, for future work.
 
-Noise and heterogeneity have a fundamental similarity which goes deeper than their mutual ability to increase information transmission in neuronal populations: noise and heterogeneity are both a form of random variation.
+Noise and heterogeneity have a fundamental similarity that goes deeper than their mutual ability to increase information transmission in neuronal populations, which is that noise and heterogeneity are both a form of random variation.
 Specifically, noise corresponds to variation within the phase space (state space) of a system over time, and heterogeneity corresponds to variation within the parameter space of a system across a population.
@@
 Finally, in analogy to stochastic resonance with noise, there is a critical level of heterogeneity that optimizes information transmission (Figure~\ref{fig:infohetero}).
-Thus noise and heterogeneity should be thought of as analogous mechanisms, but in different parts of the neuronal model (the state space versus the parameter space, respectively).
+Noise and heterogeneity should therefore be thought of as analogous mechanisms, but in different parts of the neuronal model (the state space versus the parameter space, respectively).
+
+Despite the similarities between noise and heterogeneity, there are also considerable differences.
+Perhaps the most important is that the response of a heterogeneous population is reproducible from trial to trial, whereas the response of a noisy population is only reproducible on average.
+Though we do not examine this effect here, due to our simplistic decoding method,
+many authors have commented the prevalence of reproducibility in neural systems \citep[e.g.,][]{Mainen1995}.
+Another significant difference between noise and heterogeneity is that noise increases the firing rate of neurons, whereas heterogeneity does not (on average).
+This results in noise being beneficial to the absolute information content of a population, but having a reduced effect on the information per spike (Figure~\ref{fig:infonoise}).
+A third difference is that under some conditions (LIF neurons) heterogeneity alone is able to optimize information content, whereas under others (FHN neurons) both heterogeneity and noise are required.
 
 Both our results here and the results of numerous other researchers suggest that noise and heterogeneity may be important vehicles used by neural systems to optimize information flow.
-Yet the extent to which actual biological neural systems exploit either heterogeneity or stochastic resonance for improved information encoding and transmission remains unclear, and may vary significantly from system to system.
+Yet the extent to which actual biological neural systems exploit either heterogeneity or stochastic resonance for improved information encoding and transmission remains unclear.
 A number of recent studies show that heterogeneity can be very important in encoding sensory information, such as for certain electric fishes \citep{Marsat2010} or in the rodent olfactory system \citep{Padmanabhan2010,Burton2012}.
-Another recent study shows SR effects in human auditory cortex, specifically that noise can increase correlation both within and between some cortical regions \citep{Ward2010}.
-More studies of this nature are required to determine both what neural systems show evidence of heterogeneity and stochastic resonance, and more importantly, whether these systems actually take advantage of heterogeneity and stochastic resonance for improved population coding.
+Another recent study shows stochastic resonance effects in human auditory cortex, specifically that noise can increase correlation both within and between some cortical regions \citep{Ward2010}.
+More studies of this nature are required to determine both what neural systems show evidence of heterogeneity or stochastic resonance, and more importantly, whether these systems actually take advantage of them for improved population coding.
+The differences between noise and heterogeneity suggest that the extent to which they are exploited may vary significantly from one neural system to another, depending on the physiology of the neurons involved and whether factors like firing rate are a significant constraint.
 The work presented here demonstrates the complex interaction between noise and heterogeneity due to the shared underlying mechanisms of desynchronization and linearization, and explains how these mechanisms lead to improved information processing characteristics.
@@
 
+Originally, we took the FHN neuron output to be equal to the positive part of the membrane voltage.
+This resulted in spikes that were not discrete, stereotyped events,
+but rather extended events that could transmit information in their amplitude.
+Although some researchers have argued that in vivo neurons do transmit information in spike amplitude,
+we found that this approach did not provide sensible results when the FHN model
+was pushed to the limits of its operating range,
+specifically for high levels of noise and wide ranges of heterogeneity.
+
+We therefore developed a basic method to determine when the FHN model was spiking,
+and transmitting an instantaneous stereotyped spike when this occurred.
+The FHN neuron model was considered to spike
+when the voltage was above the threshold $V_{th} = 0$ for a time of 1 ms,
+and was not allowed to spike again until the voltage was below the threshold for 1 ms.
+This 1 ms buffer time was necessary so that brief membrane voltage threshold crossings
+due to noise were not counted as spikes.
+
 \textbf{Leaky integrate-and-fire (LIF) model.} The LIF model describes a neuron as a membrane with resistance $R$ and capacitance $C$, which fires spikes when the membrane voltage $v(t)$ crosses the threshold $V_{th}$.
@@
 where $v_i(t)$ is the membrane voltage, $\tau_{RC} = RC$ is the membrane time constant, $\alpha$ is the input signal gain, $\beta_i$ is a constant bias voltage, $\eta_i(t)$ is a delta-correlated Gaussian white noise process (as with the FHN neuron model), and $s(t)$ is an input signal common to all neurons.
-When a neuron voltage reaches the threshold voltage $V_{th} = 1$, the neuron fires a spike, and begins a refractory period.
+When a neuron voltage reaches the threshold voltage $V_{th} = 1$, the neuron fires a spike and begins a refractory period.
 The length of the refractory period is given by the refractory time constant $\tau_{ref}$, and during this time the neuron voltage is held at zero.
@@
 
-\textbf{Input signal generation.} For the input signal $s(t)$, we used an aperiodic random signal with zero mean and equal power below a nominal maximum frequency.
-The signal had a standard deviation of $\sigma_s = 0.1$.
-It had a maximum frequency of 5Hz for the simulations, roughly corresponding to one fifth of the maximum firing frequency of the neurons over the range of inputs.
+\textbf{Input signal generation.} The input signal consisted of Gaussian white noise filtered with an $\alpha$ function, $\alpha(t) = (t / \tau_c) e^{-t / \tau_c}$, which models synaptic firing \citep{Galan2008}.
+The signal was normalized to have zero-mean and a standard deviation of $\sigma_s = 0.1$.
+The correlation time was chosen to be $\tau_c = 20$ ms, which is on the upper end of the times used by \cite{Mainen1995}.
+This correlation time results in a cutoff frequency of $f_c = 7.96$ Hz, which is considerably lower than the maximum firing rate of the neurons ($\sim 25$ Hz).
 
@@
 \begin{equation}
-  \begin{array}{rcl}
-    \tau \dot{r} & = & -r + \sum\limits_{i=1}^N q_i \vspace{0.5em}\\
-    q_i & = & \begin{cases} v_i & \text{if } v_i > V_{th} \\ 0 & \text{otherwise} \end{cases},
-  \end{array}
+  \tau \dot{r} = -r + \sum\limits_{i=1}^N s_i(t)
 \end{equation}
-where $N$ is the number of neurons in the encoding population, and $V_{th}$ is the firing threshold ($V_{th} = 0$ for the FHN model, and $V_{th} = 2$ for the LIF model).
-This model simulates a summing neuron which only receives the spikes (\emph{i.e.,} parts of the neuron voltage $v_i$ greater than the threshold $V_{th}$, and then low-pass filters the sum of these spikes, a basic description of how a summing neuron membrane sums and filters input voltages.
+where $N$ is the number of neurons in the encoding population, and $s_i(t_j)$ is 1 if neuron $i$ is spiking at time $t_j$ and 0 otherwise.
+This model simulates a summing neuron which only receives discrete, stereotyped spikes from input neurons, and then low-pass filters and sums these spikes.
+We chose a time constant $\tau = 20$ ms for the filter.
+This value was chosen to optimize the information throughput of the system; it does so likely because it is the same as the correlation time of the input signal.
+Other choices of the filter time constant produce the same qualitative results, but with lower levels of information content.
 
@@
 In the interest of simplicity, we decided to instead compare the input and output signals with no delay, as is consistent with previous experiments \citep{Stocks2001}.
+
+We obtain the information per spike by multiplying the mutual information by the Nyquist rate of the signal (2 times the signal bandwidth of $f_c = 7.96$ Hz).
+This results in an information rate (in bits per second), which we then divide by the average number of spikes per second of a neuron in the population, to obtain the information per spike.
 
@@
 
-\textbf{Simulation details.} The information transmission plots with a noise axis (Figures~\ref{fig:infocontour}~and~\ref{fig:infonoise}) were computed using noise levels in the range $\sigma_\eta \in [10^{-6},1]$ for FHN neurons and in the range $\sigma_\eta \in [10^{-4},1]$ for LIF neurons, with the levels logarithmically spaced in steps of $10^{0.1}$.
+\textbf{Simulation details.} The information transmission plots with a noise axis (Figures~\ref{fig:infocontour}~and~\ref{fig:infonoise}) were computed using noise levels in the range $\sigma_\eta \in [10^{-4},1]$, with the levels logarithmically spaced in steps of $10^{0.1}$.
 We chose the low end of these ranges to be an insignificant level of noise---where information content qualitatively the same as the noise-free case---and increasing until the noise completely overpowered the signal such that only a negligible amount of signal information could be recovered.
-Phase plots (Figure~\ref{fig:phase}) used a wider range of noise values, since desynchronization effects were observable starting at very small noise levels.
 For information plots with a heterogeneity axis (Figures~\ref{fig:infohetero}~and~\ref{fig:infocontour}), we chose $b_r \in [10^{-3}, 1]$, with points logarithmically spaced in steps of $10^{0.1}$.
@@
 
-At each combination of noise and heterogeneity levels, 25 simulations were performed, each having a simulation time of 4.5 seconds, with the first 0.5 seconds of each simulation being discarded to allow any transients created by the initial conditions to settle.
-The time step used varied depending on the noise level: larger levels of noise are more unstable and require a smaller time step, so a step of $\Delta t = 10^{-5}$ was used for $\sigma_\eta >= 0.1$, $\Delta t = 2.5 \times 10^{-5}$ for $\sigma_\eta >= 0.01$, and $\Delta t = 10^{-4}$ for $\sigma_\eta < 0.01$.
+At each combination of noise and heterogeneity levels, 100 simulations were performed, each having a simulation time of 4.5 seconds, with the first 0.5 seconds of each simulation being discarded to allow any transients created by the initial conditions to settle.
+A time step of $\Delta t = 10^{-4}$ was used.
 For each simulation, new sets of both random biases $b_i$ and noise $\eta_i(t)$ were generated for each neuron.
-The final traces in the figures show the mean over all 25 simulations.
+The final traces in the figures show the mean over all simulations.
 The standard error of the mean at each point was not plotted because it was small in all cases, with statistically significant differences between the traces.
@@
 The range of inputs measured was $[-0.2, 0.2]$, divided into 51 equally spaced points (intervals of $0.008$).
-At each point, a constant signal at that input value was presented to $N = 30$ neurons, with a duration of 5.5 seconds.
+At each point, a constant signal at that input value was presented to $N = 30$ identical neurons, with a duration of 5.5 seconds.
 The first 0.5 seconds of neuron output was discarded, to reduce the effects of initial condition transients, and the average spike rate for each neuron over the remaining time period was calculated.
-FHN neurons were considered to spike when they maintained a voltage level $V(t) > 0$ for at least 0.01 seconds, and LIF neurons were considered to spike whenever the voltage crossed the threshold $V_{th}$, as given by the neuron model.
 The tuning curves presented in the plots show the average firing rate of all $N = 30$ neurons for each input value.
@@
 
-We thank Jeff Orchard for feedback on an earlier draft of this manuscript.
+We thank the anonymous reviewers for their detailed and thoughtful comments, which greatly improved the quality of the manuscript.
+We also thank Jeff Orchard and Matthew S. Wilson for feedback on earlier drafts of the manuscript.
 
-\begin{thebibliography}{100}
-\providecommand{\natexlab}[1]{#1}
-\expandafter\ifx\csname urlstyle\endcsname\relax
-  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
-  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
-  \urlstyle{rm}\Url}\fi
+\begin{thebibliography}{}
+
+\bibitem[Angelo et~al., 2012]{Angelo2012}
+Angelo, K., Rancz, E.~a., Pimentel, D., Hundahl, C., Hannibal, J., Fleischmann,
+  A., Pichler, B., and Margrie, T.~W. (2012).
+\newblock {A biophysical signature of network affiliation and sensory
+  processing in mitral cells}.
+\newblock {\em Nature}, 488(7411):375--8.
+
+\bibitem[Brody and Hopfield, 2003]{Brody2003}
+Brody, C.~D. and Hopfield, J.~J. (2003).
+\newblock {Simple networks for spike-timing-based computation, with application
+  to olfactory processing.}
+\newblock {\em Neuron}, 37(5):843--52.
+
+\bibitem[Burton et~al., 2012]{Burton2012}
+Burton, S.~D., Ermentrout, G.~B., and Urban, N.~N. (2012).
+\newblock {Intrinsic heterogeneity in oscillatory dynamics limits
+  correlation-induced neural synchronization.}
+\newblock {\em Journal of neurophysiology}, 108(8):2115--33.
+
+\bibitem[Chance et~al., 2002]{Chance2002}
+Chance, F.~S., Abbott, L.~F., and Reyes, A.~D. (2002).
+\newblock {Gain modulation from background synaptic input.}
+\newblock {\em Neuron}, 35(4):773--82.
+
+\bibitem[Chelaru and Dragoi, 2008]{Chelaru2008}
+Chelaru, M.~I. and Dragoi, V. (2008).
+\newblock {Efficient coding in heterogeneous neuronal populations.}
+\newblock {\em Proceedings of the National Academy of Sciences},
+  105(42):16344--9.
 
@@
 \newblock {Stochastic resonance in models of neuronal ensembles}.
-\newblock {\em Physical review E}, 55(2):1798--1808.
+\newblock {\em Physical Review E}, 55(2):1798--1808.
 
@@
 \newblock {\em Nature}, 376:236--238.
+
+\bibitem[Ecker et~al., 2011]{Ecker2011}
+Ecker, A.~S., Berens, P., Tolias, A.~S., and Bethge, M. (2011).
+\newblock {The effect of noise correlations in populations of diversely tuned
+  neurons.}
+\newblock {\em The Journal of Neuroscience}, 31(40):14272--83.
 
@@
 
+\bibitem[Gal\'{a}n et~al., 2008]{Galan2008}
+Gal\'{a}n, R.~F., Ermentrout, G.~B., and Urban, N.~N. (2008).
+\newblock {Optimal time scale for spike-time reliability: theory, simulations,
+  and experiments.}
+\newblock {\em Journal of neurophysiology}, 99(1):277--83.
+
 \bibitem[Gammaitoni et~al., 1998]{Gammaitoni1998}
@@
 \newblock {\em Physical Review E}, 54(3):R2228--R2231.
+
+\bibitem[Hubel and Wiesel, 1962]{Hubel1962}
+Hubel, D. and Wiesel, T. (1962).
+\newblock {Receptive fields, binocular interaction and functional architecture
+  in the cat's visual cortex}.
+\newblock {\em The Journal of physiology}, 160:106--154.
 
@@
 
-\bibitem[Machens et~al., 2010]{Machens2010}
-Machens, C.~K., Romo, R., and Brody, C.~D. (2010).
-\newblock {Functional, but not anatomical, separation of "what" and "when" in
-  prefrontal cortex.}
-\newblock {\em The Journal of neuroscience : the official journal of the
-  Society for Neuroscience}, 30(1):350--60.
+\bibitem[Mainen and Sejnowski, 1995]{Mainen1995}
+Mainen, Z.~F. and Sejnowski, T.~J. (1995).
+\newblock {Reliability of spike timing in neocortical neurons.}
+\newblock {\em Science (New York, N.Y.)}, 268(5216):1503--6.
 
@@
   differential stimulus features}.
-\newblock {\em Neural computation}, 2440:2418--2440.
+\newblock {\em Neural Computation}, 2440:2418--2440.
 
@@
 
+\bibitem[Mejias and Longtin, 2012]{Mejias2012}
+Mejias, J. and Longtin, A. (2012).
+\newblock {Optimal Heterogeneity for Coding in Spiking Neural Networks}.
+\newblock {\em Physical Review Letters}, 108(22):228102.
+
+\bibitem[Osborne et~al., 2008]{Osborne2008}
+Osborne, L.~C., Palmer, S.~E., Lisberger, S.~G., and Bialek, W. (2008).
+\newblock {The neural basis for combinatorial coding in a cortical population
+  response.}
+\newblock {\em The Journal of Neuroscience}, 28(50):13522--31.
+
+\bibitem[Padmanabhan and Urban, 2010]{Padmanabhan2010}
+Padmanabhan, K. and Urban, N.~N. (2010).
+\newblock {Intrinsic biophysical diversity decorrelates neuronal firing while
+  increasing information content.}
+\newblock {\em Nature Neuroscience}, 13(10):1276--82.
+
+\bibitem[Prescott and {De Koninck}, 2002]{Prescott2002}
+Prescott, S.~a. and {De Koninck}, Y. (2002).
+\newblock {Gain control of firing rate by shunting inhibition: roles of
+  synaptic noise and dendritic saturation.}
+\newblock {\em Proceedings of the National Academy of Sciences},
+  100(4):2076--81.
+
+\bibitem[Salinas and Abbott, 1994]{Salinas1994}
+Salinas, E. and Abbott, L. (1994).
+\newblock {Vector reconstruction from firing rates}.
+\newblock {\em Journal of computational neuroscience}, 1(1):89--107.
+
+\bibitem[Shamir and Sompolinsky, 2006]{Shamir2006}
+Shamir, M. and Sompolinsky, H. (2006).
+\newblock {Implications of neuronal diversity on population coding.}
+\newblock {\em Neural Computation}, 18(8):1951--86.
+
+\bibitem[Shu et~al., 2003]{Shu2003}
+Shu, Y., Hasenstaub, A., Badoual, M., Bal, T., and McCormick, D.~a. (2003).
+\newblock {Barrages of synaptic activity control the gain and sensitivity of
+  cortical neurons.}
+\newblock {\em The Journal of Neuroscience}, 23(32):10388--401.
+
 \bibitem[Stocks, 2000]{Stocks2000}
@@
 
+\bibitem[Tateno et~al., 2004]{Tateno2004}
+Tateno, T., Harsch, a., and Robinson, H. P.~C. (2004).
+\newblock {Threshold firing frequency-current relationships of neurons in rat
+  somatosensory cortex: type 1 and type 2 dynamics.}
+\newblock {\em Journal of neurophysiology}, 92(4):2283--94.
+
 \bibitem[Tomita et~al., 1974]{Tomita1974}
@@
 \newblock {\em Prog. Theor. Phys}, 52(6):1744--1765.
+
+\bibitem[Tsubo et~al., 2007]{Tsubo2007}
+Tsubo, Y., Takada, M., Reyes, A.~D., and Fukai, T. (2007).
+\newblock {Layer and frequency dependencies of phase response properties of
+  pyramidal neurons in rat motor cortex.}
+\newblock {\em The European journal of neuroscience}, 25(11):3429--41.
 
@@
 
+\bibitem[Zohary et~al., 1994]{Zohary1994}
+Zohary, E., Shadlen, M., and Newsome, W. (1994).
+\newblock {Correlated neuronal discharge rate and its implications for
+  psychophysical performance}.
+\newblock {\em Nature}, 370:140--143.
+
 \end{thebibliography}

### 2014-02-04 dodo.py
Made all figures PDFs, and formatted them for publishing.

@@ -1,0 +1,141 @@
+
+from doit.action import CmdAction
+
+figure_extension = '.pdf'
+
+def _result(task, model):
+    return 'results/%s_%s.npz' % (task, model)
+
+
+def _figure(number, task):
+    return 'figures/figure%d_%s%s' % (number, task, figure_extension)
+
+
+def _simulate(task, func, model):
+    name = _result(task, model)
+    return {'basename': 'sim_%s_%s' % (task, model),
+            'actions': [(func, [model, name])],
+            # 'file_dep': ['scripts/%s.py' % task],
+            'uptodate': [lambda task, values: True],
+            'targets': [name],
+            'verbosity': 2,
+            }
+
+
+def _plot(num, task, func, deps):
+    target = _figure(num, task)
+    return {'basename': 'plot_%s' % (task),
+            'actions': [(func, [target])],
+            'file_dep': deps + ['scripts/makefigures.py'],
+            'targets': [target],
+            'doc': "Generate figure %d: %s" % (num, task),
+            'clean': True,
+            }
+
+
+def task_sim_info():
+    """Run simulations for information plots"""
+    def sim_info(model, save_name):
+        import scripts.info
+        scripts.info.run(model, save_name)
+
+    simulate = lambda m: _simulate('info', sim_info, m)
+    yield simulate('lif')
+    yield simulate('fhn')
+
+
+def task_plot_info():
+    deps = [_result('info', 'lif'), _result('info', 'fhn')]
+
+    def plot_infohetero(target):
+        import scripts.makefigures
+        scripts.makefigures.infoheteroplot(*deps, target=target)
+    yield _plot(2, 'infohetero', plot_infohetero, deps)
+
+    def plot_infocontour(target):
+        import scripts.makefigures
+        scripts.makefigures.infocontourplot(*deps, target=target)
+    yield _plot(3, 'infocontour', plot_infocontour, deps)
+
+    def plot_infonoise(target):
+        import scripts.makefigures
+        scripts.makefigures.infonoiseplot(*deps, target=target)
+    yield _plot(4, 'infonoise', plot_infonoise, deps)
+
+
+def task_sim_phase():
+    """Run simulations for phase plot"""
+    def sim_phase(model, save_name):
+        import scripts.phase
+        scripts.phase.run(model, save_name)
+
[…]

### 2014-02-04 scripts/makefigures.py

@@ -5,27 +5,34 @@ import scipy.io
 import scipy.signal
 import matplotlib
 import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
 
 from itertools import cycle
 
 ################################################################################
 ### setup
 
-inches_per_pt = 1. / 72.27               # Convert pt to inch
-inches_per_mm = 1. / 25.4
+# inches_per_pt = 1. / 72.27               # Convert pt to inch
+# inches_per_mm = 1. / 25.4
 
-onecol_in = 83.5 * inches_per_mm
-twocol_in = 173.5 * inches_per_mm
+# onecol_in = 83.5 * inches_per_mm
+# twocol_in = 173.5 * inches_per_mm
 
-params = {'axes.labelsize': 8,
-          'text.fontsize': 8,
-          'legend.fontsize': 8,
-          'xtick.labelsize': 8,
-          'ytick.labelsize': 8}
-plt.rcParams.update(params)
-
-textsize = 10
 usecolor = False
+textsize_title = 7
+textsize_small = 6
+textsize_mini = 5
+
+twocol_in = 4.25
+# onecol_in = 2.
+onecol_in = 0.5 * twocol_in
+
+params = {'axes.labelsize': textsize_small,
+          'text.fontsize': textsize_small,
+          'legend.fontsize': textsize_small,
+          'xtick.labelsize': textsize_small,
+          'ytick.labelsize': textsize_small}
+plt.rcParams.update(params)
 
 if usecolor:
     colors = [[0, 0, 0], [0, 0, 1], [0, 0.5, 0], [0.8, 0, 0]]
@@ -69,11 +76,16 @@ def makesubplot(x,y,plotfun,colors,widths,styles):
 
         l.set_dashes(next(styles))
 
-def titletext(s, ax=None, color='k', x_offset=0.0, y_offset=-0.01):
+def titletext(s, ax=None, color='k', x_offset=0.0, y_offset=0.0):
     if ax is None: ax = plt.gca()
-    ax.text(0.05+x_offset, 0.9-y_offset, s,
-            fontsize=textsize, color=color, transform=ax.transAxes,
-            backgroundcolor='white')
+    ax.text(0.01+x_offset, 1.05-y_offset, s,
+            fontsize=textsize_title, color=color, transform=ax.transAxes,)
+    # ax.text(0.06+x_offset, 0.88-y_offset, s,
[…]

### 2014-02-10 scripts/signals.py
Cleaned up signals.py

@@ -2,6 +2,7 @@
 import numpy as np
 import numpy.random as npr
 
+
 def equalpower(dt, t_final, max_freq, mean=0.0, std=1.0, n=None):
     """Generate a random signal with equal power below a maximum frequency
 
@@ -77,6 +78,7 @@ def lowpass_white(dt, t_final, tau, mean=0.0, std=1.0, n=None):
     if vector_out: return x.flatten()
     else:          return x
 
+
 def alpha_white(dt, t_final, tau, mean=0.0, std=1.0, n=None):
     """White noise filtered with the alpha function.
 
@@ -91,11 +93,9 @@ def alpha_white(dt, t_final, tau, mean=0.0, std=1.0, n=None):
     v = dt / tau
     e2v = np.exp(2*v)
     sigma = std / (v * dt) / np.sqrt(e2v*(e2v + 1)/(e2v - 1)**3)
-    # u = np.random.normal(size=(n, nt+2), scale=sigma)
     u = np.random.normal(size=(nt+2, n), scale=sigma)
 
     ### filter white noise
-    # x = np.zeros((n, nt))
     x = np.zeros((nt, n))
     x0 = np.random.normal(size=n, scale=std)
     x1 = x0.copy()
@@ -111,9 +111,6 @@ def alpha_white(dt, t_final, tau, mean=0.0, std=1.0, n=None):
                    tau/alpha**2 * (u[i+2,:] + 2*u[i+1,:] + u[i,:]))
         x1, x0 = x[i,:], x1
 
-    # if mean != 0:
-    #     x += mean
-
     ### fully normalize
     # x *= (std / x.std(-1))[...,None]
     # x += (mean - x.mean(-1))[...,None]
@@ -123,77 +120,3 @@ def alpha_white(dt, t_final, tau, mean=0.0, std=1.0, n=None):
 
     if vector_out: return x.flatten()
     else:          return x
-
-def lowpass_white_conv(dt, t_final, tau, mean=0.0, std=1.0, n=None):
-    vector_out = n is None
-    n = 1 if n is None else n
-
-    ### make white noise
-    nt = int(np.round(t_final / dt))
-    S = np.random.normal(size=(n, nt), scale=std*np.sqrt(2*tau/dt))
-    # S = np.random.normal(size=(n, nt), scale=std*2*np.sqrt(tau/dt))
-
[…]

### 2014-02-12 paper/manuscript.tex
Made all figures PDFs, and formatted them for publishing.
Some final changes to the manuscript.
- also prepping the repo for publication, with changes to the
  readme, dodo file, etc.

@@
 
-\graphicspath{{../figures/bw/}}
+\graphicspath{{../figures/}}
 %% \newcommand{\hidefigures}{\relax}
@@
 To explain this phenomenon, we propose
-that noise and heterogeneity both operate using two shared mechanisms:
+that noise and heterogeneity operate using two shared mechanisms:
 1) temporally desynchronizing the firing of neurons in the population,
@@
 Without noise or heterogeneity,
-all neurons in the population encode the same aspects of the input signal;
+all neurons encode the same aspects of the input signal;
 adding noise or heterogeneity allows neurons
@@
 A new form of stochastic resonance---called \emph{suprathreshold} stochastic resonance (SSR)---has emerged in the past decade \citep{McDonnell2009,McDonnell2011}.
-In neuroscience, suprathreshold SR looks at the benefits of noise to populations of nonlinear neurons receiving a common signal of which a significant portion is above the systems' firing threshold(s).
+In neuroscience, SSR looks at the benefits of noise to populations of nonlinear neurons receiving a common signal of which a significant portion is above the systems' firing threshold(s).
 This concept was first introduced by \cite{Stocks2000}, who looked at groups of binary threshold units and found that noise increases information transmission in such systems.
@@
 For the remainder of this paper, we will use the term ``noise'' to refer specifically to this stochastic offset.
-At each level of heterogeneity and noise, 100 trials were performed to achieve an accurate estimate of the mean information content, with each trial using a unique set of random biases and a unique set of stochastic offsets added to the membrane voltage.
+At each level of heterogeneity and noise, 100 trials were performed to achieve an accurate estimate of the mean information content, with each trial using a unique set of random biases and a unique set of stochastic offsets added to the membrane voltages.
 
@@
     \centering
-    \includegraphics{experiment.eps}
+    \includegraphics{figure1_experiment.pdf}
   \fi
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/bw/infohetero.eps}
+    \includegraphics[width=\textwidth]{figure2_infohetero.pdf}
   \fi
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/bw/infocontour.eps}
+    \includegraphics[width=\textwidth]{figure3_infocontour.pdf}
   \fi
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/bw/infonoise.eps}
+    \includegraphics[width=\textwidth]{figure4_infonoise.pdf}
   \fi
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/bw/phase.eps}
+    \includegraphics[width=\textwidth]{figure5_phase.pdf}
   \fi
@@
     \textbf{Noise and heterogeneity both desynchronize neurons.} To measure whether noise and heterogeneity have a desynchronizing effect on populations of neurons, we measured the standard deviation of neuron phases across the population (see \textsc{\nameref{scn:methods}}), where a smaller standard deviation indicates increased synchronization.
-Heterogeneity is seen to significantly desynchronize both FHN and LIF neurons, specifically across low to moderate levels of noise.
+Heterogeneity (individual traces) is seen to significantly desynchronize both FHN and LIF neurons, specifically across low to moderate levels of noise.
 At high levels of noise, the noise itself is sufficient to desynchronize neurons, and additional heterogeneity has no effect.
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/bw/syncraster.eps}
+    \includegraphics[width=\textwidth]{figure6_syncraster.pdf}
   \fi
@@
 
-Similar to noise, heterogeneity acts to desynchronize neurons \citep{Burton2012}.
-Adding heterogeneity to a population increases the range of distribution of neuron phases (Figure~\ref{fig:phase}), thereby desynchronizing populations of both FHN and LIF neurons (Figure~\ref{fig:syncraster}).
-In a homogeneous population, changes in the input signal cross the common firing threshold of all neurons at the same time, resulting in synchronous firing.
-Contrast this with a heterogeneous population, where changes in the input signal sequentially cross the thresholds of neurons, causing some to begin firing before others.
-For example, an increasing input signal will cause neurons with lower thresholds to begin firing before those with higher thresholds, resulting in desynchronized firing within the population of neurons.
+Similar to noise,
+heterogeneity acts to desynchronize neurons \citep{Burton2012}.
+Adding heterogeneity to a population
+increases the range of distribution of neuron phases (Figure~\ref{fig:phase}),
+indicating desynchronized firing
+for both FHN and LIF neurons (Figure~\ref{fig:syncraster}).
+Homogeneous populations exhibit synchronized firing
+because changes in the input signal cross the common firing threshold
+of all neurons at the same time.
+Contrast this with a heterogeneous population,
+where changes in the input signal
+sequentially cross the thresholds of neurons,
+causing some to begin firing before others.
+For example, an increasing input signal
+will cause neurons with lower thresholds
+to begin firing before those with higher thresholds,
+resulting in desynchronized firing within the population of neurons.
 
-High frequency input signals may cross the thresholds of many neurons in a heterogeneous population at the same time, causing a large group of neurons to begin firing simultaneously.
-LIF neurons are still able to desynchronize quickly after such a stimulus, since the LIF tuning curves used have different firing rates for different sizes of suprathreshold signal (Figure~\ref{fig:tuning}).
-Specifically, a high-frequency increase in signal amplitude starts neurons firing at approximately the same time, but starts them firing at different rates, allowing them to quickly desynchronize.
-FHN neurons, however, all have around the same firing rate regardless of the strength of the suprathreshold signal (Figure~\ref{fig:tuning}), and heterogeneity with higher-frequency signals has a reduced effect, since these signals cause many neurons to begin firing at once, unable to desynchronize due to their common firing rate.
-This discrepancy between LIF and FHN neurons results in heterogeneity having a large effect on phase in LIF neurons (Figure~\ref{fig:phase}A), whereas the desynchronizing effect of heterogeneity in FHN neurons is more modest (Figure~\ref{fig:phase}B).
+High frequency input signals may cross the thresholds of many neurons
+in a heterogeneous population in a short period of time,
+causing a large group of neurons to begin firing simultaneously.
+LIF neurons are still able to desynchronize quickly after such a stimulus,
+since the LIF tuning curves used have different firing rates
+for different sizes of suprathreshold signal (Figure~\ref{fig:tuning}).
+Specifically, a high-frequency increase in signal amplitude
+starts neurons firing at approximately the same time,
+but starts them firing at different rates,
+allowing them to quickly desynchronize.
+FHN neurons, however, all have around the same firing rate
+regardless of the strength
+of the suprathreshold signal (Figure~\ref{fig:tuning}).
+In the presence of higher-frequency input signals
+heterogeneity has a reduced effect in these neurons,
+since such input signals cause many neurons to begin firing at once,
+and they are unable to desynchronize due to their common firing rate.
+This discrepancy between LIF and FHN neurons results in heterogeneity
+having a large effect on phase in LIF neurons (Figure~\ref{fig:phase}A),
+whereas the desynchronizing effect of heterogeneity
+in FHN neurons is more modest (Figure~\ref{fig:phase}B).
 
@@
 
-Intrinsic noise makes neuron tuning curves more linear;
+Intrinsic noise makes neuron tuning curves more linear
+(Figure~\ref{fig:tuning});
 this phenomenon underlies many results in the stochastic resonance literature
@@
 In type~II neuron models (\emph{e.g.} the FHN model),
-noise can also slow firing for signals slightly above the firing threshold,
-by momentarily pushing the neuron away from its limit cycle and back to near its unstable equilibrium.
+noise can also slow firing for signals slightly above the firing threshold
+by momentarily pushing the neuron away from its limit cycle
+and back toward its weakly unstable equilibrium.
 Overall, the addition of noise has the effect
 of linearizing the response of the neuron near its firing threshold,
-resulting in a smoother transition from the silent to tonic firing regimes (Figure~\ref{fig:tuning}).
+resulting in a smoother transition from the silent to tonic firing regimes.
 This result has been previously observed and extensively studied \citep{Stocks1996,Chialvo1997,Chance2002,Prescott2002,Shu2003}.
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/bw/tuningnoisy.eps}
+    \includegraphics[width=\textwidth]{figure7_tuningnoisy.pdf}
   \fi
@@
 
-Neuron heterogeneity has a linearizing effect similar to adding noise to a population, but unlike the noisy case, where it made sense to talk about noise linearizing an individual neuron's tuning curve, here we will discuss how heterogeneity linearizes the \emph{population's} tuning curve.
-The population's tuning curve is the set of points connecting the input signal strength to the decoded output.
-Since our model uses a simple summing decoder, the population tuning curve is equal to the sum of the individual neuron tuning curves.
-For a population of neurons with uniformly varied random firing thresholds, the population tuning curve is much more linear than a single neuron's tuning curve, whereas in the homogeneous case the population's tuning curve is the same as the neurons' tuning curve (Figure~\ref{fig:tuninghetero}).
-The linearizing effect of heterogeneity thereby allows for better representation of an input signal.
-This result is related to the original idea from \cite{Hubel1962} that variation in neural receptive fields provides a better basis for representing an input signal, and reappears in the recent heterogeneity literature \citep{Angelo2012}.
+Neuron heterogeneity has a linearizing effect similar to noise,
+but unlike the noisy case,
+where it made sense to talk about noise
+linearizing an individual neuron's tuning curve,
+here we will discuss how heterogeneity
+linearizes the \emph{population's} tuning curve.
+The population's tuning curve is the set of points
+connecting the input signal strength to the decoded output.
+Since our model uses a simple summing decoder,
+the population tuning curve is equal to
+the sum of the individual neuron tuning curves.
+For a population of neurons with uniformly varied random firing thresholds,
+the population tuning curve is much more linear
+than a single neuron's tuning curve,
+whereas in the homogeneous case the population's tuning curve
+is the same as the neurons' tuning curve (Figure~\ref{fig:tuninghetero}).
+The linearizing effect of heterogeneity
+thereby allows for better representation of an input signal.
+This result is related to the original idea from \cite{Hubel1962}
+that variation in neural receptive fields
+provides a better basis for representing an input signal,
+and reappears in the recent heterogeneity literature \citep{Angelo2012}.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/bw/tuninghetero.eps}
+    \includegraphics[width=\textwidth]{figure8_tuninghetero.pdf}
   \fi
@@
 
-Desynchronization operates by affecting the temporal properties of neuronal firing, and is only observed in spiking neuron models, such as the FHN and LIF models examined in this paper.
+Desynchronization operates by affecting the temporal properties of neuronal firing, and is only observed in spiking neuron models such as the FHN and LIF models examined in this paper.
 The output of these models depends not only on the the output firing rate for a given input signal, as determined by the neuron tuning curve, but also by the timing of individual spike events.
@@
 All these experiments present a common input signal to a number of neurons, with added noise or heterogeneity.
-The common input signal may seem artificial, but it allows us to examine how the properties of individual neurons---including the sign of the input signal (``on'' and ``off'' neurons), the bias current, and the membrane voltage---affect the ability of the neuronal population to encode an input signal.
+The common input signal may seem artificial, but it allows us to examine how the properties of individual neurons---including the sign of the input signal (``on'' versus ``off'' neurons) and the bias current---affect the ability of the neuronal population to encode an input signal.
 These heterogeneities in the neurons can be viewed as modulating the input signal presented to each neuron, such that each neuron receives a unique input current.
@@
 \end{equation}
-where $v_i(t)$ is a fast (voltage) variable, $w_i(t)$ is a slow (recovery) variable, $\beta_i$ is a constant bias voltage, $e_i$ is a simple encoder equal to 1 for ``on'' neurons and $-1$ for ``off'' neurons, $s(t)$ is an input signal common to all neurons, and $\eta_i(t)$ is a Gaussian white noise process with autocorrelation $\left<\eta_i(t)\eta_j(t + \tau)\right> = \sigma^2 \delta_{ij}\delta(\tau)$.
+where $v_i(t)$ is a fast (voltage) variable, $w_i(t)$ is a slow (recovery) variable, $\beta_i$ is a constant bias voltage, $e_i$ is a simple encoder equal to 1 for ``on'' neurons and $-1$ for ``off'' neurons, $s(t)$ is an input signal common to all neurons, and $\eta_i(t)$ is a Gaussian white noise process with autocorrelation $\left<\eta_i(t)\eta_j(t + \tau)\right> = \sigma_\eta^2 \delta_{ij}\delta(\tau)$.
 We chose $\beta_i = 0.3216 - b_i$, with the offset 0.3216 chosen such that the bias $b_i$ corresponds to the threshold at which the neuron begins firing.
@@
 In the interest of simplicity, we decided to instead compare the input and output signals with no delay, as is consistent with previous experiments \citep{Stocks2001}.
+In preliminary experiments, accounting for delay resulted in only a small increase in information.
 
@@
 
-\textbf{Heterogeneity.} To introduce heterogeneity into the population, we varied the bias voltages of the neurons.
+\textbf{Heterogeneity.} To introduce heterogeneity into the population, we varied the bias voltages $b_i$ of the neurons.
 In homogeneous populations, all neurons had bias voltages equal to zero, the mean of the input signal.
-Heterogeneous populations had neurons with uniformly randomly distributed biases chosen from a range centered around the mean.
-In particular, three ranges were used: $[-0.01,0.01]$, $[-0.05,0.05]$, and $[-0.2,0.2]$, corresponding to small, medium and large levels of heterogeneity.
+Heterogeneous populations had neurons with uniformly randomly distributed biases chosen from a range $\mathcal{U}(-b_r, b_r)$, where $b_r$ is the radius of heterogeneity.
+For the phase plot (Figure~\ref{fig:phase}), three ranges were used: $b_r = 0.01, 0.05, 0.2$, corresponding to small, medium and large levels of heterogeneity.
 The ranges were chosen to be roughly logarithmically distributed, with the largest range corresponding to biases that covered two standard deviations of the input signal.
@@
 In terms of our model, this means that half the neurons of each population had $e_i = 1$, and the other half had $e_i = -1$.
-This means that half the population represents the positive and negative parts of the signal, respectively, and only a small amount of information is gained by the subthreshold SR effect, which allows each half of the population to represent both positive and negative signal components.
+This means that half the population represents the positive and negative parts of the signal, respectively, and only a small amount of information is gained by the subthreshold SR effect which allows each half of the population to represent both positive and negative signal components.
 
@@
 \begin{equation}
-  \bar\theta(t) = \tan^{-1} \left( \frac{1}{N}\sum\limits_{i=1}^N \cos\theta_i(t),
-                                  \frac{1}{N}\sum\limits_{i=1}^N \sin\theta_i(t) \right),
+  \bar\theta(t) = \tan^{-1} \left(
+  \frac{1}{N}\sum\limits_{i=1}^N \cos\left(\theta_i(t)\right),
+  \frac{1}{N}\sum\limits_{i=1}^N \sin\left(\theta_i(t)\right) \right),
 \end{equation}
@@
 
-At each combination of noise and heterogeneity levels, 100 simulations were performed, each having a simulation time of 4.5 seconds, with the first 0.5 seconds of each simulation being discarded to allow any transients created by the initial conditions to settle.
+For the information and phase plots (Figures~\ref{fig:infohetero}--\ref{fig:phase}), 100 simulations were performed at each combination of noise and heterogeneity levels, each having a simulation time of 4.5 seconds.
+Initial conditions were chosen randomly from across the neuron's working phase space (FHN neurons: $v_i \in [-2, 2]$, $w_i \in [-0.4, 1.2]$; LIF neurons: $v_i \in [0, V_{th}]$), with the first 0.5 seconds of each simulation being discarded to allow any transients created by the initial conditions to settle.
 A time step of $\Delta t = 10^{-4}$ was used.
@@
 The standard error of the mean at each point was not plotted because it was small in all cases, with statistically significant differences between the traces.
+
+Simulations for the firing rate plots (Figure~\ref{fig:syncraster}) were conducted in the same manner as for the information and phase plots, except that the figure shows the results of single trials, not averaged trials.
 

### 2014-02-12 dodo.py

@@ -1,14 +1,27 @@
 
+import os
+
+from doit import get_var
 from doit.action import CmdAction
 
 figure_extension = '.pdf'
 
+### Process arguments
+gpu = get_var('gpu', 'False')
+if gpu in ['1', 'true', 'True']:
+    os.environ['THEANO_FLAGS'] = 'device=gpu,floatX=float32'
+elif gpu not in ['0', 'false', 'False']:
+    raise ValueError("Unrecognized value '%s' for parameter 'gpu'" % gpu)
+
+
+### Helper functions
 def _result(task, model):
-    return 'results/%s_%s.npz' % (task, model)
+    return os.path.join('results', '%s_%s.npz' % (task, model))
 
 
 def _figure(number, task):
-    return 'figures/figure%d_%s%s' % (number, task, figure_extension)
+    return os.path.join(
+        'figures', 'figure%d_%s%s' % (number, task, figure_extension))
 
 
 def _simulate(task, func, model):
@@ -26,13 +39,14 @@ def _plot(num, task, func, deps):
     target = _figure(num, task)
     return {'basename': 'plot_%s' % (task),
             'actions': [(func, [target])],
-            'file_dep': deps + ['scripts/makefigures.py'],
+            'file_dep': deps + [os.path.join('scripts', 'makefigures.py')],
             'targets': [target],
             'doc': "Generate figure %d: %s" % (num, task),
             'clean': True,
             }
 
 
+### Tasks
 def task_sim_info():
     """Run simulations for information plots"""
     def sim_info(model, save_name):
@@ -134,8 +148,8 @@ def task_plot_tuninghetero():
 def task_paper():
     """Generate a PDF of the paper using pdflatex"""
     pdf = CmdAction(
-        'pdflatex -interaction=batchmode manuscript.tex', cwd='paper/')
+        'pdflatex -interaction=batchmode manuscript.tex', cwd='paper')
     return {'actions': [pdf, pdf, pdf],
-            'file_dep': ['paper/manuscript.tex'],
-            'targets': ['paper/manuscript.pdf'],
+            'file_dep': [os.path.join('paper', 'manuscript.tex')],
+            'targets': [os.path.join('paper', 'manuscript.pdf')],
             'clean': True}

### 2014-02-12 scripts/makefigures.py

@@ -1,7 +1,6 @@
 
 import numpy as np
 import scipy as sp
-import scipy.io
 import scipy.signal
 import matplotlib
 import matplotlib.pyplot as plt
@@ -302,9 +301,9 @@ def phaseplot(lifphasefile, fhnphasefile, target):
     for b, b2 in zip(lifdata['bths'], fhndata['bths']):
         assert b == b2
         if abs(b) < 1e-8:
-            legendstr.extend( [r'$b_i =$ $0$'] )
+            legendstr.extend( [r'$b_r =$ $0$'] )
         else:
-            legendstr.extend( [r'$b_i \in [%0.2f,%0.2f]$' % (b,b)] )
+            legendstr.extend( [r'$b_r =$ $%0.2f$' % b] )
 
     plt.figure(2,figsize=(twocol_in,onecol_in))
     plt.clf()
@@ -388,7 +387,7 @@ def syncrasterplot(firingfile, target):
             for t in times:
                 lines = ax.plot(t, height*np.ones(len(t)), 'k|')
                 lines[0].set_markersize(1)
-                lines[0].set_markeredgewidth(2)
+                lines[0].set_markeredgewidth(1)
                 height += 1
 
             if i == len(Nstds) - 1:

### 2014-02-13 paper/manuscript.tex
Made paper buildable without figures
- also added corresponding author information
- minor README formatting changes

@@
+%%%%% options, to be defined externally
+%% \newcommand{\hidefigures}{\relax}
+
 \documentclass[12pt]{article}
@@
 \graphicspath{{../figures/}}
-%% \newcommand{\hidefigures}{\relax}
 
@@
 \ \\
-{\bf \large Eric Hunsberger$^{\displaystyle 1}$, Matthew Scott$^{\displaystyle 2}$, Chris Eliasmith$^{\displaystyle 1}$}\\
-{$^{\displaystyle 1}$Centre for Theoretical Neuroscience, University of Waterloo, Waterloo, Canada.}\\
-{$^{\displaystyle 2}$Department of Applied Mathematics, University of Waterloo, Waterloo, Canada.}\\
+{\bf \large Eric Hunsberger$^{1*}$, Matthew Scott$^2$, Chris Eliasmith$^1$}\\
+{$^1$Centre for Theoretical Neuroscience, University of Waterloo, Waterloo, Canada.}\\
+{$^2$Department of Applied Mathematics, University of Waterloo, Waterloo, Canada.}\\
+{$^*$Corresponding author: ehunsber@uwaterloo.ca}\\
 %

### 2014-02-13 dodo.py

@@ -7,11 +7,23 @@ from doit.action import CmdAction
 figure_extension = '.pdf'
 
 ### Process arguments
-gpu = get_var('gpu', 'False')
-if gpu in ['1', 'true', 'True']:
+def _get_boolean_argument(name, default=None):
+    value = get_var(name, default)
+    if default is None and value is None:
+        raise ValueError("Required argument '%s' not provided" % name)
+
+    value = str(value).lower()
+    if value in ['1', 'true', 'yes', 'y', True]:
+        return True
+    if value in ['0', 'false', 'no', 'n', False]:
+        return False
+    raise ValueError(
+        "Unrecognized value '%s' for parameter '%s'" % (value, name))
+
+
+gpu = _get_boolean_argument('gpu', False)
+if gpu:
     os.environ['THEANO_FLAGS'] = 'device=gpu,floatX=float32'
-elif gpu not in ['0', 'false', 'False']:
-    raise ValueError("Unrecognized value '%s' for parameter 'gpu'" % gpu)
 
 
 ### Helper functions
@@ -147,8 +159,13 @@ def task_plot_tuninghetero():
 
 def task_paper():
     """Generate a PDF of the paper using pdflatex"""
+    command = ""
+    if not _get_boolean_argument('figures', True):
+        command += "\\newcommand{\\hidefigures}{\\relax}"
+    command += "\\input{manuscript.tex}"
+
     pdf = CmdAction(
-        'pdflatex -interaction=batchmode manuscript.tex', cwd='paper')
+        'pdflatex -interaction=batchmode "%s"' % command, cwd='paper')
     return {'actions': [pdf, pdf, pdf],
             'file_dep': [os.path.join('paper', 'manuscript.tex')],
             'targets': [os.path.join('paper', 'manuscript.pdf')],

### 2014-06-25 paper/manuscript.tex
Fixes to the figureless paper.
- Also some minor cosmetic edits to the references
Made a nicer preprint manuscript

@@
 
-\documentclass[12pt]{article}
+\documentclass[letterpaper,11pt]{article}
+
 \usepackage{amsmath}
@@
 \usepackage[authoryear]{natbib}
+\usepackage{nameref}
 
@@
 
-%%% margins
-\textheight 23.4cm
-\textwidth 14.65cm
-\oddsidemargin 0.375in
-\evensidemargin 0.375in
-\topmargin  -0.55in
+\usepackage{url}
+\renewcommand*{\UrlFont}{\ttfamily\small\relax}
 
-\renewcommand{\baselinestretch}{2}
-
-\interfootnotelinepenalty=10000
-
-% Different font in captions
-\newcommand{\captionfonts}{\normalsize}
-
-\makeatletter
-\long\def\@makecaption#1#2{%
-  \renewcommand{\baselinestretch}{1}% EH: stop captions running off the bottom of the page
-  \vskip\abovecaptionskip
-  \sbox\@tempboxa{{\captionfonts #1: #2}}%
-  \ifdim \wd\@tempboxa >\hsize
-    {\captionfonts #1: #2\par}
-  \else
-    \hbox to\hsize{\hfil\box\@tempboxa\hfil}%
-  \fi
-  \vskip\belowcaptionskip}
-\makeatother
-%%%%%
-
-\usepackage{nameref}
+\usepackage[top=1in, bottom=1in, left=1.5in, right=1.5in]{geometry}
 
@@
 
+% Why does higher tolerance stop the overfull lines? But it does!
+\tolerance=10000
+
+\newcommand{\horrule}[1]{\rule{\linewidth}{#1}} 	% Horizontal rule
+
+\title{
+  \vspace{-0.5in}
+  \horrule{3pt} \\[0.4cm]
+  \bf The competing benefits of noise and heterogeneity\\in neural coding
+  \horrule{1pt}
+}
+
+\newcommand{\makeauthor}[5]{
+  \normalsize
+  \begin{tabular}{c}
+    \bf #1\\
+    #2\\
+    #3\\
+    #4\\
+    #5
+  \end{tabular}
+}
+
+\author{
+  \hspace{-1em}
+  \makeauthor{Eric Hunsberger}
+             {Computational Neuroscience Research Group}
+             {University of Waterloo}
+             {Waterloo, ON, Canada}
+             {\url{ehunsber@uwaterloo.ca}}
+  \makeauthor{Matthew Scott}
+             {Department of Applied Mathematics}
+             {University of Waterloo}
+             {Waterloo, ON, Canada}
+             {\relax} \\ \\
+  \hspace{-1em}
+  \makeauthor{Chris Eliasmith}
+             {Computational Neuroscience Research Group}
+             {University of Waterloo}
+             {Waterloo, ON, Canada}
+             {}
+}
+
+\date{\today}
+
 \begin{document}
-\hspace{13.9cm}
 
-\ \vspace{20mm}\\
+\maketitle
 
-{\LARGE The competing benefits of noise and heterogeneity in neural coding}
-
-\ \\
-{\bf \large Eric Hunsberger$^{1*}$, Matthew Scott$^2$, Chris Eliasmith$^1$}\\
-{$^1$Centre for Theoretical Neuroscience, University of Waterloo, Waterloo, Canada.}\\
-{$^2$Department of Applied Mathematics, University of Waterloo, Waterloo, Canada.}\\
-{$^*$Corresponding author: ehunsber@uwaterloo.ca}\\
-%
-
-%\ \\[-2mm]
-{\bf Keywords:} Heterogeneity; stochastic resonance; population coding
-
-\thispagestyle{empty}
-\markboth{}{NC instructions}
-%
-\ \vspace{-0mm}\\
-%
-%Abstract
 \begin{center} {\bf Abstract} \end{center}
@@
 and explain these effects in terms of underlying mechanisms.
+
+{\bf Keywords:} Heterogeneity; stochastic resonance; population coding
+
+{\bf Source code:} \url{https://github.com/ctn-archive/hunsberger-neco2014}
 
@@
     \includegraphics{figure1_experiment.pdf}
+  \else
+    INSERT FIGURE 1 HERE
   \fi
@@
     \includegraphics[width=\textwidth]{figure2_infohetero.pdf}
+  \else
+    INSERT FIGURE 2 HERE
   \fi
@@
 
-LIF neurons are best able to represent the input signal
-with a moderate amount of heterogeneity ($b_r \approx 1.5 \times 10^{-1}$)
-and no noise ($\sigma_\eta = 1.0 \times 10^{-4}$).
-This was true when measuring both the absolute information content and the information per spike.
-FHN neurons, on the other hand, required both heterogeneity and noise
-to achieve maximum information content,
-because heterogeneity alone is not able to fully desynchronize FHN neurons (see \textsc{\nameref{scn:desync}}).
-However, heterogeneity alone was able to maximize the information rate of the system.
-
-We explain this difference between the information content and information rate of FHN neurons
-with the fact that noise increases the firing rates of neurons.
-While noise offers some benefit to the information content of a population of FHN neurons
-(given an optimal level of heterogeneity),
-this effect disappears when examining information per spike,
-since noise also increases the number of spikes.
-In Figure~\ref{fig:infonoise}, we examine more closely the effects of noise on information content and rate in the absence of heterogeneity.
-Noise shows a clear resonance effect when measuring the absolute information content of the population, but this resonance effect is reduced when measuring the information per spike.
-We conclude that noise operates partially by increasing the firing rates of neurons, and indeed this effect can be observed when measuring the tuning curves of neurons under noise (see Figure~\ref{fig:tuning}).
-
-The information landscapes of Figure~\ref{fig:infocontour} demonstrate that the effects of noise and heterogeneity are not independent.
-This is most apparent when examining the absolute information content of LIF populations (top left), and the information rate of FHN populations (bottom right).
-For example, LIF neurons show a resonance in heterogeneity in the absence of noise,
-but they also show a resonance in noise in the absence of heterogeneity (see also Figure~\ref{fig:infonoise}).
-However, if we choose the optimal levels of noise and heterogeneity in the absence of the other factor ($\sigma_\eta \approx 10^{-2}$ and $b_r \approx 2 \times 10^{-1}$), this combined level of noise and heterogeneity does not further increase the information content of the system.
-This indicates that the effects of noise and heterogeneity are not independent: the optimal level of noise depends on the level of heterogeneity, and vice versa.
-
 \begin{figure}
@@
     \includegraphics[width=\textwidth]{figure3_infocontour.pdf}
+  \else
+    INSERT FIGURE 3 HERE
   \fi
@@
 
+LIF neurons are best able to represent the input signal
+with a moderate amount of heterogeneity ($b_r \approx 1.5 \times 10^{-1}$)
+and no noise ($\sigma_\eta = 1.0 \times 10^{-4}$).
+This was true when measuring both the absolute information content and the information per spike.
+FHN neurons, on the other hand, required both heterogeneity and noise
+to achieve maximum information content,
+because heterogeneity alone is not able to fully desynchronize FHN neurons (see \textsc{\nameref{scn:desync}}).
+However, heterogeneity alone was able to maximize the information rate of the system.
+
+We explain this difference between the information content and information rate of FHN neurons
+with the fact that noise increases the firing rates of neurons.
+While noise offers some benefit to the information content of a population of FHN neurons
+(given an optimal level of heterogeneity),
+this effect disappears when examining information per spike,
+since noise also increases the number of spikes.
+In Figure~\ref{fig:infonoise}, we examine more closely the effects of noise on information content and rate in the absence of heterogeneity.
+Noise shows a clear resonance effect when measuring the absolute information content of the population, but this resonance effect is reduced when measuring the information per spike.
+We conclude that noise operates partially by increasing the firing rates of neurons, and indeed this effect can be observed when measuring the tuning curves of neurons under noise (see Figure~\ref{fig:tuning}).
+
 \begin{figure}
@@
     \includegraphics[width=\textwidth]{figure4_infonoise.pdf}
+  \else
+    INSERT FIGURE 4 HERE
   \fi
@@
 \end{figure}
+
+The information landscapes of Figure~\ref{fig:infocontour} demonstrate that the effects of noise and heterogeneity are not independent.
+This is most apparent when examining the absolute information content of LIF populations (top left), and the information rate of FHN populations (bottom right).
+For example, LIF neurons show a resonance in heterogeneity in the absence of noise,
+but they also show a resonance in noise in the absence of heterogeneity (see also Figure~\ref{fig:infonoise}).
+However, if we choose the optimal levels of noise and heterogeneity in the absence of the other factor ($\sigma_\eta \approx 10^{-2}$ and $b_r \approx 2 \times 10^{-1}$), this combined level of noise and heterogeneity does not further increase the information content of the system.
+This indicates that the effects of noise and heterogeneity are not independent: the optimal level of noise depends on the level of heterogeneity, and vice versa.
 
@@
     \includegraphics[width=\textwidth]{figure5_phase.pdf}
+  \else
+    INSERT FIGURE 5 HERE
   \fi
@@
     \includegraphics[width=\textwidth]{figure6_syncraster.pdf}
+  \else
+    INSERT FIGURE 6 HERE
   \fi
@@
     \includegraphics[width=\textwidth]{figure7_tuningnoisy.pdf}
+  \else
+    INSERT FIGURE 7 HERE
   \fi
@@
     \includegraphics[width=\textwidth]{figure8_tuninghetero.pdf}
+  \else
+    INSERT FIGURE 8 HERE
   \fi
@@
   correlation-induced neural synchronization.}
-\newblock {\em Journal of neurophysiology}, 108(8):2115--33.
+\newblock {\em Journal of Neurophysiology}, 108(8):2115--33.
 
@@
   membrane}.
-\newblock {\em Biophysical journal}, 1(1948):445--466.
+\newblock {\em Biophysical Journal}, 1(1948):445--466.
 
@@
   and experiments.}
-\newblock {\em Journal of neurophysiology}, 99(1):277--83.
+\newblock {\em Journal of Neurophysiology}, 99(1):277--83.
 
@@
   in the cat's visual cortex}.
-\newblock {\em The Journal of physiology}, 160:106--154.
+\newblock {\em The Journal of Physiology}, 160:106--154.
 
@@
   communication signals.}
-\newblock {\em Journal of neurophysiology}, 104(5):2543--55.
+\newblock {\em Journal of Neurophysiology}, 104(5):2543--55.
 
@@
 \newblock {Vector reconstruction from firing rates}.
-\newblock {\em Journal of computational neuroscience}, 1(1):89--107.
+\newblock {\em Journal of Computational Neuroscience}, 1(1):89--107.
 
@@
   systems}.
-\newblock {\em Physical review letters}, 84(11):2310--3.
+\newblock {\em Physical Review Letters}, 84(11):2310--3.
 
@@
   somatosensory cortex: type 1 and type 2 dynamics.}
-\newblock {\em Journal of neurophysiology}, 92(4):2283--94.
+\newblock {\em Journal of Neurophysiology}, 92(4):2283--94.
 
@@
 \newblock {Irreversible circulation and orbital revolution}.
-\newblock {\em Prog. Theor. Phys}, 52(6):1744--1765.
+\newblock {\em Progress of Theoretical Physics}, 52(6):1744--1765.
 
@@
   pyramidal neurons in rat motor cortex.}
-\newblock {\em The European journal of neuroscience}, 25(11):3429--41.
+\newblock {\em European Journal of Neuroscience}, 25(11):3429--41.
 
